### 2013-04-04 metalindelof_not_metacompact.tex
Lots of uncategorized changes

\input{./header.tex}

\begin{document}

\begin{example}
  Let $X$ be a compact, zero-dimensional space with a point-countable cover $\{U_\alpha : \alpha \in \Lim(\omega_1)\}$ which is not $\sigma$-point-finite. (Can we change this to, every uncountable subcollection is not point-finite, or every stationary subset $S$ of $\Lim(\omega_1)$ doesn't yield a point-finite subcollection?)

  For $\alpha \in \Lim(\omega_1)$, pick $\beta_{\alpha,n}$ such that $\ds\lim_{n\to\infty}\beta_{\alpha,n}=\alpha$. Assume every $\beta\in\Suc(\omega_1)$ is $\beta_{\alpha,n}$ for some $\alpha,n$. Let $\mc{B}_\alpha = \{\beta_{\alpha,0},\beta_{\alpha,1},\dots\}$

  Let $\mathbb{X} = \Lim(\omega_1) \cup (X \times \Suc(\omega_1))$, with the topology induced by $X\times\{\beta\}$ being open and a topological copy of $X$, and each $\alpha$ given the open neighborhood $B_\alpha = \{\alpha\}\cup(U_\alpha \times \mc{B}_\alpha)$.
\end{example}

\begin{theorem}
  $K\win\lfkpgame{\mathbb{X}}$.
\end{theorem}

\begin{proof}
  Idea: Seeing $\<x_i,\beta_i\>$, $K$ forbids $X\times \{\beta_i\}$, and promises to eventually forbid $B_\alpha$ for each $\alpha$ where $x_i \in U_\alpha$.
\end{proof}

\begin{theorem}
  $K\not\tactwin\lfkpgame{\mathbb{X}}$.
\end{theorem}

\begin{proof}
  Locally compact, not metacompact.

  Game idea: Suppose there exists $\alpha\in\Lim(\omega_1)$ such that one may find arbitrarily large $n<\omega$ and $x\in U_\alpha$ such that $\sigma(x,\beta_{\alpha,n})$ does not forbid a branch for $\alpha$. Then $P$ has a counter converging to $\alpha$.

  Otherwise define $f:\Lim(\omega_1)\to\omega$ such that for $n\geq f(\alpha)$, $\sigma(x,\beta_{\alpha,n})$ forbids a branch for $\alpha$ for all $x\in U_\alpha$. Then $g(\alpha) = \beta_{\alpha,f(\alpha)}$ is a downward function, and there exists a stationary subset $S\subseteq \Lim_{\omega_1}$ and some $\beta\in\Suc(\omega_1)$ such that $\beta = \beta_{\alpha,f(\alpha)}$ for all $\alpha\in S$.

  Observing that $\{U_\alpha : \alpha \in S\}$ is dfggreat
\end{proof}

% \begin{thebibliography}{9}

% \bibitem{G1}
%   G. Gruenhage,
%   \emph{Covering Properties on $X^2\setminus\Delta$, $W$-sets, and Compact Subsets of $\Sigma$-Products}.
%   Topology Appl. 17 (1984), 287-304

% \bibitem{G2}
%   G. Gruenhage,
%   \emph{Games, Covering Properites, and Eberlein Compacts}.
%   Topology Appl. 23 (1986), 291-297

% \bibitem{Michael}
%   G. Gruenhage and E. Michael, 
%   \emph{A result on shrinkable open covers}.
%   Proceedings of the 1983 topology conference (Houston, Tex., 1983), 1983, pp. 37–43.

% \bibitem{Todercevic}
%   S. Todercevic,
%   \emph{Trees and linearly ordered sets}.
%   Handbook of Set-theoretic Topology (North-Holland, Amsterdam, 1984), pp. 235-293

% \bibitem{Watson}
%   S. Watson,
%   \emph{Locally compact normal meta-Lindel\"of spaces may not be paracompact: 
%   an application of uniformization and Suslin lines}. 
%   Proc. Amer. Math. Soc. 98 (1986), no. 4, 676–680.

% \end{thebibliography}

\end{document}