### 2013-05-08 sorgenfrey/sorgenfrey.tex
added simple result to sorgenfrey

\input{../header.tex}

\begin{document}

In this paper we investigate an open question posed to us by Gruenhage:

\begin{question}
Let $P$ be the subspace of the Sorgenfrey line containing only irrational numbers. Does there exist a base $\mc{B}$ for $P$ such that for every $\mc{C}\subseteq\mc{B}$ which is also a base for $X$, there exists a locally finite subcover $\mc{C}'\subseteq\mc{C}$?
\end{question}

We begin by tackling a simpler (solved) problem:

\begin{proposition}
Let $R$ be the Sorgenfrey line, the set of real numbers with the topology generated by the base $\mc{B} = \{[a,b): a<b\in R\}$ (where $[a,b) = \{x : a\leq x < b\}$).

For every $\mc{C}\subseteq\mc{B}$ which is also a base for $R$, there exists a pairwise disjoint subcover $\mc{C}'\subseteq\mc{C}$ (and thus a locally finite subcover).
\end{proposition}

\begin{proof}
We begin by letting $b_{n,0}=n$ for each $n<\omega$, and if $b_{n,\alpha}$ is defined for some ordinal $\alpha<\omega_1$ and $b_{n,\alpha}<n+1$, we define its successor $b_{n,\alpha+1}$ as follows:
  \begin{itemize}
    \item $b_{n,\alpha} < b_{n,\alpha+1} \leq n+1$
    \item $[b_{n,\alpha},b_{n,\alpha+1})\in \mathcal{C}$

    (This is possible as $\mc C$ is a base, and there must be some element of $\mc C$ which contains $b_{n,\alpha}$ and is a subset of $[b_{n,\alpha},n+1)$.)
  \end{itemize}
(If $b_{n,\alpha}=n+1$, then let $b_{n,\alpha+1}=n+1$ as well.) Finally, if $\alpha<\omega_1$ is a limit ordinal, let $\ds b_{n,\alpha} = \lim_{\beta\to\alpha}b_{n,\beta}$.

Let $C_{n,\alpha} = [b_{n,\alpha},b_{n,\alpha+1})$. We claim $\mc{C}' = \{C_{n,\alpha} : n<\omega,\alpha<\omega_1\}$ is a pairwise disjoint cover of $R$. Pairwise disjoint is evident by definition. To see that it is a cover, suppose it wasn't and missed some $x\in[n,n+1)$. Then we have an uncountable increasing sequence of numbers $\{b_{n,\alpha}:\alpha<\omega_1\}$, which contradicts the countable chain condition on the real line.
\end{proof}

\end{document}