### 2013-07-04 miscellaneous/inverse-limit-seminar-summer-2013/results.tex
Refactored file system

\input{../header.tex}

\begin{document}

\begin{example}
Let $L\subseteq \prod_{i=1}^\infty [0,1]$ be the inverse limit space with bonding functions all equal to
\[
  f(x) = \left\{
     \begin{array}{lr}
       2x & : x \leq 0.5 \\
       2-2x & : x \geq 0.5
     \end{array}
   \right.
\]
Then the following hold:
  \begin{enumerate}
    \item The subspaces $C_t = \{\alpha\in L : \alpha(1)=t\}$ are each homeomorphic to the Cantor Set.
    \item All proper subcontinuua are homeomorphic to the unit interval $[0,1]$
    \item All proper subcontinuua are nowhere dense in the space.
  \end{enumerate}
\end{example}

\begin{proof}
For (1), consider that for each number $t\in(0,1)$, there are exactly two preimages under $f$ inside $(0,1)$, giving a natural corespondance with the Cantor tree with branches given by points in the Cantor set $2^\omega$. The arguments for $t=0,1$ are similar.

For (2), we first claim that all basic open sets are of the form $L\cap\prod_{n=1}^\infty B_n$ where $B_n=[0,1]$ for all $n\not=N$. By definition all basic open sets are of the form $L\cap\prod_{n=1}^\infty A_n$ where $A_n=[0,1]$ for all $n\not=N_1,\dots,N_m$. It's easily seen that if $N=N_m$ and $B_N = \bigcap_{i=1}^m (f^{-1})^{(N_m-N_i)}(A_{N_i})$ and $B_n=[0,1]$ otherwise, then $L\cap\prod_{n=1}^\infty A_n = L\cap\prod_{n=1}^\infty B_n$.

Using this fact, we can argue that each proper subcontinuum $K$ is of the form $L \cap \prod_{n=1}^\infty [a_n,b_n]$ where $0\leq a_N < b_N < 1$ for some $N$.

First note that every maximal basic open set missed by $K$ must be of the form $L\cap\prod_{n=1}^\infty B_n$ here $B_n=[0,1]$ for all $n\not=N$ and $[0,1] \setminus B_N$ is connected, that is, $B_N=[0,a)\cup(b,1]$. (If not, then the disconnection of $[0,1]\setminus B_N$ yields a disconnection for $K$.) Thus $K = \prod_{n=1}^\infty [0,1]\setminus ([0,a_n)\cup(b_n,1]) = \prod_{n=1}^\infty [a_n,b_n]$.

We can see that $a_n<b_n$ always; otherwise the continuum is a single point. Suppose $b_n = 1$ for all $n$, then either $a_n=0$ for all $n$ (contradiction since $K$ is a proper subcontinuum), or $a_N>0$, and $b_{N+1}\leq 1-\frac{a_N}{2}$ (contradition).

Finally, considering $[a_N,b_N]$ where $0\leq a_N<b_n<1$, we note that as only a single sequence $\alpha$ in $K$ may satisfy $\alpha(N)=t$ for each $t\in [a_N,b_N]$, the projection from $K$ onto the $N^{th}$ coordinate is a homeomorphism onto $[a_N,b_N]$.

For (3), we need only observe that for any sequence $\alpha\in K$, an open neighborhood of $\alpha$ would contain infinte sequences $\beta$ such that $\alpha(N)=\beta(N)$.

\end{proof}

\end{document}