### 2013-07-04 miscellaneous/sorgenfrey/sorgenfrey.tex

\input{../header.tex}

\begin{document}

In this paper we investigate an open question posed to us by Gruenhage:

\begin{question}
Let $P$ be the subspace of the Sorgenfrey line containing only irrational numbers. Does there exist a base $\mc{B}$ for $P$ such that for every $\mc{C}\subseteq\mc{B}$ which is also a base for $X$, there exists a locally finite subcover $\mc{C}'\subseteq\mc{C}$?
\end{question}

We begin by tackling a simpler (solved) problem:

\begin{proposition}
Let $R$ be the Sorgenfrey line, the set of real numbers with the topology generated by the base $\mc{B} = \{[a,b): a<b\}$ (where $[a,b) = \{x : a\leq x < b\}$).

For every $\mc{C}\subseteq\mc{B}$ which is also a base for $R$, there exists a pairwise disjoint subcover $\mc{C}'\subseteq\mc{C}$ (and thus a locally finite subcover).
\end{proposition}

\begin{proof}
We begin by letting $b_{n,0}=n$ for each $n<\omega$, and if $b_{n,\alpha}$ is defined for some ordinal $\alpha<\omega_1$ and $b_{n,\alpha}<n+1$, we define its successor $b_{n,\alpha+1}$ as follows:
  \begin{itemize}
    \item $b_{n,\alpha} < b_{n,\alpha+1} \leq n+1$
    \item $[b_{n,\alpha},b_{n,\alpha+1})\in \mathcal{C}$

    (This is possible as $\mc C$ is a base, and there must be some element of $\mc C$ which contains $b_{n,\alpha}$ and is a subset of $[b_{n,\alpha},n+1)$.)
  \end{itemize}
(If $b_{n,\alpha}=n+1$, then let $b_{n,\alpha+1}=n+1$ as well.) Finally, if $\alpha<\omega_1$ is a limit ordinal, let $\ds b_{n,\alpha} = \lim_{\beta\to\alpha}b_{n,\beta}$.

Let $C_{n,\alpha} = [b_{n,\alpha},b_{n,\alpha+1})$. We claim $\mc{C}' = \{C_{n,\alpha} : n<\omega,\alpha<\omega_1\}$ is a pairwise disjoint cover of $R$. Pairwise disjoint is evident by definition. To see that it is a cover, suppose it wasn't and missed some $x\in[n,n+1)$. Then we have an uncountable increasing sequence of numbers $\{b_{n,\alpha}:\alpha<\omega_1\}$, which contradicts the countable chain condition on the real line.
\end{proof}

This idea can actually be generalized for any dense subspace, but it takes some machinery:

\begin{theorem}
Let $P$ be a dense subspace of the Sorgenfrey line with the induced topology, which has the base $\mc{B} = \{[a,b) : a<b, a\in P\}$.

For every $\mc{C}\subseteq\mc{B}$ which is also a base for $P$, there exists a pairwise disjoint subcover $\mc{C}'\subseteq\mc{C}$ (and thus a locally finite subcover).
\end{theorem}
\newpage
\begin{proof}
It suffices to show for $[0,1)\cap P$. We begin by constructing a collection of functions $S\subseteq \omega^{\omega_1}$ and numbers $c_s$, $d_s$, $c_{s\concat\<-1\>}$ defined by those functions as follows:
  \begin{itemize}
    \item Let $S_0 = \{\emptyset\}$. Let $d_{\emptyset}=0$ and $c_{\<-1\>}=1$.
    \item Suppose $S_\alpha$ has been defined, as well as $d_s\leq c_{s\concat\<-1\>}$ for each $s\in S_\alpha$. For $s\in S_\alpha$, consider the following:
      \begin{itemize}
        \item If $d_s = c_{s\concat\<-1\>}$, do nothing.
        \item If $d_s < c_{s\concat\<-1\>}$ and $d_s\in P$, let $S_{\alpha+1}$ contain $s\concat\<0\>$ and define $c_{s\concat\<0\>}$, $d_{s\concat\<0\>}$, $c_{s\concat\<0,-1\>}$ such that 
          \[
            d_s = c_{s\concat\<0\>} < d_{s\concat\<0\>} \leq c_{s\concat\<0,-1\>} = c_{s\concat\<-1\>}
          \]
        where $[c_{s\concat\<0\>}, d_{s\concat\<0\>})\in \mc{C}$.
        \item If $d_s < c_{s\concat\<-1\>}$ and $d_s\not\in P$, let $S_{\alpha+1}$ contain $s\concat\<n\>$ for all $n<\omega$ and define $c_{s\concat\<n\>}$, $d_{s\concat\<n\>}$, $c_{s\concat\<n,-1\>}$ for all $n<\omega$ such that
          \[
            d_s < \dots \leq c_{s\concat\<2,-1\>} = c_{s\concat\<1\>} < d_{s\concat\<1\>} \leq c_{s\concat\<1,-1\>} = c_{s\concat\<0\>} < d_{s\concat\<0\>} \leq c_{s\concat\<0,-1\>} = c_{s\concat\<-1\>}
          \]
        where $[c_{s\concat\<n\>}, d_{s\concat\<n\>})\in \mc{C}$ for all $n<\omega$ and $c_{s\concat\<n\>} \rightarrow d_s$.
      \end{itemize}
    \item Suppose $\alpha$ is a limit ordinal and $S_\beta$ has been defined for all $\beta<\alpha$. If $s\in\omega^\alpha$ and $t\in \bigcup_{\beta<\alpha}S_\beta$ for all $t<s$, let $S_\alpha$ contain $s$ and define $d_s = \lim_{t<s}d_t$ and $c_{s\concat\<-1\>} = \lim_{t<s} c_{t\concat\<-1\>}$.
      \item Let $S=\bigcup_{n<\omega} S_\alpha$.
  \end{itemize}

  Let $S = \bigcup_{\alpha<\omega_1} S_\alpha$. By construction, $\mc{C}' = \{[c_s,d_s) : s\in S\}$ is a disjoint subcollection of $\mc{C}$. We claim it also must cover $[0,1)\cap P$.

  Suppose not: $x\in[0,1)\cap P$ is not contained in $[c_s,d_s)$ for any $s\in S$. Note that $d_\emptyset<x$ (or else $x\in[c_{\<0\>},d_{\<1\>}$)). Assume $n_\beta<\omega$ is defined for all $\beta<\alpha$, and consider $s\in\omega^\alpha$ where $s(\beta)=n_\beta$. If $d_s<x$, we claim there is a minimal $n_\alpha<\omega$ where $d_{s\concat\<n_\alpha\>}<x$.
    \begin{itemize}
      \item This possible when $d_s\not\in P$ since $d_{s\concat\<n\>}\to d_s$.
      \item This is also possible when $d_s\in P$ since $[c_{s\concat\<0\>},d_{s\concat\<0\>})$ does not contain $x$, and thus $d_s=c_{s\concat\<0\>}<d_{s\concat\<0\>}\leq x$. If $d_{s\concat\<0\>} = x$ then $d_{s\concat\<0\>}\in P$ and $[d_{s\concat\<0\>},d_{s\concat\<0,0\>})=[c_{s\concat\<0,0\>},d_{s\concat\<0,0\>})$ contains $x$, which is a contradiction.
    \end{itemize}

  Finally, we notice that by defining $f\in\omega^\omega_1$ such that $f(\alpha)=n_\alpha$, then $d_{f\rest\alpha}$ is an increasing uncountable sequence defined for all $\alpha<\omega_1$, which is a contradiction.
\end{proof}

\end{document}