### 2013-09-24 todo.tex
Updated filling results and todos

\input{header.tex}

\begin{document}
\centerline{\bf Research TODOs}

  \begin{itemize}
    \item Menger-related properties and games results
      \begin{itemize}
        \item Is there a slick characterization of $F\kmarkwin{2}\mengame{X}$ for regular/general spaces?
        \item Is $F\uparrow_{?}\mengame{X}$ or $F\uparrow_{?}\rothgame{X}$ a hereditary property under closed subsets for any type of limited information? (The Menger property is; is Rothberger?)
      \end{itemize}
    \item Filling games
      \begin{itemize}
        \item Show/disprove $F\ktactwin{3}\sfillgame{J}$ implies $F\kmarkwin{3}\fillgame{J}$.
        \item Show/disprove $F\kmarkwin{2}\kfillgame{\kappa}$ implies $F\kmarkwin{2}\mengame{\oneptlind{\kappa}}$.
      \end{itemize}
    \item Search for a class of spaces where $K\ktactwin{2}\lfkpgame{X}$ characterizes metacompact (aka implies $K\tactwin\lfkpgame{X}$)
      \begin{itemize}
        \item Investigate the ladder space suggested by G.
        \item Try zero-dimensional.
      \end{itemize}
    \item Proximity Game
      \begin{itemize}
        \item Read paper by Bell
      \end{itemize}
  \end{itemize}
\end{document}