### 2014-04-28 phd/dissertation/appendix1.tex
Initial dissertation commit

%!TEX root = dissertation.tex
% ^ leave for LaTeXTools build functionality

\appendix
\chapter*{Appendices\addcontentsline{toc}{chapter}{Appendices}}

\chapter{Notes on the style-file project}

\begin{singlespace}
These style-files have been modified by Christopher Wilson to support the new Electronic Thesis and Dissertation process. The following was written before the ETD guidelines update on Fall 2009. Changes to the original appendix have been shown in italics.

These style-files for use with \LaTeX{} {\it were} maintained by Darrel
Hankerson\footnote{Mathematics and Statistics, 221 Parker Hall,
844-3641, {\tt hankedr@auburn.edu}} and Ed
Slaminka\footnote{Mathematics and Statistics, 218 Parker,
{\tt slamiee@auburn.edu}}.

In 1990, department heads and other representatives met with Dean Doorenbos
and Judy Bush-Crofton (then responsible for manuscript approval). This
meeting was prompted by a memorandum\footnote{Originally, the memorandum
was presented to Professor Larry Wit. A copy is available on request.} from
members of the mathematics departments concerning the {\em Thesis and
Dissertation Guide\/} and the approval process. There was wide agreement
among the participants (including Dean Doorenbos) to support the basic
recommendations outlined in the memorandum. The revised {\em Guide\/}
reflected some (but not all) of the agreements of the meeting.

Ms Bush-Crofton was supportive of the plan to obtain ``official approval''
of these style files.\footnote{Followup memoranda gave a definition of
``official approval.'' Copies will be sent on request.}  Unfortunately, Ms
Bush-Crofton left the Graduate School before the process was completed. In
1994, we were revisiting some of the same problems which were
resolved at the 1990 meeting.

In Summer 1994, I sent several memoranda to Ms Ilga Trend of the Graduate
School, reminding her of the agreements made at the 1990 meeting.
Professors A. Scottedward Hodel and Stan Reeves provided additional
support.  In short, it is essential that the Graduate School honor its
commitments of the 1990 meeting. It should be emphasized that Dean
Doorenbos is to thank for the success of that meeting.

Maintaining these \LaTeX{} files has been more work than expected, in
part due to continuing changes in requirement by the graduate school.
The Graduate School occasionally has complete memory loss about the
agreements of the 1990 meeting. If the Graduate School rejects your
manuscript based on items controlled by the style-files, ask
your advisor to contact the Graduate school (and copy to me) to urge
cooperation.

Finally, there have been several requests for additions to the package
(mostly formatting changes for figures, etc.). While such changes are not
really part of the thesis-style package, it could be beneficial to collect
these options and distribute with the package (making it easier on the next
student).  I'm especially interested in changes needed by various
departments.
\end{singlespace}