### 2014-04-28 phd/dissertation/chapter1.tex

%!TEX root = dissertation.tex
% ^ leave for LaTeXTools build functionality

\chapter{Introduction}

Basic overview of combinatorial games, topological games, limited info strategies,
and applications in topology.

% \section{heading}

% \subsection{other heading}