### 2014-05-14 phd/dissertation/chapter3.tex

%!TEX root = dissertation.tex
% ^ leave for LaTeXTools build functionality

\chapter{$W$ convergence and clustering games}

Results related to Gruenhage's ``W''-convergence game and variants.

(these are related to the ``hard'' version for $\pl O$ where $\pl P$
need only play within the most recent open set)

\section{Fort spaces}

\begin{thm}
$O\not\ktactwin{k}\clusgame{\oneptcomp{\omega_1}}{\infty}$.
\end{thm}

\begin{thm}
$O\markwin \clusgame{\oneptcomp{\omega_1}}{\infty}$.
\end{thm}

\begin{thm}(Nyikos)
$O\not\markwin \congame{\oneptcomp{\omega_1}}{\infty}$.
\end{thm}

\begin{thm}
$O\not\kmarkwin{k}\clusgame{\oneptcomp{\kappa}}{\infty}$ for $\kappa>\omega_1$.
\end{thm}

(TODO: It's feasible that $k$-limit $\Leftrightarrow$ $1$-limit.)

\section{Sigma-products}

\begin{thm}
  Let $\textrm{cf}([\kappa]^{\leq\omega})=\kappa$.
  Then $F \codewin PF_{F,C}(\kappa)$.
\end{thm}

\begin{thm}
  Let $\kappa$ be the limit of cardinals $\kappa_n$ such that
  $\textrm{cf}([\kappa_n]^{\leq\omega},\subseteq)=\kappa_n$.
  Then $F \codewin \pfgame{\kappa}$.
\end{thm}

\begin{thm}
  $F\codewin\pfgame{\kappa}$ for all cardinals $\kappa$.
\end{thm}

\begin{cor}
  $O\codewin\congame{\sigmaprodr{\kappa}}{\vec{0}}$
  for all cardinals $\kappa$.
\end{cor}