### 2014-08-25 phd/dissertation/_raw_chapter4.tex

%!TEX root = dissertation.tex
% ^ leave for LaTeXTools build functionality

\chapter{Proximal Game}

Results pertaining to Bell's proximal game for uniform spaces.

\section{cut-and-paste}

\begin{defn}
  A \textbf{uniform space} $\<X,\mc D\>$ is a set $X$ paired with a filter $\mc D$ (called its \textbf{uniformity}) of relations (called \textbf{entourages}) on $X$ such that for each entourage $D\in \mc D$:
    \begin{itemize}
      \item $D$ is reflexive, i.e., the diagonal $\Delta\subseteq D$.
      \item Its inverse $D^{-1}=\{\<y,x\>:\<x,y\>\in D\}\in \mc D$.
      \item There exists $\frac{1}{2}D\in\mc D$ such that
        \[
          2(\frac{1}{2}D)=\frac{1}{2}D\circ\frac{1}{2}D=\{\<x,z\>:\exists y(\<x,y\>,\<y,z\>\in\frac{1}{2}D)\}\subseteq D
        \]
    \end{itemize}
  Note that since $\mc D$ is a filter, for each $D\in\mc D$, the symmetric relation $D\cap D^{-1}\in \mc D$.
\end{defn}

\begin{prop}
For each $D\in\mc D$ and $n<\omega$ there exists $\frac{1}{2^{n+1}}D\in\mc D$ such that
  \[2(\frac{1}{2^{n+1}}D)=\frac{1}{2^{n+1}}D\circ \frac{1}{2^{n+1}}D\subseteq \frac{1}{2^{n}}D\]
and if $2E\subseteq \frac{1}{2^{n}}D$, then $E\subseteq \frac{1}{2^{n+1}}D$.
\end{prop}

\begin{defn}
  For an entourage $D\in\mc D$, let $D[x]=\{y:(x,y)\in D\}$ be the \textbf{$D$-neighborhood} of $x$. The \text{uniform topology} for a uniform space $\<X,\mc D\>$ is generated by the base $\{D[x]:x\in X,D\in\mc D\}$.
\end{defn}

\begin{thm}
  A space $X$ is uniformizable (its topology is the uniform topology for some uniformity) if and only if $X$ is completely regular ($T_{3\frac{1}{2}}$).
\end{thm}

\begin{prop}
  If $X$ is a uniform space, then for all $x\in X$ and symmetric entourages $D$:
    \[
      x\in \frac{1}{2}D[y]\text{ and } y\in\frac{1}{2}D[z] \Rightarrow x\in D[z]
    \]
  and
    \[
      \frac{1}{2}D[x]\subseteq \cl{\frac{1}{2}D[x]}\subseteq D[x]
    \]
\end{prop}

\begin{proof}
  The first is by definition of $\frac{1}{2}D$.

  If $z\in \cl{\frac{1}{2}D[x]}$, it follows that there is $y\in \frac{1}{2}D[x]\cap\frac{1}{2}D[z]$ since $\frac{1}{2}D[z]$ is an open neighborhood of $z$. Thus $(x,z)\in D \Rightarrow z\in D[x] \Rightarrow \cl{\frac{1}{2}D[x]}\subseteq D[x]$.
\end{proof}

\begin{defn}
  For a uniform space $X$, Bell's proximity game proceeds as follows.

  In round $0$, $\pl D$ chooses an entourage $D_0$, followed by $\pl P$ choosing a point $p_0\in X$.

  In round $n+1$, $\pl D$ chooses an entourage $D_{n+1}\subseteq D_n$, followed by $\pl P$ choosing a point $p_{n+1}\in 4D_n[p_n]$.

  Player $\pl D$ wins if either $\bigcap_{n<\omega} 4D_n[p_n] = \emptyset$ or $\<p_0,p_1,\dots\>$ converges.
\end{defn}

\begin{defn}
  For a uniform space $X$, the simplified proximal game $\bellConGame{X}$ can be defined as follows:

  In round $0$, $\pl D$ chooses a symmetric entourage $D_0$, followed by $\pl P$ choosing a point $p_0\in X$.

  In round $n+1$, $\pl D$ chooses a symmetric entourage $D_{n+1}$, followed by $\pl P$ choosing a point $p_{n+1}\in\left(\bigcap_{m\leq n}D_m\right)[p_n]$.

  Player $\pl D$ wins if either $\bigcap_{n<\omega}\left(\bigcap_{m\leq n} D_m\right)[p_n]=\emptyset$ or $\<p_0,p_1,\dots\>$ converges.
\end{defn}

\begin{thm}
  $\pl D$ has a winning perfect-information strategy in Bell's game if and only if $\pl D\win\bellConGame{X}$.
\end{thm}

\begin{proof}
  Let $\sigma$ be a winning perfect information strategy for $\pl D$ in Bell's game. We define a perfect information strategy $\tau$ in the simplified game to yield symmetric entourages $\tau(p\rest n)=\sigma(p\rest n)\cap(\sigma(p\rest n))^{-1}$ for all partial attacks $p\rest n$. Note that $\tau(p\rest n)=\bigcap_{m\leq n}\tau(p\rest m)$.

  If $p$ attacks $\tau$ in the simplified game, $p(n+1)\in\left(\bigcap_{m\leq n}\tau(p\rest m)\right)[p(n)]=\tau(p\rest n)[p(n)]\subseteq\sigma(p\rest n)[p(n)]\subseteq 4\sigma(p\rest n)[p(n)]$, so $p$ attacks $\sigma$ in Bell's game. Thus either $p$ converges, or
    \[
      \emptyset
        =
      \bigcap_{n<\omega} 4\sigma(p\rest n)[p(n)]
        \supseteq
      \bigcap_{n<\omega}\tau(p\rest n)[p(n)]
        =
      \bigcap_{n<\omega}\left(\bigcap_{m\leq n}\tau(p\rest n)\right)[p(n)]
    \]

  For the other direction, let $\sigma$ be a winning perfect information strategy for $\pl D$ in the simplified game such that $\sigma(p\rest n)=\bigcap_{m\leq n}\sigma(p\rest m)$. Define the perfect information strategy $\tau$ in Bell's Game such that $4\tau(p\rest n)\subseteq \sigma(p\rest n)$ and $\tau(p\rest n)=\bigcap_{m\leq n}\tau(p\rest m)$ for all partial attacks $p\rest n$.

  If $p$ attacks $\tau$ in Bell's game, $p(n)\in 4\tau(p\rest n)\subseteq\sigma(p\rest n)=\bigcap_{m\leq n}\sigma(p\rest m)$, so $p$ attacks $\sigma$ in the simplified game. Thus either $p$ converges, or
    \[
      \emptyset
        =
      \bigcap_{n<\omega}\left(\bigcap_{m\leq n}\sigma(p\rest n)\right)[p(n)]
        =
      \bigcap_{n<\omega}\sigma(p\rest n)[p(n)]
        \supseteq
      \bigcap_{n<\omega}4\tau(p\rest n)[p(n)]
        \supseteq
      \bigcap_{n<\omega}\tau(p\rest n)[p(n)]
    \]
\end{proof}

\begin{prop}
$\pl P$ has a winning perfect-information strategy in Bell's game if and only if $\pl P \win \bellConGame{X}$.
\end{prop}

\begin{proof}
  Similar to the previous.
\end{proof}

\begin{defn}
  A uniform space is \textbf{proximal} if $\pl D \win \bellConGame{X}$.
\end{defn}

\begin{defn}
  For a space $X$ and a point $x\in X$, the \textbf{$W$-convergence-game} $\gruConGame{X}{x}$ proceeds as follows.

  In round $0$, $\pl O$ chooses a neighborhood $U_n$ of $x$, followed by $\pl P$ choosing a point $p_n\in \bigcap_{m\leq n}U_m$.

  Player $\pl O$ wins if $\<p_0,p_1,\dots\>$ converges.
\end{defn}

\begin{defn}
  A space is \textbf{$W$} if $\pl O \win \gruConGame{X}{x}$ for all $x\in X$.
\end{defn}

\begin{defn}
  For each finite tuple $(m_0,\dots,m_{n-1})$, we define the \textbf{$k$-tactical fog-of-war}
    \[
      T_k(\<m_0,\dots,m_{n-1}\>)=\<m_{n-k},\dots,m_{n-1}\>
    \]
  and the \textbf{$k$-Mark\"ov fog-of-war}
    \[
      M_k(\<m_0,\dots,m_{n-1}\>)=\<\<m_{n-k},\dots,m_{n-1}\>,n\>
    \]

  So $P\ktactwin{k}G$ if and only if there exists a winning strategy for $P$ of the form $\sigma\circ T_k$, and $P\kmarkwin{k}G$ if and only if there exists a winning strategy of the form $\sigma\circ M_k$.
\end{defn}

\begin{thm}
For all $x\in X$:
  \begin{itemize}
    \item
      $\pl D\win \bellConGame{X} \Rightarrow \pl O \win \gruConGame{X}{x}$
    \item
      $\pl D\ktactwin{2k} \bellConGame{X} \Rightarrow \pl O \ktactwin{k} \gruConGame{X}{x}$
    \item
      $\pl D\kmarkwin{2k} \bellConGame{X} \Rightarrow \pl O \kmarkwin{k} \gruConGame{X}{x}$
  \end{itemize}
\end{thm}

\begin{proof}
Let $\sigma$ witness $\pl D \ktactwin{2k}\bellConGame{X}$ (resp. $\pl D \kmarkwin{2k}\bellConGame{X}$, $\pl D\win\bellConGame{X}$). We define the $k$-tactical (resp. $k$-Mark\"ov, perfect info) strategy $\tau$ such that
  \[
    \tau\circ L_k(p)
      =
    \sigma\circ L_{2k}(\<x,p(0),\dots,x,p(|p|-1)\>)[x]
      \cap
    \sigma\circ L_{2k}(\<x,p(0),\dots,x,p(|p|-1),x\>)[x]
  \]
where $L_{2k}$ is the $2k$-tactical fog-of-war (resp. $2k$-Mark\"ov fog-of-war, identity) and $L_{k}$ is the $k$-tactical fog-of-war (resp. $k$-Mark\"ov fog-of-war, identity).

Let $p$ attack $\tau$. Consider the attack $q$ against the winning strategy $\sigma$ such that $q(2n)=x$ and $q(2n+1)=p(n)$, and let $D_n=\sigma\circ L_{2k}(q)$ and $E_n=\bigcap_{m\leq n}D_n$.

Certainly, $x\in E_{2n}[x]= E_{2n}[q(2n)]$ for any $n<\omega$. Note also for any $n<\omega$ that
    \[
      p(n) \in
      \bigcap_{m\leq n}\tau\circ L_k(p\rest n)
    \]
    \[
      =
      \bigcap_{m\leq n}\left(
        \sigma\circ L_{2k}(\<x,p(0),\dots,x,p(m-1)\>)[x]\cap
        \sigma\circ L_{2k}(\<x,p(0),\dots,x,p(m-1),x\>)[x]
      \right)
    \]
    \[
      =
      \bigcap_{m\leq n}\left(
        D_{2m}[x]\cap
        D_{2m+1}[x]
      \right) =
      \bigcap_{m\leq 2n+1} D_m[x]=E_{2n+1}[x]
    \]
so by the symmetry of $E_{2n+1}$, $x\in E_{2n+1}[p(n)]= E_{2n+1}[q(2n+1)]$. Thus $x\in \bigcap_{n<\omega} E_n[q(n)]\not=\emptyset$, and since $\sigma$ is a winning strategy, the attack $q$ converges. Since $q(2n)=x$, $q$ must converge to $x$. Thus its subsequence $p$ converges to $x$, and $\tau$ is a winning strategy in $\gruConGame{X}{x}$.
\end{proof}

\begin{cor}
For all $x\in X$:
  \begin{itemize}
    \item
      $\pl D\ktactwin{k} \bellConGame{X} \Rightarrow \pl O \ktactwin{k} \gruConGame{X}{x}$
    \item
      $\pl D\kmarkwin{k} \bellConGame{X} \Rightarrow \pl O \kmarkwin{k} \gruConGame{X}{x}$
  \end{itemize}
\end{cor}

\begin{cor}
  All proximal spaces are $W$-spaces.
\end{cor}

\begin{thm}
  Let $X\cup\{\infty\}$ be a uniformizable space such that $X$ is discrete. Then
  \begin{itemize}
    \item
      $\pl O\win \gruConGame{X\cup\{\infty\}}{\infty} \Rightarrow \pl D \win \bellConGame{X\cup\{\infty\}}$
    \item
      $\pl O\ktactwin{k} \gruConGame{X\cup\{\infty\}}{\infty} \Rightarrow \pl D \ktactwin{k} \bellConGame{X\cup\{\infty\}}$
    \item
      $\pl O\kmarkwin{k} \gruConGame{X\cup\{\infty\}}{\infty} \Rightarrow \pl D \kmarkwin{k} \bellConGame{X\cup\{\infty\}}$
  \end{itemize}
\end{thm}

\begin{proof}
  Note that the topology on $X\cup\{\infty\}$ is induced by the uniformity with equivalence relation entourages $D(U)=\Delta\cup U^2$ for each open neighborhood $U$ of $\infty$.

  Let $\sigma$ witness $\pl D \ktactwin{k}\gruConGame{X\cap\{\infty\}}{\infty}$ (resp. $\pl D \kmarkwin{k}\gruConGame{X\cap\{\infty\}}{\infty}$, $\pl D\win\gruConGame{X\cap\{\infty\}}{\infty}$). We define the $k$-tactical (resp. $k$-Mark\"ov, perfect info) strategy $\tau$ such that
    \[
      \tau\circ L(p)
        =
      D(\sigma\circ L(p))
    \]
  where $L$ is the $k$-tactical fog-of-war (resp. $k$-Mark\"ov fog-of-war, identity).

  Let $p\in (X\cup\{\infty\})^\omega$ attack $\tau$ such that $\bigcap_{n<\omega}\tau(p\rest n)[p(n)]\not=\emptyset$.

  If $\infty\in\bigcap_{n<\omega}\tau(p\rest n)[p(n)]$, it follows that $p$ is an attack on $\sigma$. Since $\sigma$ is a winning strategy, it follows that $q$ and its subsequence $p$ must coverge to $\infty$.

  Otherwise, $\infty\not\in\tau(p\rest N)[p(N)]$ for some $N<\omega$, and then $\tau(p\rest N)[p(N)]=\{p(N)\}$ implies $p\to p(N)$.

  Thus $\tau\circ L$ is a winning strategy.
\end{proof}

\begin{cor}
  Let $X\cup\{\infty\}$ be a uniformizable space such that $X$ is discrete. Then
  \begin{itemize}
    \item
      $\pl O\win \gruConGame{X\cup\{\infty\}}{\infty} \Leftrightarrow \pl D \win \bellConGame{X\cup\{\infty\}}$
    \item
      $\pl O\ktactwin{k} \gruConGame{X\cup\{\infty\}}{\infty} \Leftrightarrow \pl D \ktactwin{k} \bellConGame{X\cup\{\infty\}}$
    \item
      $\pl O\kmarkwin{k} \gruConGame{X\cup\{\infty\}}{\infty} \Leftrightarrow \pl D \kmarkwin{k} \bellConGame{X\cup\{\infty\}}$
  \end{itemize}
\end{cor}

\begin{prop} For any $x\in X$ and $k\geq 1$,
  \begin{itemize}
    \item
      $\pl O\ktactwin{k}\gruConGame{X}{x} \Leftrightarrow \pl O\tactwin\gruConGame{X}{x}$
    \item
      $\pl O\kmarkwin{k}\gruConGame{X}{x} \Leftrightarrow \pl O\markwin\gruConGame{X}{x}$
  \end{itemize}
\end{prop}

\begin{proof}
  If $\sigma$ witnesses $\pl O\ktactwin{k}\gruConGame{X}{x}$, let $\tau(\emptyset)=\sigma(\emptyset)$ and
    \[
      \tau(\<q\>)
        =
      \bigcap_{i< k}
      \sigma(\<\underbrace{x,\dots,x}_{k-i-1},q,\underbrace{x,\dots,x}_{i}\>)
    \]

  This is easily verified to be a winning strategy. The proof for $\pl O\kmarkwin{k}\gruConGame{X}{x}$ is analogous.
\end{proof}

\begin{cor}
  Let $X\cup\{\infty\}$ be a uniformizable space such that $X$ is discrete, and $k\geq 1$. Then
  \begin{itemize}
    \item
      $\pl D\ktactwin{k}\bellConGame{X\cup\{\infty\}} \Leftrightarrow O\tactwin\bellConGame{X\cup\{\infty\}}$
    \item
      $\pl D\kmarkwin{k}\bellConGame{X\cup\{\infty\}} \Leftrightarrow O\markwin\bellConGame{X\cup\{\infty\}}$
  \end{itemize}
\end{cor}

\begin{prop} For any uniform space $X$,
  \begin{itemize}
    \item
      $\pl O\ktactwin{k}\bellConGame{X} \Leftrightarrow \pl O\ktactwin{2}\bellConGame{X}$
    \item
      $\pl O\kmarkwin{k}\bellConGame{X} \Leftrightarrow \pl O\kmarkwin{2}\bellConGame{X}$
  \end{itemize}
\end{prop}

\begin{proof}
  If $\sigma$ witnesses $\pl O\ktactwin{k}\bellConGame{X}{x}$, let $\tau(\emptyset)=\sigma(\emptyset)$ and
    \[
      \tau(\<q\>)
        =
      \bigcap_{i<k}
      \sigma(\<\underbrace{q,\dots,q}_{i}\>)
    \]
    \[
      \tau(\<q,q'\>)
        =
      \bigcap_{i< k}
      \sigma(\<\underbrace{q,\dots,q}_{k-i},\underbrace{q',\dots,q'}_{i}\>)
    \]

  This is easily verified to be a winning strategy. The proof for $\pl O\kmarkwin{k}\bellConGame{X}{x}$ is analogous.
\end{proof}

\newpage

\begin{defn}
  The absolute proximal game $\bellClusGame{X}$ is analogous to $\bellConGame{X}$, except $\pl D$ may only win if $p$ converges.
\end{defn}

\begin{defn}
  A \textbf{uniformly locally compact} space is a uniformizable space with a \textbf{uniformly compact entourage} $M$ where $\cl{M[x]}$ is compact for all $x$.
\end{defn}

\begin{thm}
  For any uniformly locally compact space $X$,
      $\pl D\win \bellConGame{X} \Leftrightarrow \pl D\win \bellClusGame{X}$
\end{thm}

\begin{proof}
  Let $M$ be a uniformly locally compact entourage. Let $\sigma$ witness $\pl D\win \bellConGame{X}$ such that $\sigma(a)\subseteq M$ always (so $\cl{\sigma(a)[x]}\subseteq\cl{M[x]}$ is compact), and $a\supseteq b$  implies $\sigma(a)\subseteq\frac{1}{4}\sigma(b)$.

  Let $\tau(p\rest n)=\frac{1}{2}\sigma(p\rest n)$. If $p$ attacks $\tau$ in $\bellClusGame{X}$, then
    \[
      p(n+1)
        \in
      \tau(p\rest n)[p(n)]
        =
      \frac{1}{2}\sigma(p\rest n)[p(n)]
    \]

    and for

    \[
      x
        \in
      \cl{\sigma(p\rest (n+1))[p(n+1)]}
        \subseteq
      \cl{\frac{1}{4}\sigma(p\rest n)[p(n+1)]}
        \subseteq
      \frac{1}{2}\sigma(p\rest n)[p(n+1)]
    \]

  we can conclude $x\in\sigma(p\rest n)[p(n)]$. Thus

    \[
      \sigma(p\rest (n+1))[p(n+1)]
        \subseteq
      \cl{\sigma(p\rest (n+1))[p(n+1)]}
        \subseteq
      \sigma(p\rest n)[p(n)]
    \]

  Finally, note that $p$ attacks the winning strategy $\sigma$ in $\bellConGame{X}$, but since the intersection of a chain of nonempty compact sets is nonempty:

    \[
      \bigcap_{n<\omega} \sigma(p\rest n)[p(n)]
        =
      \bigcap_{n<\omega} \cl{\sigma(p\rest n)[p(n)]}
        \not=
      \emptyset
    \]

  We conclude that $p$ converges.
\end{proof}

\begin{cor}
  A uniformaly locally compact space $X$ is proximal if and only if $\pl D \win \bellClusGame{X}$.
\end{cor}

\begin{thm}
  For any uniformly locally compact proximal space $X$, $\pl O\win \gruClusGame{X}{H}$ for all compact $H\subseteq X$.
\end{thm}

\begin{proof}
  Let $\sigma$ witness $\pl D \win \bellClusGame{X}$ such that $p\supseteq q$ implies $\sigma(p)\subseteq \frac{1}{4}\sigma(q)$.

  Let $o(t)$ be the subsequence of $t$ consisting of its odd-indexed terms.

  We define $T(\emptyset)$, etc. as follows:

  \begin{itemize}
    \item Let $\emptyset\in T(\emptyset)$.
    \item Choose $m_\emptyset<\omega$, $h_{\emptyset,i}\in H$ for $i<m_\emptyset$, and $h_{\emptyset,i,j}\in H\cap\cl{\frac{1}{4}\sigma(\emptyset)[h_{\emptyset,i}]}$ for $i,j<m_\emptyset$ such that
      \[
        \{\frac{1}{4}\sigma(\emptyset)[h_{\emptyset,i}]:i<m_\emptyset\}
      \]
    is a cover for $H$ and such that for each $i<m_\emptyset$
      \[
        \{\frac{1}{4}\sigma(\<h_{\emptyset,i}\>)[h_{\emptyset,i,j}]:j<m_\emptyset\}
      \]
    is a cover for $H\cap\cl{\frac{1}{4}\sigma(\emptyset)[h_{\emptyset,i}]}$.
    \item Let $\<i\>\in T(\emptyset)$, $\<i,h_{\emptyset,i}\>\in T(\emptyset)$, and $\<i,h_{\emptyset,i},j\>\in T(\emptyset)$ for $i,j<m_\emptyset$.
  \end{itemize}

  Suppose $T(a)$, etc. are defined. We then define $T(a\concat\<x\>)$, etc. for
    \[
      x\in \bigcup_{s\concat\<i,h_{s,i},j\>\in\max(T(a))} \frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]
    \]
  as follows:

  \begin{itemize}
    \item Let $T(a)\subseteq T(a\concat\<x\>)$.
    \item Choose $t=s\concat\<i,h_{s,i},j,x\>$ such that $s\concat\<i,h_{s,i},j\>\in\max(T(a))$ and $x\in \frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]$.
    \item Note that, assuming $o(s)\concat\<h_{s,i}\>$ is a legal partial attack against $\sigma$, then
      \[
        x
          \in
        \frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]
          \subseteq
        \frac{1}{4}\sigma(o(s))[h_{s,i,j}]
      \]
    and
      \[
        h_{s,i,j}
          \in
        \cl{\frac{1}{4}\sigma(o(s))[h_{s,i}]}
          \subseteq
        \frac{1}{2}\sigma(o(s))[h_{s,i}]
      \]
    implies
      \[
        x
          \in
        \sigma(o(s))[h_{s,i}]
      \]
    and thus $o(s)\concat\<h_{s,i},x\>=o(t)$ is a legal partial attack against $\sigma$.
    \item Choose $m_t<\omega$, $h_{t,k}\in H\cap \cl{\frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]}$ for $k<m_t$, and $h_{t,k,l}\in H\cap\cl{\frac{1}{4}\sigma(t)[h_{t,k}]}$ for $k,l<m_t$ such that
      \[
        \{\frac{1}{4}\sigma(o(t))[h_{t,k}]:k<m_t\}
      \]
    is a cover for $H\cap \cl{\frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]}$ and such that for each $k<m_t$
      \[
        \{\frac{1}{4}\sigma(o(t)\concat\<h_{t,k}\>)[h_{t,i,j}]:l<m_t\}
      \]
    is a cover for $H\cap\cl{\frac{1}{4}\sigma(o(t))[h_{t,k}]}$.
    \item Note that, assuming $o(t)$ is a legal partial attack against $\sigma$, then
      \[
        h_{t,k}
          \in
        \cl{\frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]}
          \subseteq
        \frac{1}{2}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]
      \]
    and
      \[
        x
          \in
        \frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]
      \]
    implies
      \[
        h_{t,k}
          \in
        \sigma(o(s)\concat\<h_{s,i}\>)[x]
      \]
    and thus $o(t)\concat\<h_{t,k}\>$ is a legal partial attack against $\sigma$.
    \item Let $t\in T(a\concat\<x\>)$, $t\concat\<k\>\in T(a\concat\<x\>)$, $t\concat\<k,h_{t,k}\>\in T(a\concat\<x\>)$, and $t\concat\<k,h_{t,k},l\>\in T(a\concat\<x\>)$ for $k,l<m_t$.
    \item Note that assuming
      \[
        \{\frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}] : s\concat\<i,h_{s,i},j\>\in\max(T(a))\}
      \]
    covers $H$, then since
      \[
        \{\frac{1}{4}\sigma(o(t)\concat\<h_{t,k}\>)[h_{t,k,l}] : s\concat\<i,h_{s,i},j,x,k,h_{t,k},l\>\in\max(T(a\concat\<x\>))\setminus\max(T(a))\}
      \]
    covers $H\cap \frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]$, we have that
      \[
        \{\frac{1}{4}\sigma(o(t)\concat\<h_{t,k}\>)[h_{t,k,l}] : t\concat\<k,h_{t,k},l\>\in\max(T(a\concat\<x\>))\}
      \]
    covers $H$.
  \end{itemize}

  With this we may define the perfect information strategy $\tau$ for $\pl O$ in $\gruConGame{X}{H}$ such that:
  \[
    \tau(p\rest n) = \bigcup_{s\concat\<i,h_{s,i},j\>\in\max(T(p\rest n))} \frac{1}{4}\sigma(o(s)\concat\<h_{s,i}\>)[h_{s,i,j}]
  \]

  If $p$ attacks $\tau$, then it follows that $T(p\rest n)$ is defined for all $n<\omega$, so let $T(p)=\bigcup_{n<\omega} T(p\rest n)$. We note $T(p)$ is an infinite tree with finite levels:
    \begin{itemize}
      \item $\emptyset$ has exactly $m_\emptyset$ successors $\<i\>$.
      \item $s\concat\<i\>$ has exactly one successor $t\concat\<i,h_{s,i}\>$
      \item $s\concat\<i,h_{s,i}\>$ has exactly $m_s$ successors $t\concat\<i,h_{s,i},j\>$
      \item $s\concat\<i,h_{s,i},j\>$ has either no successors or exactly one successor $t\concat\<i,h_{s,i},j,x\>$
      \item $t=s\concat\<i,h_{s,i},j,x\>$ has exactly $m_t$ successors $t\concat\<k\>$
    \end{itemize}

  Let $q'=\<i_0,h_0,j_0,x_0,i_1,h_1,j_1,x_1,\dots\>$ correspond to this infinite branch in $T(p)$, and let $q=o(q')=\<h_0,x_0,h_1,x_1,\dots\>$. Note that by the construction of $T(p)$, $q$ is an attack on the winning strategy $\sigma$ in $\bellClusGame{X}$, so it must converge. Since every other term of $q$ is in $H$, it must converge to $H$. Then since $q$ is a subsequence of $p$, $p$ must cluster at $H$.
\end{proof}

\begin{cor}
  For any uniformly locally compact proximal space, $\pl O\win \gruConGame{X}{H}$ for all compact $H\subseteq X$.
\end{cor}

\begin{proof}
  $\pl O \win \gruConGame{X}{H}$ if and only if $\pl O \win \gruClusGame{X}{H}$.
\end{proof}

\begin{cor}
  A compact uniform space $X$ is Corson compact if and only if it is proximal.
\end{cor}

\begin{proof}
  A characterization of Corson compact is having a $W$-set diagonal. If $X$ is proximal compact, then $X^2$ is proximal compact, and its compact diagonal is a $W$-set.
\end{proof}

\newpage

\begin{thm}
  $\pl O \prewin \gruConGame{X}{H}$ if and only if there exists a countable base
  around $H$.
\end{thm}

\begin{proof}
  Let $\{U_n:n<\omega\}$ be a countable base around $H$. We define the
  predetermined strategy $\sigma(n)=\bigcap_{m\leq n}U_m$. Let $p$ attack
  $\sigma(n)$ - then if $U$ is any neighborhood of $H$, we may choose
  $H\subseteq U_m\subseteq U$, and note that $\sigma(n)\subseteq U_m$ for
  $n\geq m$, and thus $p(n)\in U_m\subseteq U$ for all $n\geq m$. Thus
  $\sigma$ is a winning strategy.

  For the other direction, suppose there does not exist a countable base
  around $H$, and let $\sigma(n)$ be an arbitrary predetermined strategy.
  Since $\{\bigcap_{m\leq n}\sigma(m):n<\omega\}$ is not a countable base
  around $H$, we may choose an open set $U$ around $H$ such that
  $\bigcap_{m\leq n}\sigma(m)\not\subseteq U$ for all $n<\omega$. We may
  easily verify that if $p(n)\in\bigcap_{m\leq n}\sigma(m)\setminus U$ for
  all $n<\omega$, then $p$ is a successful counterattack to $\sigma$.
\end{proof}

\begin{cor}
  $X$ is first countable if and only if $\pl O \prewin \gruConGame{X}{x}$ for
  all $x\in X$
\end{cor}

\begin{cor}
  $\pl D \prewin \bellConGame{X}$ implies $X$ is first countable.
\end{cor}

\begin{defn}
  Scattered Eberlein compact spaces are known as \term{strong Eberlein
  compact} spaces.
\end{defn}

\begin{thm}[folklore]
  Scattered compact first-countable spaces are metrizable.
\end{thm}

\begin{cor}
  If $X$ is scattered compact and $\pl O \prewin\gruConGame{X}{x}$ for all
  $x\in X$ (or $\pl D \prewin\bellConGame{X}$), then $X$ is metrizable.
\end{cor}

\begin{ex}
  $\pl D \not\prewin \bellConGame{\oneptcomp\omega_1}$
\end{ex}

\begin{proof}
  There does not exist a countable base around $\infty$, so
    $\pl O \not\prewin \gruConGame{X}{\omega_1}$.
\end{proof}

\begin{ex}
  $\pl O \tactwin \gruConGame{\oneptcomp\kappa}{\infty}$ and
  $\pl D \tactwin \bellConGame{\oneptcomp\kappa}$ for all cardinals $\kappa$
\end{ex}

\begin{proof}
  For $\gruConGame{\oneptcomp\kappa}{\infty}$, let
    $\sigma()=\sigma(\infty)=\oneptcomp\kappa$
  and
    $\sigma(x)=\oneptcomp\kappa\setminus\{x\}$
  otherwise.
\end{proof}

\newpage

\begin{thm}
  If $H$ is a closed subset of $X$, then
    $
      \pl D \limitwin \bellConGame{X}
        \Rightarrow
      \pl D \limitwin \bellConGame{H}
    $
  where $\limitwin$ is any of $\win$, $\ktactwin{k}$, or $\kmarkwin{k}$.
\end{thm}

\begin{proof}
  Let $\sigma\circ L$ witness $\pl D \limitwin \bellConGame{X}$. We define
  $\tau\circ L$ for $\pl D$ in $\bellConGame{H}$ as follows:
    \[
      \tau\circ L(p\rest n)
        =
      \sigma\circ L(p\rest n)
        \cap
      H^2
    \]

  Let $p$ attack $\tau\circ L$. $p$ also attacks the winning strategy
  $\sigma\circ L$, so either
    \[
      \bigcap_{n<\omega}
      \left(\bigcap_{m\leq n}\tau\circ L(p\rest n)\right)
      [p(n)]
        \subseteq
      \bigcap_{n<\omega}
      \left(\bigcap_{m\leq n}\sigma\circ L(p\rest n)\right)
      [p(n)]
        =
      \emptyset
    \]
  or $p$ converges in $X$, and thus converges in $H$.
\end{proof}

\begin{thm}
  If $\pl D\limitwin \bellConGame{X_i}$ for $i<\omega$, then
  $\pl D\limitwin \bellConGame{\prod_{i<\omega}X_i}$, where $\limitwin$ is either
  $\win$ or $\kmarkwin{k}$.
\end{thm}

\begin{proof}
  A subbase for $\prod_{i<\omega}X_i$ is
    \[
      \{\pi_i^{-1}(D):i<\omega, D\in\mc D_i\}
    \]
  where $\pi_i$ is the natural projection from
  $\left(\prod_{i<\omega}X_i\right)^2$ onto $X_i^2$. (See Bell.)

  For $p\in \left(\prod_{i<\omega}X_i\right)^\omega$, let $p_i\in X_i^\omega$
  such that $p_i(n)=p(n)(i)$.

  Let $\sigma_i\circ L$ witness $\pl D\limitwin \bellConGame{X_i}$ for
  $i<\omega$, and assume without loss of generality that
  $\sigma_i\circ L$ always yields $X_i^2$ before round $i$.

  Then we define the strategy $\tau\circ L$ for $\pl D$ in
  $\bellConGame{\prod_{i<\omega}X_i}$ as follows:
    \[
      \tau\circ L(p\rest n)
        =
      \bigcap_{i\leq n} \pi_i^{-1}(\sigma_i\circ L(p_i\rest n))
    \]

  Let $p$ attack $\tau\circ L$. If
    $
      \bigcap_{n<\omega}
      \left(\bigcap_{m\leq n}\sigma_i(p_i\rest n)\right)[p_i(n)]
      = \emptyset
    $
  for any $i<\omega$, it easily follows that
    $
      \bigcap_{n<\omega}
      \left(\bigcap_{m\leq n}\tau(p\rest n)\right)[p(n)]
      = \emptyset
    $.

  Otherwise, we assume that for each $i<\omega$, $p_i$ converges to some
  $x_i\in X_i$. Thus $p$ converges to $x=\<x_0,x_1,\dots\>$.
\end{proof}

Note: I expect I should be able to do some clever things
assuming $S(\kappa,\omega,\omega)$ to get a similar result for sigma
products of dimension $\kappa$.

\begin{ex}
  $\pl D \markwin \bellConGame{(\oneptcomp\kappa)^\omega}$
\end{ex}

\begin{proof}
  $\pl D \tactwin \bellConGame{\oneptcomp\kappa}$ + previous result
\end{proof}

\newpage

\begin{lem}
  $\pl O\prewin\gruClusGame{X}{S}$ if and only if $\pl O\prewin\gruConGame{X}{S}$.
\end{lem}

\begin{proof}
  Suppose that $\sigma$ is a predetermined winning strategy for
  $\gruClusGame{X}{S}$. Let $p$ attack $\sigma$, and $q$ be a subsequence of $p$.
  It follows that $q$ also attacks $\sigma$, so $q$ clusters at $S$. Thus $p$
  conveges to $S$, and $\sigma$ is a predetermined winning strategy for
  $\gruConGame{X}{S}$.
\end{proof}

\begin{thm}
  For any predetermined absolutely proximal space $X$,
  $\pl O\prewin \gruConGame{X}{H}$ for all compact $H\subseteq X$.
\end{thm}

\begin{proof}
  Let $\sigma(n)$ be a winning predetermined strategy for $\pl D$ in the
  absolutely proximal game such that
  $\sigma(n+1)\subseteq \sigma(n)$.
  For a given tree $T$, let $\max(T)$ denote its maximal nodes.

  \bigskip

  First we define $T(0)\subseteq \omega^{\leq 2}$.

  \begin{itemize}
    \item Let $\emptyset\in T(0)$.
    \item Choose
      \[
        m_{\emptyset}<\omega
      \]
      and for $i<m_{\emptyset}$ choose
      \[
        h_{\<i\>}\in H
      \]
      and for $i,j<m_{\emptyset}$ choose
      \[
        h_{\<i,j\>}
          \in
        H\cap\cl{\frac{1}{4}\sigma(0)[h_{\<i\>}]}
      \]
      such that
        \[
          \left\{\frac{1}{4}\sigma(0)[h_{\<i\>}]:i<m_{\emptyset}\right\}
        \]
      is a cover for $H$ and such that for each $i<m_{\emptyset}$
        \[
          \left\{\frac{1}{4}\sigma(1)[h_{\<i,j\>}]:j<m_{\emptyset}\right\}
        \]
      is a cover for
      $H\cap\cl{\frac{1}{4}\sigma(0)[h_{\<i\>}]}$.
    \item Let $\<i\>$ and $\<i,j\>$ be in $T(0)$ for $i,j<m_{\emptyset}$.
  \end{itemize}

  \bigskip

  Now suppose $T(n)\subseteq \omega^{\leq 2n+2}$ is defined.
  We then define $T(n+1)\subseteq \omega^{\leq 2n+4}$ as follows:

  \begin{itemize}
    \item Let $T(n)\subseteq T(n+1)$.
    \item For each $t\concat\<i,j\>\in\max(T(n))$, choose
        \[m_{t\concat\<i,j\>}<\omega\]
        and for $k<m_{t\concat\<i,j\>}$ choose
          \[
            h_{t\concat\<i,j,k\>}
              \in
            H\cap \cl{\frac{1}{4}\sigma(2n+2)[h_{t\concat\<i,j\>}]}
          \]
        and for $k,l<m_{t\concat\<i,j\>}$ choose
          \[
            h_{t\concat\<i,j,k,l\>}
              \in
            H\cap\cl{\frac{1}{4}\sigma(2n+3)[h_{t\concat\<i,j,k\>}]}
          \]
        such that
      \[
        \left\{\frac{1}{4}\sigma(2n+2)[h_{t\concat\<i,j,k\>}]
          :
        k<m_{t\concat\<i,j\>}\right\}
      \]
    is a cover for $H\cap \cl{\frac{1}{4}\sigma(2n+1)[h_{t\concat\<i,j\>}]}$,
    and such that for each $k<m_{t\concat\<i,j\>}$
      \[
        \left\{\frac{1}{4}\sigma(2n+3)[h_{t\concat\<i,j,k\>}]:l<m_t\right\}
      \]
    is a cover for $H\cap\cl{\frac{1}{4}\sigma(2n+2)[h_{t\concat\<i,j,k\>}]}$.
    \item For each $t\in\max(T(n))$ and each $i,j<m_t$, put $t\concat\<i\>$
      and $t\concat\<i,j\>$ in $T(n+1)$.
  \end{itemize}

  \bigskip

  We now define the predetermined strategy $\tau$ for $\pl O$ in
  $\gruClusGame{X}{H}$ such that:
  \[
    \tau(n)
      =
    \bigcup_{t\in\max(T(n))}
      \bigcap_{m<2n+2}
        \frac{1}{4}\sigma(m)[h_{t\rest m+1}]
  \]
  noting that $\tau(n)=\bigcap_{m\leq n}\tau(m)$ by definition.

  Since
  $\left\{\frac{1}{4}\sigma(2n+2)[h_{t\concat\<i\>}]:i<m_{t}\right\}$ is a cover
  for $H\cap \frac{1}{4}\sigma(2n+1)[h_{t}]$, since \\
  $\left\{\frac{1}{4}\sigma(2n+1)[h_{t\concat\<i,j\>}]:j<m_{t}\right\}$ is
  a cover for $H\cap \frac{1}{4}\sigma(2n)[h_{t\concat\<i\>}]$, and since
  $\left\{\frac{1}{4}\sigma(0)[h_{\<i\>}]:i<m_{\emptyset}\right\}$ is a
  cover for $H$, it follows that $\tau(n)$ contains $H$ and is $\tau$ is
  a legal strategy.

  Let $p$ be an attack against $\tau$ such that
  $p(n)\in \tau(n)$. If we can construct an attack $q$
  against $\sigma$ which shares a subsequence of $p$, then $p$ must cluster
  since $q$ must converge.
  To find such a $q$, we construct a subtree $T'\subseteq T$.

  We begin by setting $T'(0)=T(0)$.

  For $n<\omega$, suppose $T'(n)\subseteq T(n)$ is defined such that:
    \begin{itemize}
      \item If $t'\concat\<i,j\>\in T'(n)$, then $|t'|\leq 2n$
      \item If $s'\leq t'\in T'(n)$, then $s\in T'(n)$.
      \item If $t'\in T'(n)\setminus\max(T'(n))$ and $|t'|$ is even,
            then $t'\concat\<i,j\>\in T'(n)$ for $i,j<m_{t'}$.
    \end{itemize}

  Since $p(n)\in\tau(n)$, there exists some $t_n\in\max(T(n))$ such that
  \[
    p(n)
      \in
    \bigcap_{m<2n+2}
    \frac{1}{4}\sigma(m)[h_{t_n\rest m+1}]
  \]
  and in turn, there exists $t_n'\concat\<i,j\>\in\max(T'(n))$
  with $t_n'\concat\<i,j\>\leq t_n$ and
  \[
    p(n)
      \in
    \bigcap_{m<|t_n'|+2}
    \frac{1}{4}\sigma(m)[h_{t_n'\concat\<i,j\>\rest m+1}]
  \]
  since $|t_n'|\leq 2n$.

  Let $p_{t_n'\concat\<i,j\>}=p(n)$

  Take note that, in particular,
  \[
    p_{t_n'\concat\<i,j\>}
      \in
    \sigma(|t_n'|)[h_{t_n'\concat\<i\>}]
  \]
  and
  \[
    p_{t_n'\concat\<i,j\>}
      \in
    \frac{1}{4}\sigma(|t_n'|+1)[h_{t_n'\concat\<i,j\>}]
  \]

  We then define
  \[
    T'(n+1)
      =
    T'(n)
      \cup
    \{t_n'\concat\<i,j,k\>: k\leq m_{t_n'\concat\<i,j\>}\}
      \cup
    \{t_n'\concat\<i,j,k,l\>: k,l\leq m_{t_n'\concat\<i,j\>}\}
  \]
  while noting that for all $k\leq m_{t_n'\concat\<i,j\>}$,
  \[
    h_{t_n'\concat\<i,j,k\>}
      \in
    H\cap \cl{\frac{1}{4}\sigma(|t_n'|+2)[h_{t_n'\concat\<i,j\>}]}
      \subseteq
    \frac{1}{2}\sigma(|t_n'|+1)[h_{t_n'\concat\<i,j\>}]
  \]
  and thus
  \[
    h_{t_n'\concat\<i,j,k\>}\in\sigma(|t_n'|+1)[p_{t_n'\concat\<i,j\>}]
  \]

  Finally, we let $T'=\bigcup_{n<\omega}T'(n)$. Since
  $T'$ is an infinite tree with finite levels, we may
  pick an infinite branch $b$. From $b$, we construct the sequence
  \[
    q=\<h_{b\rest 1}, p_{b\rest 2}, h_{b\rest 3}, p_{b\rest 4}, \dots\>
  \]
  and claim it attacks $\sigma$ and thus must converge. If so, since
  $\<p_{b\rest2},p_{b\rest4},\dots\>$ is a subsequence of $p$, $p$ must
  cluster. To see this, recall that for some $t_n'$:
  \[
    p_{b\rest2n+2}
      =
    p_{t_n'\concat\<i,j\>}
      \in
    \sigma(|t_n'|)[h_{t_n'\concat\<i\>}]
  \]
  and
  \[
    h_{b\rest2n+3}
      =
    h_{t_n'\concat\<i,j,k\>}
      \in
    \sigma(|t_n'|+1)[p_{t_n'\concat\<i,j\>}]
  \]

  We have thus proven $\pl O\prewin\gruClusGame{X}{H}$, and thus
  $\pl O\prewin\gruConGame{X}{H}$.
\end{proof}

\newpage

\begin{ex}
  Let $X=I\times 2$ be the Alexandrov double interval. Then
  $\pl D \not\prewin \bellConGame{X}$, but $\pl D \markwin \bellConGame{X}$.
\end{ex}

\begin{proof}
  We assume that the uniformity on $X$ is given by entourages
  \[
    D(\epsilon,F) =
    \{\<x,0\>,\<y,0\>: |x-y|<\epsilon \}
      \cup
    \{\<x,1\>,\<y,0\>: |x-y|<\epsilon \vee x\not\in F\}
  \]
  \[
      \cup
    \{\<x,0\>,\<y,1\>: |x-y|<\epsilon \vee y\not\in F\}
      \cup
    \{\<x,1\>,\<y,1\>: x=y\}
  \]

  That is, points are $D(\epsilon,F)$-close if they are the same point, or
  the first coordinates are within $\epsilon$ of each other while neither
  second coordinate is in $F$.

  Suppose $\pl D$ had a predetermined winning strategy
  $\sigma(n)=D(\epsilon_n,F_n)$. Then
  $\pl P$ can choose $x\not\in \bigcup_{n<\omega} F_n$, and play $\<x,1\>$
  during even rounds, and $\<x_{2n+1},0\>$ where $|x-x_{2n+1}|<\epsilon_{2n}$
  during odd rounds, preventing convergence.

  However, assume $\pl D$ uses the Mark\"ov strategy
  $\sigma(x,n)=D(2^{-n},\{x\})$.
  If $\pl P$ repeats a point of the form $\<x,1\>$, then since
  $D(2^{-n},\{x\})[\<x,1\>]=\{\<x,1\>\}$, $\pl P$ must repeat $\<x,1\>$ for
  the rest of the game, and $\pl D$ wins. Otherwise, $\pl P$ cannot repeat
  points played in $I\times\{1\}$, and as the first
  coordinates form a Cauchy sequence and converge to some $z$, any open set
  about $\<z,0\>$ contains all but finitely many points of $\pl P$'s sequence,
  and $\pl D$ wins.
\end{proof}

\begin{thm}
  For any uniformly locally compact space $X$,
      $\pl D\prewin \bellConGame{X} \Leftrightarrow \pl D\prewin \bellClusGame{X}$
\end{thm}

\begin{proof}
  Let $M$ be a uniformly locally compact entourage.
  Let $\sigma$ witness $\pl D\prewin \bellConGame{X}$ such that
  $\sigma(n)\subseteq M$ always
  (so $\cl{\sigma(a)[x]}\subseteq\cl{M[x]}$ is compact),
  $\sigma(n+1)\subseteq\frac{1}{4}\sigma(n)$.

  Let $\tau(n)=\frac{1}{2}\sigma(n)$. If $p$ attacks $\tau$ in
  $\bellClusGame{X}$, then
    \[
      p(n+1)
        \in
      \tau(n)[p(n)]
        =
      \frac{1}{2}\sigma(n)[p(n)]
    \]

    and for

    \[
      x
        \in
      \cl{\sigma(n+1)[p(n+1)]}
        \subseteq
      \cl{\frac{1}{4}\sigma(n)[p(n+1)]}
        \subseteq
      \frac{1}{2}\sigma(n)[p(n+1)]
    \]

  we can conclude $x\in\sigma(n)[p(n)]$. Thus

    \[
      \sigma(n+1)[p(n+1)]
        \subseteq
      \cl{\sigma(n+1)[p(n+1)]}
        \subseteq
      \sigma(n)[p(n)]
    \]

  Finally, note that $p$ attacks the winning strategy $\sigma$ in
  $\bellConGame{X}$, but since the intersection of a chain of nonempty compact
  sets is nonempty:

    \[
      \bigcap_{n<\omega} \sigma(n)[p(n)]
        =
      \bigcap_{n<\omega} \cl{\sigma(n)[p(n)]}
        \not=
      \emptyset
    \]

  We conclude that $p$ converges.
\end{proof}

\begin{prop}
If $\pl D \prewin \bellConGame{X}$, then $X$ has a $G_\delta$ diagonal.
\end{prop}

\begin{proof}
  If $\pl D \prewin \bellConGame{X}$ with strategy $\sigma$, then consider
  $\<x,y\>\in \bigcap_{n<\omega} \sigma(n)$. It follows that
  $\<x,y,x,y,\dots\>$ attacks $\sigma$, and
    $
      \{x,y\}\subseteq \bigcap_{n<\omega}
      \sigma(n)[x]\cap\bigcap_{n<\omega}\sigma(n)[y]
      \not=0
    $
  so it must converge, and $x=y$. Thus
  $\bigcap_{n<\omega}\sigma(n) = \Delta$ is $G_\delta$.
\end{proof}

\begin{ex}
The Sorgenfrey line $S$ has a $G_\delta$ diagonal but $\pl P\win\bellConGame{S}$.
\end{ex}

\begin{cor}
  For $X$ with uniformity $\mathbb{D}$ inducing the compact Hausdorff topology
  $\tau$, the following are equivalent:
    \begin{enumerate}[(a)]
      \item $\pl D \prewin \bellConGame{X}$
      \item $\pl D \prewin \bellClusGame{X}$
      \item $X$ has a $G_\delta$ diagonal
      \item $\mathbb{D}$ is metrizable
      \item $\tau$ is metrizable
    \end{enumerate}
\end{cor}

\begin{proof}
  For compact Hausdorff spaces, it is well known that there is exactly one
  uniformity inducing the topology. Thus $(d)\Leftrightarrow(e)$. Since $X$
  is uniformly locally compact, $(a)\Leftrightarrow(b)$. Also, compact spaces
  with a $G_\delta$ diagonal are metrizable, so $(c)\Rightarrow(e)$.
  Bell noted $(d)\Rightarrow(a)$
  for arbitrary uniform spaces, and the previous proposition shows
  $(a)\Rightarrow(c)$.
\end{proof}

\begin{thm}
  A uniformly locally compact space with a $G_\delta$ diagonal is metrizable.
\end{thm}

\begin{proof}
  Based on several folklore results.

  Uniformly locally compact implies the topological sum of $\sigma$-compact
  spaces implies paracompact. Locally compact plus $G_\delta$ diagonal implies
  locally metrizable. Locally metrizable plus paracompact characterizes
  metrizable.
\end{proof}

\begin{cor}
  If $X$ is uniformly locally compact, then $\pl D \prewin \bellConGame{X}$
  implies $X$'s topology is metrizable.
\end{cor}

\newpage

\begin{ex}
  Let $R$ be the Michael Line. Then $\pl P\win \bellConGame{X}$.
\end{ex}

\begin{proof}
  During round $0$, $\pl P$ may choose $m(0)=0$ and $p(0)=1$,
  and during round $n+1$,
  $\pl P$ may choose $m(n+1)>m(n)$ and
  $p(n+1)=p(n)+\frac{1}{10^{m(n+1)}}$ such that $p$ is a legal attack.

  It follows that $p$ ``converges'' to
  $x=\sum_{n<\omega}\frac{1}{10^{m(n)}}$, except $x$ is an irrational
  number composed of $1$s separated by strings of $0$s of strictly
  increasing size.
\end{proof}

\begin{ex}
  Let $\omega_1$ be given a ladder topology:
    \begin{itemize}
      \item All successor ordinals are isolated.
      \item Strictly increasing sequences (ladders) $L_\alpha:\omega\to\alpha$
            are defined for each limit ordinal $\alpha$ such that $L_\alpha$ converges to $\alpha$ in the order topology, and each limit
            $\alpha$ is given neighborhoods of the form
            $L(\alpha,m)=\{\alpha\}\cup\{L_\alpha(n):n\geq m\}$.
      \item $\omega_1=\bigcup_{\alpha\in\omega_1^L} L(\alpha,0)$
    \end{itemize}

  Let
    \[
      A(\alpha,n)
        =
      [L(\alpha,0)\setminus L(\alpha,n)]^1
        \cup
      \{\oneptcomp\omega_1 \setminus (L(\alpha,0)\setminus L(\alpha,n))\}
    \]
    \[
      B(\alpha)
        =
      \{L(\alpha,0), \oneptcomp\omega_1 \setminus L(\alpha,0)\}
    \]

  Finite refinements of $A(\alpha,n)$ and $B(\alpha)$ give partitions
  witnessing a uniformization of the ladder topology.

  Then $\bellConGame{\oneptcomp\omega_1}$ is indetermined.
\end{ex}

\begin{proof}
  $\pl D\not\win\bellConGame{\oneptcomp\omega_1}$ since $\oneptcomp\omega_1$
  isn't Corson compact.

  Let $\sigma$ be a perfect information strategy for $\pl P$.

  Let
  \[
    C(\sigma)
      =
    \{
      \delta\in\omega_1^L
        :
      \omega_1\setminus D[\infty] \subseteq \delta
        \Rightarrow
      \sigma(D) < \delta
    \}
  \]
  and note that by the closing up lemma, $C(\sigma)$ is always club.

  We define a counterattack for $\pl D$ with $D_0=D(F_0,0)$ and
  $F_0=\emptyset$ as her
  initial move, and let $p_0=\sigma(D_0)$ be the initial move of $\pl P$.
  We fix $p_0<\delta_0<\delta_1<\dots\in C(\sigma)$ such that
  $\lim_{n\to\infty}\delta_n=\delta_\infty\in C$. Enumerate all the
  limit ordinals $<\delta_n$ as $\alpha_{n,m}$ for $m<\omega$. Let
  $i_n<\omega$ be the least integer such that
  $L_{\delta_\infty}(i_n)>\delta_n$.

  If $D_0,\dots,D_n$ are the first $n$ moves of $\pl D$, and
  $p_0,\dots,p_n$ are the first $n$ moves of $\pl P$ determined by $\sigma$,
  then we define
    \[
      D_{n+1}
        =
      \bigwedge_{j\leq n}
      \left(
        A(\delta_j,n)
          \wedge
        B(\delta_j)
      \right)
        \wedge
      \bigwedge_{j,k\leq n}
      \left(
        A(\alpha_{j,k},n)
          \wedge
        B(\alpha_{j,k})
      \right)
        \wedge
      A(\delta_\infty,i_n)
    \]

  (TODO: show this is a winning counterattack)
\end{proof}