### 2014-11-25 phd/dissertation/chapter5.tex
begin work on chapter 5

%!TEX root = dissertation.tex
% ^ leave for LaTeXTools build functionality

\chapter{Gruenhage's Locally Finite Games}

A variation of $\gruConGame{X}{x}$, also due to Gruenhage,
may be used to characterize various covering properties, particularly
for locally compact spaces. We assume all spaces are $T_1$.

\section{Characterizations using $\gruKPGame{X}$, $\gruKLGame{X}$}

\begin{game}
  Let $\gruKPGame{X}$ denote the \term{Gruenhage compact/point game}
  with players $\pl K$, $\pl P$. During round $n$, $\pl K$ chooses
  a compact subset $K_n$ of $X$, followed by $\pl P$ choosing a point
  $p_n\in X$ such that $p_n\not\in \bigcup_{m\leq n}K_m$.

  $\pl K$ wins the game if the collection $\{\{p_n\}:n<\omega\}$ is locally
  finite in the space, and $\pl P$ wins otherwise.
\end{game}

This game is often formulated by requiring that the collection
$\{\{p_n\}:n<\omega\}$ be discrete. With the knowledge of at least the latest
move of the opponent, $\pl K$ may guarantee that if $\{\{p_n\}:n<\omega\}$
is locally finite then it is also discrete, since she may require
$p_n\in K_{n+1}$. Thus this formulation is essentially equivalent to the usual
formulation for all existing applications.

We may relate this game to $\gruConGame{X}{x}$ as follows:

\begin{thm}
  If $X$ is locally compact, then for $k<2$:
    \begin{itemize}
      \item $\pl K\win\gruKPGame{X}$ if and only if
        $\pl O\win\gruConGame{\oneptcomp{X}}{\infty}$.
      \item $\pl K\kmarkwin{k}\gruKPGame{X}$ if and only if
        $\pl O\kmarkwin{k}\gruConGame{\oneptcomp{X}}{\infty}$.
      \item $\pl K\ktactwin{k}\gruKPGame{X}$ if and only if
        $\pl O\ktactwin{k}\gruConGame{\oneptcomp{X}}{\infty}$.
    \end{itemize}
\end{thm}

\begin{proof}
  Let $L$ be any of $\nu_0$, $\nu_1$, $\mu_0$, $\mu_1$, or the identity.
  For any sequence $s$ of points in $\oneptcomp X$, let $s'$ be the
  subsequence of non-$\infty$ points in $s$.

  If $\sigma\circ L$ is a winning strategy for $\pl K$ in $\gruKPGame{X}$, let
  $\tau\circ L$ be a strategy for $\pl O$ in $\gruConGame{\oneptcomp X}{\infty}$
  such that $\tau(L(s)) = \oneptcomp X\setminus \sigma(L(s)')$. Then for
  any legal attack $p$ against $\tau$, $p'$ is a legal attack against $\sigma$.
  (The proof of this claim uses the fact that $k<2$.)
  If $p'$ is a finite sequence, then $p$ converges to $\infty$.
  Otherwise, the set $\{\{p'(n)\}:n<\omega\}$ is locally finite in $X$, so
  $\{p'(n):n<\omega\}$ is a closed discrete subset of $X$. Then for every
  neighborhood $U$ of $\infty$, $X\setminus U$ is contained in a compact set,
  so it cannot contain a closed discrete subset. Thus $p'$ and $p$
  converge to $\infty$.

  If $\sigma\circ L$ is a winning strategy for $\pl O$ in
  $\gruConGame{\oneptcomp X}{\infty}$, let $\tau\circ L$ be a strategy for
  $\pl K$ in $\gruKPGame{X}$ such that $\tau(L(s))= X\setminus\sigma(L(s))$.
  Then for any legal attack $p$ against $\tau$, $p$ is a legal attack against
  $\sigma$, so the sequence $p$ converges to $\infty$. For any point $x\in X$
  distinct from the $p(n)$, we may choose a neighborhood $U_x$ of $x$ in $X$
  missing every point in $p(n)$.
  For every $n<\omega$, we may choose a neighborhood $U_{p(n)}$ of $p(n)$ in
  $X$ which misses every distinct $p(m)$. Thus $\{\{p(n)\}:n<\omega\}$ is a
  locally finite collection.
\end{proof}

The reason why $\gruConGame{\oneptcomp X}{\infty}$ and $\gruKPGame{X}$ are
not completely equivalent is due to the fact that $\pl P$ may hide information
from $\pl O$ by playing $\infty$. These moves cannot trivially be ignored
using a $k$-limited information strategy for $k\geq 2$, since $\pl P$ may
ensure that only one ``useful'' move may be seen by $\pl K$ at a time by playing
$k$ $\infty$s between each non-$\infty$.

Applications of limited information strategies for $\pl K$ are already known;
see \cite{MR858337}.

\begin{thm}
  The following are equivalent for a locally compact space $X$:
    \begin{itemize}
      \item $X$ is metacompact
      \item $\pl K \tactwin \gruKPGame{X}$.
    \end{itemize}
\end{thm}

\begin{thm}
  The following are equivalent for a locally compact space $X$:
    \begin{itemize}
      \item $X$ is $\sigma$-metacompact
      \item $\pl K \markwin \gruKPGame{X}$.
    \end{itemize}
\end{thm}

In addition, it's trivial to show the following.

\begin{prop}
  The following are equivalent for any space $X$:
    \begin{itemize}
      \item $X$ is compact
      \item $\pl K \ktactwin{0} \gruKPGame{X}$.
    \end{itemize}
\end{prop}

\begin{proof}
  A $0$-tactic is seeded with zero information about the moves of the opponent
  or the round number, so it must be a constant function valued at $X$.
\end{proof}

A similar game may be considered by allowing the second player to choose
compact sets rather than points, which also provided in \cite{MR858337}
a game-theoretic characterization of a covering property for locally compact
spaces.

\begin{game}
  Let $\gruKLGame{X}$ denote the \term{Gruenhage compact/compact game}
  with players $\pl K$, $\pl L$. This game proceeds analogously to
  $\gruKPGame{X}$, except the second player $\pl L$ chooses comapct sets $L_n$
  missing $\bigcup_{m\leq n}K_n$,
  and $\pl K$ wins if the collection $\{L_n:n<\omega\}$ is locally finite.
\end{game}

As above, this formulation is equivalent to requiring $\{L_n:n<\omega\}$ be
discrete when considering strategies for $\pl K$ which use at least the latest
move of $\pl L$.

\begin{thm}
  The following are equivalent for a locally compact space $X$:
    \begin{itemize}
      \item $X$ is paracompact
      \item $\pl K \win \gruKLGame{X}$.
    \end{itemize}
\end{thm}

\section{Predetermined strategies for $\gruKPGame{X}$, $\gruKLGame{X}$}

Adding knowledge of the round number to a tactic changes the characterization
from metacompact to $\sigma$-metacompact. In fact, the analogous result holds
for $0$-tactics to $0$-marks, known as predetermined strategies since they
rely only on the round number and not the moves of the opponent.

\begin{thm}
  If $X$ is a locally compact Lindel\"of space, then $\pl K \prewin G_{K,L}(X)$.
\end{thm}

\begin{proof}
  For each $x\in X$, let $U_x$ be an open neighborhood of $x$ with $\cl{U_x}$
  compact. Then as $X$ is Lindel\"of, choose $x_n\in X$ for $n<\omega$ such that
  $\{U_{x_n}:n<\omega\}$ covers $X$. Define the predetermined strategy $\sigma$
  for $\pl K$ by $\sigma(n)=\cl{U_{x_n}}$.

  Let $L:\omega\to \mc K(X)$ legally attack $\sigma$, so
  $L(n)\cap\bigcup_{m\leq n}\sigma(m)=\emptyset$. For each $x\in X$,
  choose $n<\omega$ with $x\in U_{x_n}$. Then $U_{x_n}$ is a neighborhood of
  $x$ which intersects finitely many $L(n)$, so $\{L(n):n<\omega\}$ is
  locally finite.
\end{proof}

\begin{defn}
  A space $X$ is \textbf{hemicompact} if there exist compact sets $K_n$ for
  $n<\omega$ such that every compact set in $X$ is a subset of some $K_n$.
\end{defn}

\begin{thm}
  If $\pl K\prewin \gruKPGame{X}$, then $X$ is hemicompact.
\end{thm}

\begin{proof}
  Let $\sigma$ be a winning predetermined strategy for $\pl K$ in
  $\gruKPGame{X}$. If $C\in \mc K(X)$ is compact, then for each $x\in C$ let
  $U_x$ be an open neighborhood of $x$ which intersects finitely many
  $\sigma(n)$. Thus there exists an open neighborhood of $C$ which intersects
  finitely many $\sigma(n)$.
\end{proof}

\begin{cor}
  The following are equivalent for any locally compact space $X$:
    \begin{itemize}
      \item $X$ is Lindel\"of.
      \item $X$ is $\sigma$-compact.
      \item $X$ is hemicompact.
      \item $\pl K \prewin \gruKPGame{X}$.
      \item $\pl K \prewin \gruKLGame{X}$.
    \end{itemize}
\end{cor}

\section{Compactly generated spaces}

\begin{defn}
  A space $X$ is \term{compactly generated} if a set is closed if and
  only if its intersection is closed in every compact subspace. Such
  spaces are also known as \term{$k$-spaces}.
\end{defn}

All locally compact spaces are $k$-spaces. As will be shown,
the games $\gruKPGame{X}$, $\gruKLGame{X}$ are equivalent for $\pl K$'s
predetermined strategies in Hausdorff $k$-spaces.

\begin{defn}
  A space $X$ is a $k_\omega$-space if there exist compact sets $K_n$ for
  $n<\omega$ such that a set is closed if and
  only if its intersection is closed in every subspace $K_n$.
\end{defn}

\begin{thm}
  If $X$ is a Hausdorff hemicompact $k$-space, then
  $\pl K \prewin \gruKLGame{X}$.
\end{thm}

\begin{proof}
  Let $K_n$ witness hemicompactness for $X$. Define the predetermined strategy
  $\sigma$ for $\pl K$ by $\sigma(n)=K_n$.

  Let $L:\omega\to\mc{K}(X)$ be a legal attack against $\sigma$, and let
  $L_\omega = \bigcup_{n<\omega}L_(n)$. For each compact set (TODO FINISH)
\end{proof}

\begin{cor}
  The following are equivalent for any Hausdorff $k$-space $X$:
    \begin{itemize}
      \item $X$ is hemicompact.
      \item $X$ is $k_{\omega}$.
      \item $\pl K \prewin \gruKPGame{X}$.
      \item $\pl K \prewin \gruKLGame{X}$.
    \end{itemize}
\end{cor}

% \section{Non-locally compact spaces}

% \begin{prop}
%   If $X=\omega\cup\mathcal{F}\subset\beta\omega$ (a non-$k$ space),
%   then $K\not\prewin \gruKLGame{X}$.
% \end{prop}

% \begin{prop}
%   If a selective ultrafilter $\mathcal{F}$ exists (independent of ZFC),
%   then $K \not\prewin \gruKPGame{X}$
%   for $X=\omega \cup \{\mathcal{F}\}\subset \beta\omega$.
% \end{prop}

% \begin{thm}
%   Then there is an ultrafilter $\mathcal{F}$ such that
%   then $K \prewin \gruKPGame{X}$ and $K \tactwin \gruKPGame{X}$
%   for $X=\omega \cup \{\mathcal{F}\}\subset \beta\omega$.
% \end{thm}

% \begin{thm}
%   Let $M$ be the \term{metric fan}, a non-locally compact $k$ space.
%   Then $P \win \gruKPGame{M}$.
% \end{thm}

% \begin{thm}
%   Let $S$ be the \term{sequential fan}, a non-locally compact $k$ space.
%   Then $K \prewin \gruKPGame{S}$ and $K\tactwin \gruKPGame{S}$.
% \end{thm}

% (TODO: Maybe consider claim
% $K \prewin \gruKPGame{X}$ $\Rightarrow$ $X$ is a $k$-space)

% \section{Cantor tree space example}

% \begin{ex}
%   Let $X$ be a zero-dimensional, compact L-space
%   (hereditarally Lindeloff and non-separable).
%   It is a fact that there exists a point-countable collection
%   $\mathcal{U}=\{U_\alpha : \alpha<\omega_1\}$ of clopen sets in $X$, and it is
%   also true that any point-finite subcollection of $\mathcal{U}$ is countable.

%   Let $C = \{c_\alpha : \alpha <\omega_1\}$ be any uncountable subset of the
%   Cantor space $2^\omega$. Let $X_s = X \times \{s\}$ for each
%   $s \in 2^{<\omega}$, and $U_{\alpha,s} = U_\alpha \times \{s\}$.

%   Finally, let \[\mathbb{X} = C \cup \bigcup_{s\in 2^{<\omega}} X_s\] be a
%   tree of $2^{<\omega}$ copies of $X$, and where
%   \[c_\alpha \cup \bigcup_{n < \omega} U_{\alpha, x_\alpha \restriction n}\]
%   is an open set about each $c_\alpha$.
% \end{ex}

% \begin{prop}
% $K \win \gruKPGame{\mathbb{X}}$.
% \end{prop}

% \begin{thm}(cor of G, game-theoretic proof by me)
% $K\not\tactwin\gruKPGame{\mathbb{X}}$.
% \end{thm}

% \begin{thm}
% $K\not\ktactwin{k}\gruKPGame{\mathbb{X}}$.
% \end{thm}