### 2014-12-04 phd/dissertation/abstract.tex
finish first draft of dissertation

%!TEX root = dissertation.tex
% ^ leave for LaTeXTools build functionality

\begin{abstract}
A topological game $G(X)$ is a two-player game which characterizes
properties of a topological space $X$ based upon the existence of winning
perfect-information
strategies for players in the game. If the property $P$ is characterized by a
player having a winning perfect information strategy, then a property
stronger than $P$ is characterized by that player having a winning
limited information strategy. This paper investigates the existence of
limited information strategies for four types of topological games
from the literature, and the topological properties
characterized by those strategies.
\end{abstract}