### 2014-12-04 phd/dissertation/acknowledgments.tex

%!TEX root = dissertation.tex
% ^ leave for LaTeXTools build functionality

\begin{acknowledgments}
I'm very grateful for the support of all my friends and family during
the completion of this manuscript, particularly for all the faculty members
and graduate student colleagues I've worked with over the years at
Auburn. I'd like especially to thank my
adviser Gary Gruenhage for all his support and insight on writing this
paper and beginning my mathematical career.

This dissertation is dedicated to my wife Jessica, for her love and
encouragement.
\end{acknowledgments}