### 2015-03-06 phd/talks-and-presentations/2015-03-11-defense.tex
update through chapter 4 intro

% $Header: /home/vedranm/bitbucket/beamer/solutions/generic-talks/generic-ornate-15min-45min.en.tex,v 90e850259b8b 2007/01/28 20:48:30 tantau $

\documentclass{beamer}

% This file is a solution template for:

% - Giving a talk on some subject.
% - The talk is between 15min and 45min long.
% - Style is ornate.

% Copyright 2004 by Till Tantau <tantau@users.sourceforge.net>.
%
% In principle, this file can be redistributed and/or modified under
% the terms of the GNU Public License, version 2.
%
% However, this file is supposed to be a template to be modified
% for your own needs. For this reason, if you use this file as a
% template and not specifically distribute it as part of a another
% package/program, I grant the extra permission to freely copy and
% modify this file as you see fit and even to delete this copyright
% notice.

\mode<presentation>
{
  \usetheme{Warsaw}
  % or ...

  \setbeamercovered{transparent}
  % or whatever (possibly just delete it)
}

      \newtheorem{proposition}[theorem]{Proposition}
      \theoremstyle{definition}
      \newtheorem{game}[theorem]{Game}
      \newtheorem{question}[theorem]{Question}

\usepackage[english]{babel}
% or whatever

\usepackage[latin1]{inputenc}
% or whatever

\usepackage{times}
\usepackage[T1]{fontenc}
% Or whatever. Note that the encoding and the font should match. If T1
% does not look nice, try deleting the line with the fontenc.

\usepackage{marvosym} % For \Smiley
\usepackage{verbatim} % for \verbatiminput

\usepackage{tikz}
\usetikzlibrary{matrix} % for diagrams

\title
{Limited information strategies for topological games}

\subtitle
{PhD Defense} % (optional)

\author%[Author, Another] % (optional, use only with lots of authors)
{Steven~Clontz~http://stevenclontz.com}%\inst{1} \and S.~Another\inst{2}}
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[Auburn University] % (optional, but mostly needed)
{
  %\inst{1}%
  Department of Mathematics and Statistics\\
  Auburn University}
  %\and
  %\inst{2}%
  %Department of Theoretical Philosophy\\
  %University of Elsewhere}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date[15-03-11] % (optional)
{March 11, 2015}

% If you have a file called "university-logo-filename.xxx", where xxx
% is a graphic format that can be processed by latex or pdflatex,
% resp., then you can add a logo as follows:

 % \pgfdeclareimage[height=1cm]{university-logo}{auburn_logo.png}
 % \logo{\pgfuseimage{university-logo}}

% Delete this, if you do not want the table of contents to pop up at
% the beginning of each subsection:
%\AtBeginSubsection[]
%{
%  \begin{frame}<beamer>{Outline}
%    \tableofcontents[currentsection,currentsubsection]
%  \end{frame}
%}

% If you wish to uncover everything in a step-wise fashion, uncomment
% the following command:

%\beamerdefaultoverlayspecification{<+->}

% Strategy uparrow shortcuts
\newcommand{\win}{\uparrow}
\newcommand{\prewin}{\underset{\text{pre}}{\uparrow}}
\newcommand{\markwin}{\underset{\text{mark}}{\uparrow}}
\newcommand{\tactwin}{\underset{\text{tact}}{\uparrow}}
\newcommand{\kmarkwin}[1]{\underset{#1\text{-mark}}{\uparrow}}
\newcommand{\ktactwin}[1]{\underset{#1\text{-tact}}{\uparrow}}
\newcommand{\codewin}{\underset{\text{code}}{\uparrow}}
\newcommand{\limitwin}{\underset{\text{limit}}{\uparrow}}
\newcommand{\notwin}{\not\uparrow}
\newcommand{\notprewin}{\underset{\text{pre}}{\not\uparrow}}
\newcommand{\notmarkwin}{\underset{\text{mark}}{\not\uparrow}}
\newcommand{\nottactwin}{\underset{\text{tact}}{\not\uparrow}}
\newcommand{\notkmarkwin}[1]{\underset{#1\text{-mark}}{\not\uparrow}}
\newcommand{\notktactwin}[1]{\underset{#1\text{-tact}}{\not\uparrow}}
\newcommand{\notcodewin}{\underset{\text{code}}{\not\uparrow}}
\newcommand{\notlimitwin}{\underset{\text{limit}}{\not\uparrow}}

\newcommand{\oneptcomp}[1]{#1^*}
\newcommand{\oneptlind}[1]{#1^\dagger}
% \newcommand{\sharp}[1]{#1^{\#}}

% Games
\newcommand{\gruConGame}[2]{Gru_{O,P}^{\to}\left({#1},{#2}\right)}
\newcommand{\gruConGameHard}[2]{Gru_{O,P}^{\to,\star}\left({#1},{#2}\right)}
\newcommand{\gruClusGame}[2]{Gru_{O,P}^{\leadsto}\left({#1},{#2}\right)}
\newcommand{\gruClusGameHard}[2]{Gru_{O,P}^{\leadsto,\star}\left({#1},{#2}\right)}

\newcommand{\gruKPGame}[1]{Gru_{K,P}\left({#1}\right)}
\newcommand{\gruKLGame}[1]{Gru_{K,L}\left({#1}\right)}

\newcommand{\cloPFGame}[1]{PtFin_{F,C}\left({#1}\right)}

\newcommand{\menGame}[1]{Men_{C,F}\left({#1}\right)}
\newcommand{\rothGame}[1]{Roth_{C,S}\left({#1}\right)}
\newcommand{\rothAltGame}[1]{Roth_{P,O}\left({#1}\right)}

\newcommand{\cloFillStrictGame}[1]{Fill^{\cup,\subset}_{C,F}\left({#1}\right)}
\newcommand{\cloFillGame}[1]{Fill^{\cup,\subseteq}_{C,F}\left({#1}\right)}
\newcommand{\cloFillInitialGame}[1]{Fill^{1,\subseteq}_{C,F}\left({#1}\right)}
\newcommand{\cloFillIntGame}[1]{Fill^{\cap}_{C,F}\left({#1}\right)}

\newcommand{\bellUniGame}[1]{Bell_{D,P}^{\textrm{uni}}\left({#1}\right)}
\newcommand{\bellConGame}[1]{Bell_{D,P}^{\rightharpoonup}\left({#1}\right)}
\newcommand{\bellConHardGame}[1]{Bell_{D,P}^{\rightharpoonup,\star}\left({#1}\right)}
\newcommand{\bellAbsConGame}[1]{Bell_{D,P}^{\to}\left({#1}\right)}
\newcommand{\bellAbsConHardGame}[1]{Bell_{D,P}^{\to,\star}\left({#1}\right)}

\newcommand{\SigmaProdR}[1]{\Sigma\mathbb{R}^{#1}}
\newcommand{\sigmaprodtwo}[1]{\Sigma2^{#1}}

\newcommand{\concat}{{^\frown}}
\newcommand{\rest}{\restriction}

\newcommand{\cl}[1]{\overline{#1}}

\newcommand{\pow}[1]{\mc{P}(#1)}

\newcommand{\<}{\langle}
\renewcommand{\>}{\rangle}

\newcommand{\al}[1]{{#1}^*}

\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\mb}[1]{\mathbb{#1}}

\newcommand{\po}{\mathbb{P}}
\newcommand{\pok}{\po_\kappa}

\newcommand{\Lim}{\mathrm{Lim}}
\newcommand{\Suc}{\mathrm{Suc}}

\newcommand{\ds}{\displaystyle}

\newcommand{\st}[2]{st\left(#1,#2\right)}

\newcommand{\alcomp}{\al\parallel}

\newcommand{\rank}{\textrm{rank}}
\newcommand{\dom}{\textrm{dom}}

\renewcommand{\mod}{\,\textrm{mod}}

\newcommand{\zip}{\bowtie}
\newcommand{\ran}[1]{\text{range}(#1)}

\newcommand{\cf}[1]{\textrm{cf}(#1)}

\newcommand{\alcompS}[1]{S(#1)}

\newcommand{\scish}{almost-$\sigma$-(relatively compact)}

\usepackage{mathrsfs}
\newcommand{\pl}[1]{\mathscr{#1}}

\newcommand{\term}{\textit}

\newcommand{\bmGame}[1]{{BM}_{E,N}(#1)}

\begin{document}
\renewcommand{\pause}{}
\newcommand{\vspacing}{\vspace{1em}}
\newcommand{\vpause}{\pause\vspacing}

\begin{frame}
  \titlepage
\end{frame}

\section{Topological Games}

\subsection{Defintion}

\begin{frame}
  A \term{topological game} is a two-player game $G(X)$ of length
  $\omega=\{0,1,2,\dots\}$ defined for certain topological spaces $X$.

  \vpause

  During each round $n$, the first and second player take turns choosing
  certain topological objects from $X$ (e.g. point, open set, open cover, etc.).

  \vpause

  At the ``end'' of the game, a winner is declared by inspecting the sequences
  of choices made throughout the game.

  \vpause

  The study of such games involves finding when a player has a
  \term{winning strategy} which defeats every possible counterattack by
  the opponent.

  \vpause

  {\tiny See Telgarsky's excellent survey on topological games for more
  details: \cite{MR892457}}
\end{frame}

\subsection{Example}

\begin{frame}
  \small
  \begin{game}
    The \term{Banach-Mazur Game}
    $\bmGame{X}$ (1935) \cite{MR666400}

    \begin{figure}
      \includegraphics[width=0.6\linewidth]{bmGame.pdf}
    \end{figure}

  The first player $\pl N$ wins the game if the intersection of all the chosen
  open sets is nonempty.
  \end{game}
\end{frame}

\begin{frame}
  \begin{theorem}
    $X$ is Baire if and only if $\pl N$ lacks a winning strategy
    in the Banach Mazur game ($\pl N\not\win\bmGame{X}$).
  \end{theorem}

  \pause

  Thus the topological property of being a Baire space has a game-theoretic
  characterization using $\bmGame{X}$.

  \vpause

  By considering \term{limited information strategies}, we may characterize
  more properties.
\end{frame}

\begin{frame}
  Consider the following:

  \begin{theorem}
    $X$ is $\alpha$-favorable $\Rightarrow$
    $X$ is weakly $\alpha$-favorable $\Rightarrow$
    $X$ is Baire
  \end{theorem}

  \pause

  $\alpha$-favorability is characterized by $\pl E\tactwin\bmGame{X}$:
  player $\pl E$ has a \term{tactical} winning strategy which only considers
  the most recent move of the opponent.

  \vpause

  This is stronger than weak
  $\alpha$-favorability, characterized by $\pl E\win\bmGame{X}$.
  In this case $\pl E$ still has a winning strategy, but it may
  rely on perfect information of the history of the game.
\end{frame}

\subsection{Motivation}

\begin{frame}
  By characterizing topological properties using the theory of
  topological games, we introduce new proof techniques for demonstrating
  the structure of given topological spaces.

  \vpause

  The aim of my dissertation was to investigate four topological games
  from the literature with unknown limited information implications.

  \vpause

  In doing so I uncovered several new results in general topology,
  advancing research done by
  G. Gruenhage, P. Nyikos, F. Gavin, R. Telg{\' a}rksy, J. Bell,
  M. Scheepers, and others.
\end{frame}

\section{Gruenhage's Convergence/Clustering Games}

\subsection{Definition}

\begin{frame}
  \small
  \begin{game}
  Gruenhage's convergence game $\gruConGame{X}{x}$
  and clustering game $\gruClusGame{X}{x}$ proceed as follows:
    \begin{figure}
      \includegraphics[width=0.6\linewidth]{convergenceGame.pdf}
    \end{figure}

  $\pl O$ wins the game if the points chosen by $\pl P$ converge/cluster
  to the given point $x\in X$. Otherwise, $\pl P$ wins.
  \end{game}

  \tiny
  Note that $\pl O$ need not know anything about the history of the game
  to play each round.
\end{frame}

\subsection{Motivation}

\begin{frame}
  If $\pl O\win \gruConGame{X}{x}$, then $x$ is called a $W$-point in $X$.
  Obviously, all points of first-countablity are $W$-points, but
  $\pl O\win\gruConGame{\oneptcomp\kappa}{\infty}$ also, where $\infty$
  is the added point in the one-point compactification $\oneptcomp\kappa$
  of uncountable discrete $\kappa$.

  \vpause

  Points of first-countability may in fact be characterized by this game
  as well:
  \begin{theorem}
    $x$ has a countable local base in $X$ if and only if
    $\pl O\prewin \gruConGame{X}{x}$ ($\pl O$ has
    a winning \term{predetermined}
    strategy using only the round number).
  \end{theorem}

  \pause

  If every point in a space $X$ is a $W$-point, then $X$ is a $W$-space.
\end{frame}

\subsection{Results}

\begin{frame}
  A variation of this game which is harder for $\pl O$ yields some
  difficult infinite combinatorial questions:

  \begin{game}
  Gruenhage's hard convergence game $\gruConGameHard{X}{x}$
  and hard clustering game $\gruClusGameHard{X}{x}$ proceed as follows:
    \begin{figure}
      \includegraphics[width=0.6\linewidth]{convergenceHardGame.pdf}
    \end{figure}
  \end{game}
\end{frame}

\begin{frame}
  Nyikos observed in \cite{MR1031771} that:
  \begin{theorem}
    $\pl O \notmarkwin \gruConGameHard{\oneptcomp\omega_1}{\infty}$.

    {\small ($\pl O$ cannot guarantee a win using a \term{Markov strategy}
    which considers only the round number and most recent move.) }
  \end{theorem}

  \vpause

  Some more work shows that in fact
  \begin{theorem}
    $\pl O \notkmarkwin{k} \gruConGameHard{\oneptcomp\omega_1}{\infty}$.

    {\small ($\pl O$ cannot guarantee a win using a $k$-\term{Markov strategy}
    which considers only the round number and $k$ most recent moves.) }
  \end{theorem}
\end{frame}

\begin{frame}
  Interestingly, the strategy which prevents convergence won't
  prevent clustering as well unless the cardinality of the space is
  sufficiently large.

  \begin{theorem}
    $\pl O\markwin \gruClusGameHard{\oneptcomp\omega_1}{\infty}$, but
    $\pl O\notkmarkwin{k}\gruClusGameHard{\oneptcomp\omega_2}{\infty}$.
  \end{theorem}

  \pause

  But knowledge of the round number is used non-trivially in doing so.

  \begin{theorem}
    $\pl O\notktactwin{k}\gruClusGameHard{\oneptcomp\omega_1}{\infty}$.
  \end{theorem}
\end{frame}

\section{Gruenhage's Locally Finite Games}

\subsection{Definition}

\begin{frame}
  \small
  \begin{game}
  Gruenhage's locally finite games $\gruKPGame{X}$ and
  $\gruKLGame{X}$ proceed as follows:
    \begin{figure}
      \includegraphics[width=0.6\linewidth]{compactPointGame.pdf}
    \end{figure}

  $\pl K$ wins the game if the points/sets chosen by $\pl P$/$\pl L$ are
  locally finite in the space. Otherwise, $\pl P$/$\pl L$  wins.
  \end{game}
\end{frame}

% \subsection{Proximal Game}

% \begin{frame}
%   \term{Proximal Game} (2014)
%   \cite{MR3239205}

%   \vspacing

%   \begin{figure}
%     \includegraphics[width=0.8\linewidth]{proximalGame.pdf}
%   \end{figure}
%   {\tiny for compact $T_1$ $0$-dim spaces}

%   \vpause

%   The first player $\pl D$ wins the game if the points chosen by the
%   second player $\pl P$ converge. If $\pl D$ has a winning
%   strategy for this game, call $X$ \term{proximal}.
% \end{frame}

% \begin{frame}
%   Some results related to the Proximal Game due to Jocelyn Bell:

%   \begin{proposition}
%     If $X$ is metrizable, then $X$ is proximal.
%   \end{proposition}

%   \begin{theorem}
%     If $X$ is proximal, then $X$ is collectionwise normal.
%   \end{theorem}

%   \begin{theorem}
%     $\Sigma$-products and closed subspaces of proximal spaces are proximal.
%   \end{theorem}

%   \pause

%   \begin{corollary}
%     The $\Sigma$-product of metrizable spaces is collectionwise normal.
%     \cite{MR0461410} \cite{MR716576}
%   \end{corollary}
% \end{frame}

% \section{Main Result}

% \subsection{Corson compact}

% \begin{frame}
%   A \term{Corson compact} space is a space homeomorphic to a compact
%   subset of the $\Sigma$-product of real lines.

%   \vpause

%   Peter Nyikos observed:

%   \begin{proposition}
%     Every Corson compact space is proximal compact. \cite{nyikosProximalPreprint}
%   \end{proposition}

%   \pause

%   C. and Gruenhage showed in \cite{MR3227201} that any compact proximal
%   space must be Corson compact, using another game-theoretic characterization
%   of Corson compact due to Gruenhage:
% \end{frame}

% \subsection{Showing proximal compact implies Corson compact}

% \begin{frame}
%   \small
%   \term{Diagonal Game} (1984) \cite{MR752278}:

%   \begin{figure}
%     \includegraphics[width=0.7\linewidth]{diagonalGame.pdf}
%   \end{figure}
%   {\tiny for compact $T_1$ $0$-dim spaces}

%   \vpause

%   The first player $\pl O$ wins the game if any open set containing the
%   diagonal also contains infinitely many of $\pl P$'s chosen points.

%   \begin{theorem}
%     For $X$ compact, $\pl O$ has a winning strategy for the diagonal game
%     if and only if $X$ is Corson compact.
%   \end{theorem}
% \end{frame}

% \begin{frame}
%   One may use a winning strategy $\sigma$ for $\pl D$ in the proximal game to
%   construct a strategy $\tau$ for $\pl O$ in the diagonal game.

%   \begin{figure}
%     \includegraphics[width=0.7\linewidth]{convertingStrategy.pdf}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   In general:
%   \[
%     \tau(a)
%       =
%     \bigcup_{s\concat\<i,h_{s,i},j\>\in\max(T(a))}
%     \frac{1}{4}\sigma(o_s\concat\<h_{s,i}\>)[h_{s,i,j}]
%   \]

%   \vspacing

%   Using the strategy $\tau$ defined for every proximal compact space,
%   $\pl O$ cannot be defeated in the diagonal game, and therefore all
%   proximal compacts are Corson compact. \qed
% \end{frame}

% \section{Questions and References}

% \begin{frame}
%   Open questions:

%   \begin{itemize}
%     \item If compactness is dropped, does the proximal game characterize
%     all copies of \textit{closed} subspaces of a $\Sigma$-product of reals?
%     (Nyikos)
%     \item If the winning strategy for the proximal game is \term{Markov}
%     (relies on only the latest move and round number) for a compact space,
%     does that imply that the space is \term{Eberlein} compact?
%     (This holds for the diagonal game.)
%   \end{itemize}
% \end{frame}

\begin{frame}[allowframebreaks]
  \tiny
  \bibliographystyle{plain}
  \bibliography{../dissertation/bibliography}
\end{frame}

\begin{frame}
  Any questions?
\end{frame}

\end{document}