### 2015-05-28 postdoctoral/strategic_fan_tightness.tex
add strategic fan tightness result

\documentclass[11pt]{article}

\usepackage{../clontzStyle}
\usepackage{../clontzDefinitions}

\begin{document}

  \begin{definition}
    Let \(\menGame{X}\) be the Menger game played on \(X\).
    Let \(\barmanDowFanGame{X}{x}\) be the countable fan tightness
    game introduced by Barman and Dow in \cite{MR2678950}: \(\pl B\)
    chooses a set \(B_n\) with \(x\in\cl{B_n}\), followed by
    \(\pl F\) choosing \(F_n\in[B_n]^{<\omega}\), where \(\pl F\)
    wins if \(x\in\cl{\bigcup_{n<\omega} F_n}\).
  \end{definition}

  \begin{theorem}
    \(X^k\) is Menger for all \(k<\omega\) if and only if
    \(C_p(X)\) has countable fan tightness.
  \end{theorem}

  \begin{proof}
    Arhangel'ski{\u\i} in \cite{MR526012}.
  \end{proof}

  \begin{theorem}
    If \(\pl F\markwin\menGame{X}\), then
    \(\pl F\markwin\barmanDowFanGame{C_p(X)}{\vec 0}\).
  \end{theorem}

  \begin{proof}
    Essentially \cite[2.6]{MR2868880}, once one observes that
    \(\pl F\markwin\menGame{X}\) characterizes
    \(\sigma\)-relative-compactness (equivalent to \(\sigma\)-compactness
    in regular spaces). Note further that this
    property is preserved for finite powers.
  \end{proof}

  \begin{theorem}
    If \(\pl F\win\menGame{X^k}\) for all \(k<\omega\), then
    \(\pl F\win\barmanDowFanGame{C_p(X)}{\vec 0}\).
  \end{theorem}

  \begin{proof}
    For a sequence of blades
    \(s\concat\<B\>\in(\mc P(X))^{\leq\omega}\) satisfying
    \(\vec 0\in \cl{s(i)}\) for all \(i<|s|\), let
    \(t_k^m(s\concat\<B\>)\) be the sequence defined by
      \[
        t_k^m(s\concat\<B\>)(i)
          =
        \{
          (f^{-1}[(-2^{-i-m},2^{-i-m})])^k
        :
          f\in s\concat\<B\>(i)
        \}
      \]
    for each \(i\leq|s|\). In the case \(i=|s|\), note
      \[
        t_k^m(s\concat\<B\>)(|s|)
          =
        \{
          (f^{-1}[(-2^{-|s|-m},2^{-|s|-m})])^k
        :
          f\in B
        \}
      \]

    We claim that \(t_k(s\concat\<B\>)(i)\) is an open cover of \(X^k\) for
    each \(i\leq|s|\).
    Indeed, for \(\alpha\in X^k\), \(U(\vec 0,\text{ran}\alpha,2^{-i-m})\) is
    an open neighborhood of \(\vec 0\), and thus must hit some
    \(f\in s\concat\<B\>(i)\). Note then that
    \(\alpha\in(f^{-1}[(-2^{-i-m},2^{-i-m})])^k\).

    For a sequence \(s\) of length \(\leq\omega\), define its offset by
    \(n\), \(s\offset n\), as follows. Let
    \(s\offset n=\emptyset\) when \(n\geq|s|\) and otherwise
    satsify \(|s\offset n|+n=|s|\) and \(s\offset n(i)=s(n+i)\).

    If \(\sigma_k\) is the winning strategy for \(\pl F\) in \(\menGame{X^k}\),
    then let \(\tau^m\) be a strategy for \(\pl F\) in
    \(\barmanDowFanGame{C_p(X)}{\vec 0}\) satisfying
      \[
        \sigma_k(t_k^m(s\concat\<B\>))
          \subseteq
        \{
          (f^{-1}[(-2^{-|s|-m},2^{-|s|-m})])^k
        :
          f\in \tau^m(s\concat\<B\>)
        \}
      \]
    and \(\tau\) be a winning strategy satisfying
      \[
        \tau(s\concat\<B\>)
          =
        \bigcup_{m\leq|s|}
        \tau^m((s\offset m)\concat\<B\>)
      \]

    Attack \(\tau\) with \(a\). Let \(\alpha\in X^k\) and \(m<\omega\).
    Since \(t_k^m(a\offset m)\) is an attack on \(\sigma\) in
    \(\menGame{X^k}\), there is some \(n<\omega\) where
    \(\alpha\in\bigcup\sigma_k(t_k^m((a\offset m)\rest(n+1)))\),
    and therefore \(\alpha\in (f^{-1}[(-2^{-n-m},2^{-n-m})])^k\) for
    some \(f\in\tau^m((a\offset m)\rest(n+1))\), so
    \(
      f
        \in
      U(\vec 0,\text{ran}\alpha,2^{-n})
        \cap
      \tau(a\rest(n+m+1))
    \).
  \end{proof}

\newpage
\bibliographystyle{plain}
\bibliography{../bibliography}

\end{document}