### 2015-09-23 todo.tex
man these are out of date, but not gonna fix them today

\input{header}

\begin{document}
\centerline{\bf Research TODOs}

  \begin{itemize}
    \item write these TODOs
  \end{itemize}
\end{document}