### 2015-10-08 postdoctoral/S_prime_kappa.tex
incomplete kappa_n generalization

\documentclass[11pt]{article}

\usepackage{../clontzStyle}
\usepackage{../clontzDefinitions}

\begin{document}

  \begin{remark}
    Scheeper's \(S(\kappa)\) requiring injections is stronger than my
    \(S'(\kappa)\) requiring finite-to-one maps. Dow suggests that
    \(S'(\omega_\omega)\) holds in ZFC by the following.
  \end{remark}

  \begin{definition}
    A topological space is said to be \(\omega\)-bounded if each countable
    subset of the space has compact closure.
  \end{definition}

  \begin{theorem}
    For each \(k<\omega\) there exists a topology on \(\omega_k\) which
    is \(\omega\)-bounded and locally countable.
  \end{theorem}

  \begin{proof}
    Assume we've defined such a topology for \(\omega_k\)
    such that \(\gamma+1=[0,\gamma]\) is clopen for all \(\gamma<\omega_k\).
    Note that the usual linear order on \(\omega_1\) satisfies these requirements.

    Let \(\alpha<\omega_{k+1}\), and suppose we've defined compatible topologies on
    \(\omega_k\cdot (\beta+1)\) for all \(0\leq\beta<\alpha\).
    If \(\alpha=\beta+1\), then let
    \(\omega_k\cdot(\alpha+1)=\omega_k\cdot(\beta+2)\) be the topological sum
    of the previously defined \(\omega_k\cdot(\beta+1)\) and the
    previously defined
    \(\omega_k\cdot(\beta+2)\setminus\omega_k\cdot(\beta+1)\cong\omega_k\).
    Similarly, if \(cf(\alpha)>\omega\), then let \(\omega_k\cdot(\alpha+1)\)
    be the topological sum of \(\bigcup_{\beta<\alpha}\omega_k\cdot(\beta+1)\)
    and the previously defined
    \(\omega_k\cdot(\alpha+1)\setminus\omega_k\cdot\alpha\cong\omega_k\).

    The remaining case is where \(\alpha\) is the limit of increasing
    \(\alpha_n\) for \(n<\omega\). Fix a bijection
    \(
      f_\alpha
        :
      \omega_k\cdot(\alpha+1)\setminus\omega_k\cdot\alpha
        \to
      \omega_k\cdot\alpha
    \). Points in \(\omega_k\cdot(\alpha_n+1)\) for some \(n<\omega\)
    have their usual base induced by that previously defined topology. So let
    \(\gamma\in\omega_k\cdot(\alpha+1)\setminus\omega_k\cdot\alpha\).
    Basic open neighborhoods of \(\gamma\) are of the form
    \(
      W
        \cup
      f_\alpha[W]
        \setminus
      \omega_k\cdot(\alpha_n+1)
    \),
    where \(n<\omega\) and
    \(W\subseteq\gamma+1\) is any countable neighborhood of \(\gamma\).

    We wish to show that \(\omega_{k+1}\) with the topology induced by
    \(\bigcup_{\alpha<\omega_{k+1}}\omega_k\cdot(\alpha+1)\) is \(\omega\)-bounded
    and locally countable. If
    \(\gamma\in\omega_k\cdot(\alpha+1)\setminus\omega_k\cdot\alpha\) where
    \(cf(\alpha)\not=\omega\), then we immediately see that it is in a clopen
    copy of \(\omega_k\) giving us local countability
    immediately. Otherwise, \(\gamma\) has a basic open neighborhood of the form
    \(
      W
        \cup
      f_\alpha[W]
        \setminus
      \omega_k\cdot(\alpha_n+1)
    \),
    which is obviously countable.

    Let \(C\) be a countable subset of \(\omega_k\cdot(\alpha+1)\).
    In the case that \(\alpha=\beta+1\), we may use the \(\omega\)-boundedness
    of each part in the clopen partition \(\omega_k\cdot(\beta+1)\) and
    \(\omega_k\cdot(\beta+2)\setminus\omega_k\cdot(\beta+1)\cong\omega_k\)
    to conclude that the closure of \(C\) is compact.
    Similarly, if \(cf(\alpha)>\omega\),
    then we may use the  \(\omega\)-boundedness
    of each part in the clopen partition
    \(\bigcup_{\beta<\alpha}\omega_k\cdot(\beta+1)\) and
    \(\omega_k\cdot(\alpha+1)\setminus\omega_k\cdot\alpha\cong\omega_k\)
    to conclude that the closure of \(C\) is compact.

    The remaining case is again where \(\alpha\) is the limit of increasing
    \(\alpha_n\) for \(n<\omega\). Then
    TODO: generalize for \(k+1\)
    % \(
    %   C
    %     \subseteq
    %   [\gamma',\gamma]
    %     \cup
    %   f_\alpha[[\omega_k\cdot\alpha,\gamma]]
    % \)
    % for some \(\omega_k\cdot\alpha\leq\gamma'<\gamma<\omega_k\cdot(\alpha+1)\).
    % Its closure is compact:
    % the closure operation does not add any ordinals greater than \(\gamma\),
    % and any open cover
    % contains another basic open neighborhood of \(\gamma\) such as
    % \(
    %   [\gamma'',\gamma]
    %     \cup
    %   f_\alpha[[\omega_k\cdot\alpha,\gamma]]
    %     \setminus
    %   \omega_k\cdot(\alpha_m+1)
    % \)
    % which misses only the compact set \([\gamma',\gamma'']\) and the
    % closure of the countable set
    % \(
    %   f_\alpha[[\omega_k\cdot\alpha,\gamma]]
    %     \cap
    %   \omega_k\cdot(\alpha_{\min(m,n)}+1)
    % \), which is compact by the \(\omega\)-boundedness of
    % \(\omega_k\cdot(\alpha_{\min(m,n)}+1)\).

    Finally, since every countable subset of \(\omega_{k+1}\) is contained in
    some \(\omega_k\cdot(\alpha+1)\), we conclude \(\omega_{k+1}\) is
    \(\omega\)-bounded.
  \end{proof}

  \begin{theorem*}
    \(S'(\omega_\omega)\).
  \end{theorem*}

\end{document}