### 2015-10-30 postdoctoral/gamma_functions.tex
nailed the answer for \gamma

\documentclass[11pt]{article}

\usepackage{../clontzStyle}
\usepackage{../clontzDefinitions}

\renewcommand{\int}{\textrm{int}}
\renewcommand{\cl}{\textrm{cl}}
\newcommand{\lexTimes}{\times_{\textrm{lex}}}
\newcommand{\vect}{\vec}

\begin{document}

  \begin{definition}
    Let a V-map be a u.s.c. idempotent surjection.
  \end{definition}

  \begin{definition}
    For any LOS \(\<L,\leq\>\), let \(\check L\) be the collection of
    leftward subsets of \(L\)
    (subsets for which \(b\in L,a\leq b\Rightarrow a\in L\))
    linearly ordered by \(\subseteq\), and let \(\hat L\) be the collection
    of left-closed subsets of \(L\) (leftward subsets which are closed)
    linearly ordered by \(\subseteq\).
  \end{definition}

  \begin{proposition}
    \(\check L\), \(\hat L\) are compact.
  \end{proposition}

  \begin{proof}
    Each subset \(S\) has an infimum \(\cap S\) and a supremum \(\cup S\)
    (or \(\cl(\cap S)\)).
  \end{proof}

  Note that \(\check L\)
  is not a ``compactification'' as \(L\) does not necessarily
  embed as a dense subspace of \(\check L\): if \(L=I\), we might attempt to embed
  \(t\mapsto [0,t)\), but then note that the subspace topology induces the
  reverse Sorgenfrey interval as \(([0,s),[0,t])=([0,s),[0,t)]\) is open.
  However \(\hat L\) is
  the typical way of compactifying a linearly ordered space \(L\),
  provided \(L\) lacks a least element (otherwise the empty set is an [easily
  removable] isolated point in \(\hat L\)). Note that we \textbf{always}
  assume that \(\emptyset\in\hat L\):

  \begin{example}
    \(\hat I\cong\{-\infty\}\cup I\) where
    \(\emptyset\mapsto-\infty\) and \([0,t]\mapsto t\).
  \end{example}

  \begin{example}
    For limit ordinals \(\alpha\), \(\hat \alpha\cong \alpha+1\),
    and for all other infinite ordinals, \(\hat\alpha\cong\alpha\).
    (The addition of a new least isolated point is of course irrelevant).
  \end{example}

  \begin{definition}
    For any compact LOTS \(K\) with minimum \(0\) and maximum \(1\),
    let \(\gamma\) be the V-map on \(K\) where \(\gamma(0)=K\) and \(\gamma(t)=\{1\}\)
    for \(t>0\).
  \end{definition}

  \begin{definition}
    For any LOTS \(M\) with minimum element \(0\),
    let \(\nu\) be the V-map on \(M\) where \(\nu(0)=K\) and \(\nu(t)=\{t\}\)
    for \(t>0\).
  \end{definition}

  Note for \(K=M=2\) that \(\gamma=\nu\).

  \begin{theorem}
    \(X=\varprojlim \{2, \gamma, L\}\cong \check L\)
  \end{theorem}

  \begin{proof}
    We start by placing an order on \(X\). Let \(\vect x<\vect y\) if
    there exists \(a\in L\) with \(\vect x(a)=0,\vect y(a)=1\). We claim this is
    a total order inducing the topology on \(X\).

    We first observe that if \(\vect x(b)=1\), then for all \(a\leq b\),
    \(\vect x(a)\in\gamma(1)=\{1\}\). If \(\vect x\not=\vect y\), then assume
    without loss of generality that \(\vect x(a)=0,\vect y(a)=1\), so
    \(\vect x<\vect y\). Also, whenever \(\vect x(b)=1\), we have that \(b<a\),
    so \(\vect y(b)=1\), preventing \(\vect y<\vect x\). Finally if
    \(\vect x<\vect y\) and \(\vect y<\vect z\), take \(a,b\) with
    \(\vect x(a)=0\), \(\vect y(a)=1\),\(\vect y(b)=0\),\(\vect z(b)=1\). It
    follows that \(a<b\) so \(\vect z(a)=1\) and \(\vect x<\vect z\).

    Consider the basic open set \(B(\vect x,F)\) for a finite set
    \(F\in [L]^{<\omega}\)
    about the sequence \(\vect x\in X\) which contains all sequences
    \(\vect y\) agreeing with \(\vect x\) on \(F\). If \(\vect x(a)=1\) for all
    \(a\in F\), then let \(\vect w\in X\) be \(0\) on the maximum of \(F\),
    and \(1\) for anything less. It follows that
    \(B(\vect x,F)=(\vect w,\rightarrow)\). If \(\vect x(a)=0\) for all
    \(a\in F\), then let \(\vect y\in X\) be \(1\) on the minimum of \(F\),
    and \(0\) for anything greater. It follows that
    \(B(\vect x,F)=(\leftarrow,\vect y)\). Finally if \(\vect x(a)=1\) and
    \(\vect x(b)=0\) for \(a<b\) in \(F\) and nothing between \(a,b\) is in
    \(F\), then let \(\vect w\in X\) be \(0\) on \(a\)
    and \(1\) for anything less, and let \(\vect y\in X\) be \(1\) on \(b\)
    and \(0\) for anything greater. It follows that
    \(B(\vect x,F)=(\vect w,\vect y)\).

    Let \(\phi\) evaluate each \(\vect x\in X\subseteq 2^L\) as the
    characteristic function for a subset of \(L\). It's easy to see that
    \(\phi\) is an order isomorphism between \(\<X,\leq\>\) and
    \(\<\check L,\subseteq\>\).
  \end{proof}

  \begin{corollary}
    \(
      \varprojlim \{2, \gamma, \alpha\}
      \cong
      \alpha+1
    \)
    for every ordinal \(\alpha\).
  \end{corollary}

  \begin{proof}
    Since \(\check\alpha=\alpha+1\) (actually equals, not just homeomorphic!),
    we get \(\varprojlim^\star \{2, \gamma, \alpha\}
      \cong \check\alpha =
    \alpha+1\) for free.
    Note that C and Varagona used this in (TODO create citataion) to break
    metrizability in uncountable-ordinal-indexed inverse limits (for any V-map
    there exists a two-point set \(2\) such that \(f\rest 2\supseteq\gamma\),
    that is, ``\(f\) has condition \(\Gamma\)'').
  \end{proof}

  \bigskip

  We may generalize theorem 8 as follows:

  \begin{theorem}
    If \(M\) is a LOTS with minimum \(0\) and maximum \(1\),
    then \(\varprojlim\{M,\gamma,L\}\cong \hat L\lexTimes M/\sim\),
    where \(\<(\leftarrow,l_0],1\>\sim\<(\leftarrow,l_1],0\>\) if
    \(l_0<l_1\) and \((l_0,l_1)=\emptyset\),
    and where \(\<A,m\>\sim\<A,m'\>\) if \(A\in\hat L\setminus L\).
  \end{theorem}

  \begin{proof}
    Let \(\rho(\vect x)=\cl\{l\in L:\vect x(l)>0\}\), \(v(\vect 0)=0\),
    and \(v(\vect x)=\min\{\vect x(l):l\in\rho(\vect x)\}\) otherwise.
    Say \(\vect x<\vect y\) if \(\rho(\vect x)\subsetneq\rho(\vect y)\)
    or both \(\rho(\vect x)=\rho(\vect y)\) and \(v(\vect x)<v(\vect y)\).
    The reader may verify that this is a linear order on
    \(\varprojlim\{M,\gamma,L\}\), and
    \(\theta(\vect x)=\<\rho(\vect x),v(\vect x)\>\in\hat L\lexTimes M/\sim\)
    preserves order.
    For each left-closed set \(A\) and \(m\in M\),
    let \(\vect x_{A,m}(l)=1\) for \(l\in A\) unless \(l\) is the supremum
    element of \(A\),
    \(\vect x_{A,m}(l)=m\) if \(l\) is the supremum of \(A\),
    and \(\vect x_{A,m}(l)=0\) for \(l\not\in A\).
    To complete the proof, we should demonstrate that the linear order we
    defined induces the topology of the inverse limit, and that \(\theta\)
    is a surjection.

    A basic open set in \(\varprojlim\{M,\gamma,L\}\subseteq L^m\)
    is of the form
    \([U,F]\) where \(U(l)\) is an open interval in \(M\) for each
    \(l\in F\in [L]^{<\omega}\), and
    \([U,F]=\{\vect x:l\in F\Rightarrow\vect x(l)\in U(l)\}\).
    If we assume that \([U,F]\) is non-empty, one of the following
    must hold:

    \begin{itemize}
      \item
        \(U[l_0]=(a,b)\) for some \(l_0\in F\).
        Then \([U,F]=[U,\{l_0\}]\), and note that
        \(
          [U,\{l_0\}]
            =
          (\vect x_{(\leftarrow,l_0],a},\vect x_{(\leftarrow,l_0],b})
        \).
      \item
        \(U(l_0)=(a,1]\) and \(U(l_1)=[0,b)\) for some \(l_0<l_1\in L\)
        and \([U,F]=[U,\{l_0,l_1\}]\). Then
        \([U,\{l_0,l_1\}]=(\vect x_{l_0,a},\vect x_{l_1,b})\).
    \end{itemize}
    In the other direction, consider \(\vect y\in(\vect x,\vect z)\).
    \begin{itemize}
      \item
        In the case that \(l_0\in\rho(\vect y)\setminus\rho(\vect x)\)
        and \(l_1\in\rho(\vect z)\setminus\rho(\vect y)\), let
        \(U(l_0)=(0,1]\), \(U(l_1)=[0,v(\vect z))\) and note
        \(\vect y\in[U,\{l_0,l_1\}]\subseteq(\vect x,\vect z)\).
      \item
        In the case that \(l_0\in\rho(\vect y)\setminus\rho(\vect x)\),
        \(\rho(\vect y)=\rho(\vect z)\), and \(v(\vect y)<v(\vect z)\),
        it follows that \(\rho(\vect y)=\rho(\vect z)=(\leftarrow,l_1]\),
        so let \(U(l_0)=(0,1]\), \(U(l_1)=[0,v(\vect z))\) and note
        \(\vect y\in[U,\{l_0,l_1\}]\subseteq(\vect x,\vect z)\).
      \item
        In the case that
        \(\rho(\vect x)=\rho(\vect y)\), \(v(\vect x)<v(\vect y)\),
        and \(l_1\in\rho(\vect z)\setminus\rho(\vect y)\),
        it follows that \(\rho(\vect x)=\rho(\vect y)=(\leftarrow,l_0]\),
        so let \(U(l_0)=(v(\vect x),1]\), \(U(l_1)=[0,v(\vect z))\) and note
        \(\vect y\in[U,\{l_0,l_1\}]\subseteq(\vect x,\vect z)\).
      \item
        In the case that
        \(\rho(\vect x)=\rho(\vect y)=\rho(\vect z)\) and
        \(v(\vect x)<v(\vect y)<v(\vect z)\),
        it follows that
        \(\rho(\vect x)=\rho(\vect y)=\rho(\vect z)=(\leftarrow,l_0]\),
        so let \(U(l_0)=(v(\vect x),v(\vect z))\) and note
        \(\vect y\in[U,\{l_0\}]=(\vect x,\vect z)\).
    \end{itemize}

    We conclude by
    showing that \(\theta\) is a surjection. If \(B\in\hat L\setminus L\)
    and \(m\in M\), consider \(\<B,m\>\).
    \(B\) lacks a supremum in \(L\), so
    \(\vect x_{B,0}(l)=1\) for \(l\in B\) and \(\vect x_{B,0}(l)=0\) otherwise.
    So \(\theta(\vect x_{B,0})=\<\cl B,1\>=\<B,1\>\sim\<B,m\>\) for all
    \(m\in M\). Otherwise, \(B=(\leftarrow,l_1]\) for some \(l_1\in L\).
    Let \(m>0\).
    Then
    \(
      \theta(\vect x_{(\leftarrow,l_1],m})
        =
      \<\cl (\leftarrow,l_1],v(\vect x_{(\leftarrow,l_1],m})\>
        =
      \<(\leftarrow,l_1],m\>
    \).
    Finally, we want to map onto \(\<(\leftarrow,l_1],0\>\).
    If there exists \(l_0<l_1\) with \((l_0,l_1)=\emptyset\), then
    \(
      \theta(\vect x_{(\leftarrow,l_1],0})
        =
      \theta(\vect x_{(\leftarrow,l_0],1})
        =
      \<(\leftarrow,l_0],1\>
    \)
    will suffice. Otherwise,
    \(
      \theta(\vect x_{(\leftarrow,l_1],0})
        =
      \<\cl (\leftarrow,l_1),v(\vect x_{(\leftarrow,l_1),0})\>
        =
      \<(\leftarrow,l_1],0\>
    \).
  \end{proof}

  Here are some applications:

  \begin{example}
    \(
      \varprojlim\{2,\gamma,I\}
        \cong
      (\hat I\setminus\emptyset)\lexTimes 2
        \cong
      I\lexTimes 2
        \cong
      \check I
    \) (of course, this could be found quicker with theorem 8).
  \end{example}

  \begin{example}
    \(
      \varprojlim\{I,\gamma,I\}
        \cong
      (\hat I\setminus\emptyset)\lexTimes I
        \cong
      I\lexTimes I
    \).
  \end{example}

  \begin{example}
    For infinite ordinals \(\alpha\),
    \(
      \varprojlim\{I,\gamma,\alpha\}
        \cong
      (\alpha\lexTimes[0,1))\cup\{\infty\}
    \).
    In particular, \(\alpha=\kappa\) for an infinite cardinal \(\kappa\)
    gives the closed long ray of length \(\kappa\).
  \end{example}

  \bigskip

  We introduce an alternate definition of an arbitrarily indexed
  inverse limit.

  \begin{definition}
    Let \(\varprojlim^\star\{X,f,L\}\subseteq\varprojlim\{X,f,L\}\) satisfy
    that \(\vect x(a)=\lim_{t\to a}\vect x(t)\) for all \(a\in L\)
    (for any open neighborhood
    \(U\)of \(\vect x(a)\) there is \(b<a\) where \(\vect x(t)\in U\)
    for all \(t\in(b,a]\)).
  \end{definition}

  \begin{theorem}
    \(Y=\varprojlim^\star \{2, \gamma, L\}\cong \hat L\).
  \end{theorem}

  \begin{proof}
    Consider \(Y\) as a subspace of \(X=\varprojlim \{2, \gamma, L\}\) with
    the linear order described above. We claim that if \(\phi\) is the
    characteristic function for a subset of \(L\), then \(\phi\)
    is an order isomorphsim between \(\<Y,\leq\>\) and
    \(\<\hat L,\subseteq\>\).

    Let \(A\) be a left-closed subset of \(L\). Let \(\vect x(a)=1\) when
    \(a\in A\) and \(\vect x(a)=0\) otherwise. Then \(\vect x\in Y\) and
    \(\phi(\vect x)=A\).

    Let \(\vect x,\vect y\in Y\). If
    \(\phi(\vect x)=\phi(\vect y)=A\), then \(A\) is a
    left-closed set where \(\vect x(a)=\vect y(a)=1\) for \(a\in A\)
    and \(\vect x(a)=\vect y(a)=0\) otherwise, so \(\vect x=\vect y\).

    Finally let \(\vect x<\vect y\), so there exists \(a\in L\) with
    \(\vect x(a)=0\), \(\vect y(a)=1\). Then
    \(
      \phi(\vect x)
        \subseteq
      (\leftarrow,a)
        \subseteq
      \phi(\vect y)
    \). Thus \(\phi\) preserves order.
  \end{proof}

  \begin{corollary}
    \(
      \varprojlim^\star \{2, \gamma, \alpha\}
      \cong
      \alpha+1
    \)
    for every infinite limit or finite ordinal \(\alpha\).
  \end{corollary}

  \begin{proof}
    If \(\alpha\) is finite, then of course all (leftward) sets are
    closed and we get \(\hat\alpha=\check\alpha=\alpha+1\) for free.
    Otherwise, as observed previously \(\hat\alpha\)
    is homeomorphic to its usual compactification \(\alpha+1\) for
    limit ordinals.
  \end{proof}

  In fact, \(\hat\alpha=\alpha+1\setminus L(\alpha)\) where \(L(\alpha)\)
  is the colleciton of all limit ordinals less than \(\alpha\), which also
  shows \(\hat\alpha\cong\alpha\) for infinite successor ordinals \(\alpha\).

\newpage
\bibliographystyle{plain}
\bibliography{../bibliography}

\end{document}