### 2015-11-09 postdoctoral/S_prime_kappa.tex
start writing up notes based on alan's input

\documentclass[11pt]{article}

\usepackage{../clontzStyle}
\usepackage{../clontzDefinitions}

\begin{document}

  \begin{remark}
    Scheeper's \(S(\kappa)\) requiring injections is stronger than my
    \(S'(\kappa)\) requiring finite-to-one maps. Dow suggests that
    \(S'(\omega_\omega)\) holds in ZFC by the following.
  \end{remark}

  \begin{definition}
    A topological space is said to be \(\omega\)-bounded if each countable
    subset of the space has compact closure.
  \end{definition}

  \begin{theorem}
    For each \(k<\omega\) there exists a topology on \(\omega_k\) which
    is \(\omega\)-bounded and locally countable.
  \end{theorem}

  \begin{proof}
    TODO: write up Alan's proof
  \end{proof}

  \begin{definition}
    A Kurepa family \(\mc K\)
    is a family of countable sets such that for each \(A\in\mc K\),
    \(\mc K\rest A=\{K\cap A:K\in\mc K\}\) is countable.
  \end{definition}

  \begin{corollary}
    There exists a Kurepa family cofinal in \([\omega_k]^\omega\)
    for each \(k<\omega\).
  \end{corollary}

  \begin{proof}
    This is actually a corollary of an observation of Todorcevic communicated
    by Dow in [TODO cite Gen Prog in Top I]:
    if every Kurepa family of size at most \(\theta\)
    extends to a cofinal Kurepa family, then the same is true of \(\theta^+\).
    So the result follows as
    every Kurepa family of size \(\omega\) extends to
    the cofinal Kurepa family \([\omega]^\omega\). (TODO am I understanding
    this right?)

    We may alternatively
    obtain the result from the previous topological argument by using the family
    \(\mc K\) of compact sets in the constructed topology on
    \(\omega_k\) as our witness. Of course, every Lindel\"of set in
    a locally countable space is countable. Thus \(\mc K\)
    is cofinal since for every countable set \(A\), \(\cl A\) is compact
    and countable.
    It is Kurepa since for every compact (countable) set \(A\), (TODO).
  \end{proof}

  \begin{theorem}
    \(S'(\omega_k)\) for each \(k<\omega\).
  \end{theorem}

  \begin{proof}
    Let \(k<\omega\), and \(\mc K\) be a cofinal Kurepa family on \(\omega_k\).
    (TODO finish)
  \end{proof}

\end{document}