### 2015-12-10 articles/almost-compatible-functions-and-topological-games/article.tex
begin cleaning omega_2 cohen reals argument, including deflation of math mode

\documentclass{amsart}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}

\usepackage{tikz}
\usetikzlibrary{matrix}

\usepackage{../../clontzDefinitions}

      \theoremstyle{plain}
      \newtheorem{theorem}{Theorem}
      \newtheorem{lemma}[theorem]{Lemma}
      \newtheorem{corollary}[theorem]{Corollary}
      \newtheorem{proposition}[theorem]{Proposition}
      \newtheorem{conjecture}[theorem]{Conjecture}
      \newtheorem{question}[theorem]{Question}
      \newtheorem{claim}[theorem]{Claim}

      \theoremstyle{definition}
      \newtheorem{definition}[theorem]{Definition}
      \newtheorem{example}[theorem]{Example}
      \newtheorem{game}[theorem]{Game}

      \theoremstyle{remark}
      \newtheorem{remark}[theorem]{Remark}

      \theoremstyle{plain}
      \newtheorem*{theorem*}{Theorem}
      \newtheorem*{lemma*}{Lemma}
      \newtheorem*{corollary*}{Corollary}
      \newtheorem*{proposition*}{Proposition}
      \newtheorem*{conjecture*}{Conjecture}
      \newtheorem*{question*}{Question}
      \newtheorem*{claim*}{Claim}

      \theoremstyle{definition}
      \newtheorem*{definition*}{Definition}
      \newtheorem*{example*}{Example}
      \newtheorem*{game*}{Game}

      \theoremstyle{remark}
      \newtheorem*{remark*}{Remark}

\parskip=0.5em

% dow stuff
          \def\Succ{\mathop{Succ}}
          \def\dom{\mathop{dom}}
          \def\val{\operatorname{val}}
          \def\N{\mathbb N}
          \def\Nstar{\mathbb N^*}

\begin{document}

% \title{Proximal compact spaces are Corson compact\tnoteref{t1}}
% \tnotetext[t1]{2010 Mathematics Subject Classification. 54E15, 54D30, 54A20.}
\title{Almost compatible functions and infinite length games}

% \author[aub]{S.~Clontz\fnref{fn1}}
% \ead{steven.clontz@auburn.edu}
% \author[aub]{G.~Gruenhage\fnref{fn2}}
% \ead{gruengf@auburn.edu}

% \address[aub]{Department of Mathematics, Auburn University,
%  Auburn, AL 36830}

\author{Steven Clontz}
\address{Department of Mathematics and Statistics, UNC Charlotte,
Charlotte, NC 28262}
\email{steven.clontz@gmail.com}
\author{Alan Dow}
\address{Department of Mathematics and Statistics, UNC Charlotte,
Charlotte, NC 28262}
\email{adow@uncc.edu}

% \keywords{TODO}
% \subjclass[2010]{}

\begin{abstract}
  \(\alcompSP\kappa\) asserts the existence of pairwise
  almost compatible finite-to-one functions \(A\to\omega\) for each
  countable subset \(A\) of \(\kappa\).
  The existence of winning \(2\)-Markov strategies in
  several infinite-length games, including
  the Menger game on the one-point Lindel\"ofication \(\oneptlind\kappa\)
  of \(\kappa\), are guaranteed by \(\alcompSP\kappa\). \(\alcompSP\kappa\) is
  implied by the existence of cofinal Kurepa families of size \(\kappa\),
  and thus holds for all cardinals less than \(\aleph_\omega\).
  It's consistent that \(\alcompSP{\aleph_\omega}\) fails, but
  there must always be a winning \(2\)-Markov strategy for the second
  player in the Menger game on \(\oneptlind{\omega_\omega}\).
\end{abstract}

\maketitle

\section{Introduction}

  \begin{definition}
    Two functions \(f,g\) are almost compatible, that is, \(f\sim g\) when
    \(\{a\in\dom f\cap\dom g:f(a)\not=g(a)\}\) is finite.
  \end{definition}

  Marion Scheepers used almost compatible functions in \cite{MR1129143} in order
  to study the existence of limited information strategies on a variation
  of the meager-nowhere dense game he introduced in \cite{MR1183703}.

  \begin{game}
    Let \(\schFillStrictGame\kappa\) denote
    \term{Scheepers' strict countable-finite union game}
    with two players \(\pl C\), \(\pl F\). In round \(0\), \(\pl C\) chooses
    \(C_0\in[\kappa]^{\leq\omega}\), followed by \(\pl F\) choosing
    \(F_0\in[\kappa]^{<\omega}\). In round \(n+1\), \(\pl C\) chooses
    \(C_{n+1}\in[\kappa]^{\leq\omega}\) such that \(C_{n+1}\supset C_n\), followed
    by \(\pl F\) choosing \(F_{n+1}\in[\kappa]^{<\omega}\).

    \(\pl F\) wins the game if
    \(\bigcup_{n<\omega} F_n\supseteq\bigcup_{n<\omega} C_n\); otherwise,
    \(\pl C\) wins.
  \end{game}

  Of course, with perfect information this game is trivial: during round
  \(n\) player \(\pl F\) simply chooses \(n\) ordinals from each of the
  \(n\) countable sets played by \(\pl C\). However, if \(\pl F\) is limited
  to using information from the last \(k\) moves by \(\pl C\) during each
  round, the task becomes more difficult. Call such a strategy a
  \term{\(k\)-tactical strategy} or \term{\(k\)-tactic}; if using the round
  number is allowed, then the strategy is called a \term{\(k\)-Markov strategy}
  or a \term{\(k\)-mark}.

  \begin{definition}
    The statement \(\alcompS\kappa\) (given as \(S(\kappa,\aleph_0,\omega)\)
    in \cite{MR1129143} and \(S(\kappa)\) in \cite{clontzMengerGamePreprint})
    claims that there exist one-to-one functions
    \(f_A:A\to\omega\) for each \(A\in[\kappa]^{\leq\aleph_0}\) such that
    the collection \(\{f_A:A\in[\kappa]^{\leq\aleph_0}\}\) is pairwise
    almost compatible.
  \end{definition}

  In the same paper, Scheepers noted that \(\alcompS{\omega_1}\) holds in \(ZFC\),
  and that it's possible to force \(\mf c\) to be arbitrarily large
  while preserving \(\alcompS{\mf c}\). However, \(\alcompS{\mf c^+}\) always fails.
  This axiom may be applied to obtain a winning \(2\)-tactic for \(\pl F\)
  in the countable-finite game.

  In \cite{clontzMengerGamePreprint}, Clontz related this game to a game
  which may be used to characterize the Menger covering property of a
  topological space.

  \begin{game}
    Let \(\menGame{X}\) denote the \term{Menger game} with players \(\pl C\),
    \(\pl F\).
    In round \(n\), \(\pl C\) chooses an open cover \(\mc U_n\), followed by
    \(\pl F\)
    choosing subset \(F_n\) of \(X\) which may be finitely covered by \(\mc U_n\).

    \(\pl F\) wins the game if \(X = \bigcup_{n<\omega}F_n\),
    and \(\pl C\) wins otherwise.
  \end{game}

  This characterization is slightly different than the typical characterization
  in which the second player first chooses a specific
  finite subcollection \(\mc F_n\) of the cover itself and lets
  \(F_n=\bigcup \mc F_n\), denoted as \(G_{fin}(\mc O,\mc O)\)
  in \cite{MR1378387}.
  However, it's easily seen that these games are equivalent
  for perfect information strategies (so both characterize
  the Menger property), and this characterization is
  more convenient for our concerns.

  \begin{definition}
    Let \(\oneptlind\kappa=\kappa\cup\{\infty\}\) where \(\kappa\) is
    discrete and \(\infty\)'s neighborhoods are the co-countable sets
    containing it.
  \end{definition}

  The relationship between \(\schFillStrictGame\kappa\) and
  \(\menGame{\oneptlind\kappa}\) is strong; in both games \(\pl C\) essentially
  chooses a countable subset of \(\kappa\) followed by \(\pl F\) choosing
  a finite subset of that choice, and it's easy to see the winning perfect
  information strategy for \(\pl F\) in both games.
  In addition, it was shown in
  \cite{clontzMengerGamePreprint} that when \(\alcompS{\kappa}\) holds,
  \(\pl F\) has a winning \(2\)-Markov strategy in
  \(\menGame{\oneptlind\kappa}\).

  One source of motivation is to make progress on the following open question:

  \begin{question}\label{mainQuestion}
    Does there exist a topological space \(X\) for which
    \(\pl F\win\menGame{X}\) but \(\pl F\notkmarkwin{2}\menGame{X}\)?
    (That is, the second player can win the Menger game on \(X\)
    with perfect information but not with \(2\)-Markov information.)
  \end{question}

  \section{One-to-one and finite-to-one almost compatible functions}

  We may weaken Scheeper's \(\alcompS{\kappa}\) as follows:

  \begin{definition}
    The statement \(\alcompSP{\kappa}\) weakens \(\alcompS{\kappa}\) by only
    requiring the witnessing almost-compatible functions \(f_A:A\to\omega\)
    to be finite-to-one.
  \end{definition}

  \begin{proposition}
    \(\alcompS{\kappa}\) and \(\alcompSP{\kappa}\) need only be witnessed by
    functions \(\{f_A:A\in\mc S\}\) for some family \(\mc S\)
    cofinal in \([\kappa]^{\leq\aleph_0}\).
  \end{proposition}

  \begin{proof}
    For each \(A\in[\kappa]^{\leq\aleph_0}\) choose \(A'\supseteq A\)
    from \(\mc S\) and let \(g_A=f_{A'}\rest A\).
  \end{proof}

  In the next section we will show that \(\alcompSP{\kappa}\) is sufficient
  for the applications to the Scheepers and Menger games.
  In the meantime, we will demonstrate that \(\alcompSP{\kappa}\) is strictly
  weaker than \(\alcompS{\kappa}\).

  Recall the following.

  \begin{definition}
    A Kurepa family \(\mc K\subseteq[\kappa]^{\aleph_0}\) on \(\kappa\)
    satisfies that
    \(\mc K\rest A=\{K\cap A:K\in\mc K\}\) is countable
    for each \(A\in[\kappa]^{\aleph_0}\).
    Let \(\kurepaS\kappa\) be the statement claiming there exists
    a Kurepa family on \(\kappa\)
    cofinal in \([\kappa]^{\aleph_0}\).
  \end{definition}

  \begin{theorem}
    \(\kurepaS\kappa \Rightarrow \alcompSP{\kappa}\).
  \end{theorem}

  \begin{proof}
    Let \(\mc K=\{K_\alpha:\alpha<\theta\}\)
    be a cofinal Kurepa family on \(\kappa\).
    We first define \(f_\alpha:K_\alpha\to\omega\) for each \(\alpha<\theta\).

    Suppose we've already defined pairwise almost compatible finite-to-one functions
    \(\{f_\beta:\beta<\alpha\}\). To define
    \(f_\alpha\), we first recall that \(\mc K\rest K_\alpha\) is countable,
    so we may choose \(\beta_n<\alpha\) for \(n<\omega\) such that
    \(
      \{K_\beta:\beta<\alpha\}\rest K_\alpha \setminus \{\emptyset\}
        =
      \{K_\alpha\cap K_{\beta_n}:n<\omega\}
    \).
    Let \(K_\alpha=\{\delta_{i,j}:i\leq\omega,j<w_i\}\) where
    \(w_i\leq\omega\) for each \(i\leq\omega\),
    \(
      K_\alpha\cap \left(K_{\beta_n}\setminus\bigcup_{m<n}K_{\beta_m}\right)
        =
      \{\delta_{n,j}:j<w_n\}
    \),
    and
    \(
      K_\alpha\setminus\bigcup_{n<\omega}K_{\beta_n}
        =
      \{
        \delta_{\omega,j}:j<w_\omega
      \}
    \).
    Then let \(f_\alpha(\delta_{n,j})=\max(n,f_{\beta_n}(\delta_{n,j}))\) for
    \(n<\omega\) and \(f_\alpha(\delta_{\omega,j})=j\) otherwise.

    We should show that \(f_\alpha\) is finite-to-one. Let \(n<\omega\).
    Since \(f_\alpha(\delta_{m,j})\geq m\), we only consider the finite cases
    where \(m\leq n\). Since each
    \(f_{\beta_m}\) is finite-to-one, \(f_{\beta_m}(\delta_{m,j})\leq n\)
    for only finitely many \(j\). Thus
    \(f_\alpha(\delta_{m,j})=\max(m,f_{\beta_m}(\delta_{m,j}))\) maps to
    \(n\) for only finitely many \(j\).

    We now want to demonstrate that \(f_\alpha\sim f_{\beta_n}\) for all
    \(n<\omega\). Note \(\delta_{m,j}\in K_{\beta_n}\) implies
    \(m\leq n\).
    For \(m=n\), we have
    \(f_\alpha(\delta_{n,j})=\max(n,f_{\beta_n}(\delta_{n,j}))\) which differs
    from \(f_{\beta_n}(\delta_{n,j})\) for only the finitely many \(j\) which
    are mapped below \(n\) by \(f_{\beta_n}\).
    For \(m<n\) and \(\delta_{m,j}\in K_{\beta_n}\),we have
    \(f_\alpha(\delta_{m,j})=\max(m,f_{\beta_m}(\delta_{m,j}))\) which can
    only differ
    from \(f_{\beta_n}(\delta_{m,j})\) for only the finitely many \(j\) which
    are mapped below \(m\) by \(f_{\beta_m}\) or the finitely many \(j\)
    for which the
    almost compatible \(f_{\beta_n}\sim f_{\beta_m}\) differ.

    Finally for any \(\beta<\alpha\), we may conclude \(f_\alpha\sim f_\beta\)
    since there is some \(\beta_n\) with
    \(K_\alpha\cap K_\beta=K_\alpha\cap K_{\beta_n}\),
    \(f_\alpha\sim f_{\beta_n}\), and \(f_{\beta_n}\sim f_\beta\).
  \end{proof}

  We now construct a topology on \(\omega_n\) for each \(n<\omega\) which
  will witness a Kurepa family of size \(\aleph_n\). A similar construction
  was previously shown by Juh\'asz et. al. in \cite{MR539228},
  and the relationship of
  Kurepa families and such spaces has also been investigated in a preprint
  of Nyikos \cite{NYIKOSKUREPA}.

  \begin{proposition}
    Let \(X\) be a \(T_2\) space with a base of countable and compact
    neighborhoods. Then \(X\) is locally metrizable with
    a base of compact open countable sets.
  \end{proposition}

  \begin{proof}
    For each point \(x\) let \(K\) be a countable and compact neighborhood
    of \(x\), and it follows that it is contained in a countable, open,
    and locally compact neighborhood \(W\) of \(x\), which in turn
    is zero-dimensional and metrizable. So choose \(V\) clopen
    in \(W\) such that \(x\in V\subseteq K\); \(V\) is a compact open
    neighborhood of \(x\) in \(X\).
  \end{proof}

  \begin{definition}
    A topological space is said to be \term{\(\omega\)-bounded}
    if each countable subset of the space has compact closure.
    As in \cite{MR539228} we call a
    \(T_2\), locally countable, \(\omega\)-bounded space
    \term{splendid}, and let \(\splendidS\kappa\) represent
    the claim that there exists a splendid space of cardinality \(\kappa\).
  \end{definition}

  \begin{proposition}
    Let \(X\) be a \(T_2\) space with cardinality less than \(\aleph_\omega\)
    which is locally countable and
    \(\omega\)-bounded. Then the closure operation preserves cardinality and
    weight.
  \end{proposition}

  \begin{proof}
    Note that the closure of any countable neighborhood is compact, and
    any Lindel\"of set is countable. This space is locally metrizable
    and thus first-countable, so cardinality and weight coincide for any
    subspace. The result is obvious if \(A\) is countable; otherwise
    let \(A=\{a_\alpha:\alpha<\omega_{n+1}\}\) and since basic neighborhoods
    are countable note any limit point of \(A\) is
    a limit point of \(A_\beta=\{a_\alpha:\alpha<\beta\}\)
    for some \(\beta<\omega_{n+1}\).
    Thus \(\cl A=\bigcup_{\beta<\omega_{n+1}}\cl{A_\beta}\) and
    by induction \(|\cl A|=|A|\).
  \end{proof}

  \begin{lemma}
    Let \(X\) be a \(T_2\) space with cardinality less than \(\aleph_\omega\)
    which is locally countable and locally compact, and such that
    its closure operation preserves cardinalities.
    Then \(X\) has an \(\omega\)-bounded extension \(\tilde X\)
    with the same properties where \(\tilde X\setminus X\) has the same
    cardinality as \(X\).
  \end{lemma}

  \begin{proof}
    We prove this by induction on \(n\).
    If \(n=0\), then we can just use the one-point compactification of two
    copies of \(X\). So suppose \(n>0\) and
    that \(X=\omega_n\) has an appropriate topology. Note that \(X\) has
    a base of countable and compact neighborhoods since the closure operation
    preserves cardinalities.

    For each \(\alpha<\omega_n\), \(\gamma_\alpha\) may be chosen such that
    both the closure of the set \(\alpha\) in \(X\) and a countable
    neighborhood of the point \(\alpha\) are subsets of \(\gamma_\alpha\).
    Note that the set
    \(\{\lambda<\omega_n:\alpha<\lambda\Rightarrow\gamma_\alpha<\lambda\}\)
    is a cub subset of \(\omega_n\) containing a cub subset \(C\)
    of limit ordinals.
    Now for each \(\lambda\in C\), the set \(\lambda\) is
    open as \(\alpha<\lambda\) belongs to the neighborhood
    \(\gamma_\alpha\subsetneq\lambda\). Also, if \(\lambda\) has uncountable
    cofinality, then for \(\beta\geq\lambda\) and any countable neighborhood
    \(U\) of \(\beta\), \(U\cap\lambda=U\cap\alpha\)
    for some \(\alpha<\lambda\); thus \(U\setminus\cl\alpha=U\setminus\lambda\)
    is a neighborhood of \(\beta\), showing that \(\lambda\)
    is clopen.

    Let \(\tilde X=\omega_n\times 2\).
    By induction on \(\lambda \in C\) we will define compatible topologies
    for \(\tilde X_\lambda=\omega_n\times\{0\}\cup\lambda\times\{1\}\)
    such that
    \begin{itemize}
      \item \(\omega_n \times \{0\}\) is an open copy of \(X\),
      \item \(\lambda\times 2\) is open, and when \(\cf\lambda>\omega\) also closed,
      \item the space has a base of countable and compact neighborhoods, and
      \item when \(\lambda\) is a successor,
        for each \(\alpha<\lambda\) the closure of \(\alpha\times 2\) is
        an \(\omega\)-bounded  subset of \(\lambda\times 2\).
    \end{itemize}

    We first consider the case \(n=1\).
    If \(\lambda\) is a limit in \(C\), then
    \(
      \tilde X_\lambda
        =
      \bigcup_{\mu\in C\cap\lambda}\tilde X_\mu
    \)
    satisfies the induction requirements.
    Otherwise we choose an increasing sequence of
    ordinals \(\{\alpha_k : k\in \omega\}\) with limit \(\lambda\)
    such that \(\alpha_0\) is the predecessor of \(\lambda\) in \(C\),
    or \(\alpha_0=0\) if \(\lambda\) is the least element of \(C\).

    The subspace \(\cl\lambda\times\{0\}\cup\alpha_0\times2\) of \(X\)
    is countable and locally compact; therefore it is
    metrizable and zero-dimensional.
    So we may choose increasing sets \(U_k\) for \(k<\omega\) which are
    clopen in this topology and satisfy
    \[
      \cl{\alpha_k\times\{0\}\cup\alpha_0\times2}
        =
      \cl{\alpha_k}\times\{0\}\cup\alpha_0\times2
        \subseteq
      U_k
        \subseteq
      \lambda\times\{0\}\cup\alpha_0\times 2
    \]
    Note that \(U_k\) is also clopen in \(\tilde X_{\alpha_0}\) since it is
    closed in \(\cl\lambda\times\{0\}\cup\alpha_0\times2\) and open in
    \(\lambda\times\{0\}\cup\alpha_0\times 2\).

    We need only describe a base for the points
    \(\<\alpha,1\>\in(\lambda\setminus\alpha_0)\times\{1\}\).
    We do so by letting
    \(\<\alpha,1\>\) be isolated when \(\alpha\not\in\{\alpha_k:k<\omega\}\),
    and giving \(\<\alpha_k,1\>\) the open neighborhoods
    \((U_k\cup((\alpha_k+1)\times\{1\}))\setminus K\) for each compact
    subset \(K\) of \(U_k\cup(\alpha_k\times\{1\})\); that is,
    \(\<\alpha_k,1\>\) is the one point compactifying
    \(U_k\cup(\alpha_k\times\{1\})\).

    The first two requirements of our inductive hypothesis are obviously
    satsified. Note points in \(\lambda\times 2\) are covered by
    the compact countable neighborhood \(U_k\cup((\alpha_k+1)\times\{1\})\)
    for some \(k<\omega\), and for points in
    \((\omega_n\setminus\lambda)\times\{0\}\) we may use a compact
    countable neighborhood from \(X\).
    For the final requirement, note that for \(\alpha<\lambda\), we may
    choose \(\alpha<\alpha_k<\lambda\) and note that
    \(\alpha\times 2\) is contained in the compact
    subset \(U_k\cup((\alpha_k+1)\times\{1\})\) of \(\lambda\times2\).

    For the case \(n>1\), we may assume that the successors in \(C\)
    have uncountable cofinality.
    We again proceed by induction on \(\lambda\in C\).  Again when
    \(\lambda\) is a limit in \(C\),
    \(
      \tilde X_\lambda
        =
      \bigcup_{\mu\in C\cap\lambda}\tilde X_\mu
    \)
    satisfies the given
    requirements; in particular if \(\alpha<\lambda\), then
    \(\alpha<\mu<\lambda\) for some successor \(\mu\in C\) with uncountable
    cofinality. As such, the closure of \(\alpha\times2\) is
    an \(\omega\)-bounded subset of the clopen \(\mu\times 2\) and
    therefore also of \(\lambda\times 2\).
    In case \(\lambda\) is not a limit of \(C\), then \(\lambda\)
    has uncountable cofinality and a predecessor \(\mu\in C\).
    We therefore have that \(\lambda\times\{0\}\) is clopen
    in \(\omega_n\times \{0\}\). Since the cardinality of
    \(\lambda\times \{0\} \cup \mu\times 2\) is less than \(\aleph_n\),
    we may simply apply the induction hypothesis
    to choose an appropriate topology for \(\lambda\times 2\).

    As a result, \(\tilde X=\bigcup_{\lambda\in C}\tilde X_\lambda\)
    is
    \(\omega\)-bounded as any countable set is contained in some
    \(\alpha\times 2\) for \(\alpha<\lambda\in C\).
  \end{proof}

  \begin{theorem}
    For each \(k< \omega\), there is a \(T_2\), locally countable,
    \(\omega\)-bounded topology on \(\omega_k\). That is, \(\splendidS{\aleph_k}\)
    for all \(k<\omega\).
  \end{theorem}

  \begin{proof}
    Apply the previous lemma to \(\omega_n\) with the discrete topology.
  \end{proof}

  \begin{lemma}
    The family of compact open sets in a locally countable, \(\omega\)-bounded
    topological space \(X\)
    is a Kurepa family cofinal in \([X]^\omega\). That is,
    \(\splendidS\kappa \Rightarrow \kurepaS\kappa\).
  \end{lemma}

  \begin{proof}
    Let \(\mc K\) collect all compact
    open subsets of \(X\). Of course, every Lindel\"of set in
    a locally countable space is countable, and the closure of every
    countable set is a compact countable set; thus \(\mc K\)
    is cofinal in \([X]^\omega\).
    It is Kurepa since every countable set is contained in a countable
    compact open subspace of \(X\); this subspace has a countable
    base of compact open sets, which closed under finite unions enumerates
    all compact open subsets of the subspace.
  \end{proof}

  \begin{corollary}
    \(\kurepaS{\aleph_k}\) for all \(k<\omega\).
  \end{corollary}

  Alternatively, the previous corollary may be obtained
  via an observation of Todorcevic
  communicated by Dow in \cite{MR1229125}:
  if every Kurepa family of size at most \(\kappa\)
  extends to a cofinal Kurepa family, then the same is true of \(\kappa^+\).

  Nyikos points out in \cite{NYIKOSKUREPA} that a cofinal Kurepa family may
  be used to construct a locally metrizable, \(\omega\)-bounded,
  zero-dimensional space with appropriate cardinality,
  but whether this can be
  strengthened to locally countable and \(\omega\)-bounded (as asked in
  \cite{MR539228}) remains an open question.

  Also left open is this extension of the question asked in \cite{NYIKOSKUREPA}
  and \cite{MR539228} on the possible equivalence of \(\splendidS\kappa\)
  and \(\kurepaS\kappa\).

  \begin{question}
    May any of the implications in the theorem
    \(\splendidS\kappa\Rightarrow\kurepaS\kappa\Rightarrow\alcompSP\kappa\)
    be reversed?
  \end{question}

  Regardless, we have obtained our desired result.

  \begin{corollary}
    \(\alcompSP{\aleph_k}\) for all \(k<\omega\).
  \end{corollary}

  \section{Consistency results}

  As noted in \cite{MR1229125}, Jensen's one-gap two-cardinal theorem under
  \(V=L\) introduced in \cite{MR0309729} implies that \(\kurepaS\kappa\)
  holds for all cardinals \(\kappa\).

  \begin{corollary}[\(V=L\)]
    \(\alcompSP{\kappa}\) for all cardinals \(\kappa\).
  \end{corollary}

  Weakening to the continuum hypothesis, we see an obvious consequence.

  \begin{corollary}[\(CH\)]
    \(\alcompSP{\mf c^+}\), but \(\neg\alcompS{\mf c^+}\).
  \end{corollary}

  But \(CH\) is not required to have \(\alcompS{\aleph_2}\) fail.
  We take the following lemma from Kunen.

  \begin{lemma}[\cite{MR597342}]
    Assume that $G\subset \operatorname{Fn}(\omega_2,2)$ is a generic filter,
    and let $\mu\in \omega_2$. Then the final model $V[G]$ can be
    regarded as forcing  with $\operatorname{Fn}(\omega_2\setminus \mu,
    2)$ over the model $V[G_\mu]$.
    In addition, for each $\operatorname{Fn}(\omega_2,2)$-name $\dot A$
    for $A\subseteq\omega$ (abusing notation and letting \(n\) represent
    its own $\operatorname{Fn}(\omega_2,2)$-name and
    $\operatorname{Fn}(\omega_2\setminus \mu, 2)$-name),
    there is a canonical name $\dot A(G_\mu)$ where,
    $$\dot A(G_\mu) = \{ (n,p\restriction [\mu,\omega_2))  :
    (n,p)\in \dot A\ \ \mbox{and} \ \ p\restriction \mu\in G_\mu\}$$
    and the valuation of $\dot A(G_\mu)$ by the tail
    of the generic, $G_{\omega_2\setminus \mu}$, is the same as
    the valuation of $\dot A$ by the full generic.
  \end{lemma}

 \begin{lemma}[\cite{MR597342}]
    Assume that \(G\subset \operatorname{Fn}(\omega_2,2)\) is a generic filter,
    and let \(\mu\in \omega_2\). Then the final model \(V[G]\) can be
    regarded as forcing  with \(\operatorname{Fn}(\omega_2\setminus \mu,
    2)\) over the model \(V[G_\mu]\).
    In addition, for each \(\operatorname{Fn}(\omega_2,2)\)-name \(\dot A\)
    for \(A\subseteq\omega\) (abusing notation and letting \(n\) represent
    its own \(\operatorname{Fn}(\omega_2,2)\)-name and
    \(\operatorname{Fn}(\omega_2\setminus \mu, 2)\)-name),
    there is a canonical name \(\dot A(G_\mu)\) where,
    \[\dot A(G_\mu) = \{ (n,p\restriction [\mu,\omega_2))  :
    (n,p)\in \dot A\ \ \mbox{and} \ \ p\restriction \mu\in G_\mu\}\]
    and the valuation of \(\dot A(G_\mu)\) by the tail
    of the generic, \(G_{\omega_2\setminus \mu}\), is the same as
    the valuation of \(\dot A\) by the full generic.
  \end{lemma}

  \begin{theorem}
    It follows from the consistency of \(ZFC\) that
    there exists a model of \(ZFC\) for which \(\mf c=\aleph_2\)
    and \(\neg\alcompS{\aleph_2}\).
  \end{theorem}

  \begin{proof}
    The forcing poset is \(\operatorname{Fn}(\omega_2, 2)\).
    Let \(\{ \dot f_{A} : A\in [\omega_2]^\omega\}\) be a family of
    names such that \(f_{A}\) is a one-to-one function from \(A\) into
    \(\omega\) in the ground model. Let \(M\) be an elementary submodel
    of the universe such that
    \(|M|=\omega_1 =\mathfrak c\) and \(M^\omega\subset M\).

    Let \(\lambda = M\cap \omega_2=\sup(M\cap\omega_2)\).
    Consider the name
    \[
      \dot f_{[\lambda,\lambda+\omega)}
        =
      \{
        ( (\lambda+k,m), p)
      :
        p\in \mathop{Fn}(\omega_2,2);k,m\in \omega
      \}
    \]
    This is (almost) equivalent to
             saying
             that \(p\) forces that \(\dot f_{[\lambda,\lambda+\omega)}(\lambda+k) =
             m\). We don't take all such \(p\), in fact for each \(k,m\) it is enough
            to take a
             countable set of such \(p\) to get an equivalent name
             (Kunen calls it a nice name if we take, for each \(k,m\) an
            antichain that is maximal among such conditions).
            Given any such name \(\dot f\), let \(\operatorname{supp}(\dot f)\)
            denote the union of the domains of conditions \(p\) appearing in the
            name.

             Also let \(Y\) equal \(\operatorname{supp}(\dot
             f_{[\lambda,\lambda+\omega)})\setminus \lambda\).
             Let \(\delta\) denote the order type of
             \(Y\) and let the 2-parameter notation
             \(\varphi_{\mu,\lambda}\) be the order-preserving function from
            \(\mu\cup Y\) onto  the ordinal \(\mu+\delta\).  This lifts canonically to
            an order-preserving bijection \(\varphi_{\mu,\lambda}:
            \operatorname{Fn}(\mu\cup Y,2) \mapsto
            \operatorname{Fn}(\mu+\delta,2)\).
            Similarly, we make sense of the name
             \(\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})\), call it \(F_M\).
            Here simply, for each tuple \((~(k,m)~, p)\in \dot
            f_{[\lambda,\lambda+\omega)}\),
             we have that \((~(k,m)~,\varphi_{\mu,\lambda}(p))\) is in \(F_M\).
            Again, let \(\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})\)
            be interpreted in the above sense as giving \(F_M\) (which is an element
            of \(M\)).  Note that we do not regard \(\delta\) as fixed here, but
            rather simply depending on the \(\operatorname{supp}(\dot
            f_{[\lambda,\lambda+\omega)})\) described above. Other values
            replacing \(\lambda>\mu\) will result in their own set \(Y\)
            and canonical map \(\varphi_{\mu,\lambda}\); but one thing we do have to
            assume (or arrange) for other values \(\alpha\) replacing \(\lambda\)
            is that \(\operatorname{supp}(\dot f_{[\alpha,\alpha+\omega)})\)
            intersected with
            \(\alpha\) is contained in \(\mu\).

            Now the object \(F_M\) is an
            element of \(M\), and \(M\) believes this statement is true:
            \[
            (\forall \beta\in\omega_2)~ (\exists \beta<\lambda\in \omega_2)~~~
            \operatorname{supp}(\dot f_{[\lambda,\lambda+\omega)})
            \cap \lambda\subset \mu \ \ \mbox{and}\ \
            F_M = \varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})\]

            But now, this  means that,
            not only is there  an \(\alpha\in M\),
            \( F_M = \varphi_{\mu,\alpha}(\dot f_{[\alpha,\alpha+\omega)})\)
            but also that there is an increasing sequence
             \(\{\alpha_\xi : \xi  \in \omega_1\}\subset \lambda\) of
             such \(\alpha\)'s satisfying that, for each \(\xi\)
            we have that \(\operatorname{supp}(\dot
            f_{[\alpha_\xi,\alpha_\xi+\omega)})\) is contained in \(\alpha_{\xi+1}\).

            Choose such a sequence.
             This means that if we let \(A = \bigcup_{n>0} [\alpha_n,\alpha_n+\omega)\) we
             have the name \(\dot f_A\) in \(M\). This then means that all
             the \(( (\beta,m) , p)\) appearing in \(\dot f_A\) have the property
             that \(\mathop{dom}(p)\) is contained in \(M\).
            There is, within \(M\), a name \(\dot g\) satisfying that
             \(\dot f_A(\alpha_n+k) = \dot f_{[\alpha_n,\alpha_n+\omega)}(\alpha_n+k)\)
            for all \(k >\dot g(n)\).

            \bigskip

            We now apply the above Lemma using \(\mu = \mu_0\) and we are now
            working in the extension \(V[G_\mu]\). We will abuse the notation
            and use \(\dot f_{[\alpha_n,\alpha_n+\omega)}\) instead of
              \(\dot f_{[\alpha_n,\alpha_n+\omega)}(G_{\mu})\) as defined in
              the Lemma.
            We work for a contradiction. Something special has now happened,
             namely,  the supports of the names \(\{
            \dot f_{[\alpha_n, \alpha_n+\omega)} :  0< n<\omega\}\) are pairwise disjoint
            and also disjoint from the support of the name
             \(\dot f_{[\lambda,\lambda+\omega)}\) (under the same convention about
             \(G_\mu\).  And not only that, these names are pairwise isomorphic (in
             the way that they all map to \(F_M\)).

            \bigskip

            Since \(A\) is disjoint from \([\lambda,\lambda+\omega)\),
            there must be an integer \(\ell\)
            together with a condition
             \(q\in \mathop{Fn}(\omega_2\setminus \mu,2)\) satisfying that for all
             \(n>\ell\) , \(q\) forces that

            \centerline{
             ``if \(k>\dot g(n)\) (since  \(\alpha_n+k\in A\)) then
             \(\dot f_{[\alpha_n,\alpha_n+\omega)}(\alpha_n+k) \neq
            \dot f_{[\lambda,\lambda+\omega)}(\lambda +k)\)''.}

            Choose \(n\) large enough so that \(\mathop{dom}(q) \cap [\alpha_n,
            \mu_{n+1})\) is
            empty.
             Choose \(q_1<q\restriction \lambda\) (in \(M\)) so that
            \[ \varphi_{\mu,\alpha_n}(q_1\restriction \operatorname{supp}(
             \dot f_{[\alpha_n,\alpha_n+\omega)}) =
            \varphi_{\mu,\lambda}(q\restriction \operatorname{supp}(
             \dot f_{[\lambda,\lambda+\omega)}) \]
            and then (again in \(M\)) choose \(q_2 < q_1\)
            so that it both forces a value
             \(L\) on \(\ell+\dot g(n)\)
            and subsequently forces a value \(m\) on
             \(\dot f_{[\alpha_n, \alpha_n+\omega)}(\alpha_n+L+1)\).
            But now, again calculate
            \[ q_3 = \varphi_{\mu,\lambda}^{-1} \circ \varphi_{\mu,\alpha_n}
            (q_2\restriction \operatorname{supp}(\dot
            f_{[\alpha_n,\alpha_n+\omega)}))\]
            and, by the isomorphisms, we have that \(q_3\) forces that
             \(\dot f_{[\lambda,\lambda+\omega)}(\lambda+L+1) = m\).

            Technically (or with more care)  all of this is taking place in the
            poset \(\operatorname{Fn}(\omega_2\setminus \mu,2)\) and this means
            that \(q_3\) and \(q\) are all compatible with each other.

            Follow
            the bouncing ball: it suffices to consider
             \(q(\beta)=e\) and to assume that \(q_3(\beta)\) is defined.
            Since \(q_3(\beta)\) is defined, we have that
            there is a  \(\beta'\in\mathop{dom}(q_2)\) such that \(
             \varphi_{\mu,\lambda}(\beta) = \varphi_{\mu,\alpha_n}(\beta')\),
            and that \(q_3(\beta) = q_2(\beta')\).
            But, by definition of \(q_1\), \(\beta'\in\mathop{dom}(q_1)\)
            and even  that \(q_1(\beta') = q(\beta)\).
             Then, since \(q_2<q_1\), we have that \(q_2(\beta')=q_1(\beta') =
             q(\beta)\). This completes the circle that \(q_3(\beta) = q(\beta)\).
            \bigskip

            Finally, our contradiction is that \(q_3\cup q_2\cup q\)
            forces that
             \(k=L+1\) violates the quoted statement above.
            \end{proof}

  We are also able to force \(\alcompSP{\kappa}\) to fail for every cardinal
  other than the first \(\omega\)-many we've already guaranteed.

\begin{theorem} It follows from the existence of a $2$-huge cardinal
that there is a model of $ZFC$ for which $\lnot \mathcal A'(\aleph_{\omega})$.
\end{theorem}

            \begin{proof}
            We will need the model constructed in \cite{MR1045371} in
             which an instance of Chang's conjecture
             $(\aleph_{\omega+1},\aleph_\omega) \rlap{$\rightarrow$}\,{\rightarrow}
             (\aleph_1,\aleph_0)$ is shown to fail.

            We can take as a given (as shown in \cite[Theorem 5]{MR1045371}) that we may
            assume that we have a model $V$
            of GCH in which there are regular limit cardinals $\kappa<\lambda$
            satisfying that
            $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
            (\kappa^{+\omega+1},\kappa^{+\omega})$.

            What this says is that if $L$ is a countable language
            with at least one unary relation symbol $R$ and
             $M$ is a model of $L$  with base set $\lambda^{+\omega+1}$
            in which the interpretation of $R$ has cardinality
             $\lambda^{+\omega}$, then $M$ has an elementary submodel
             $N$ of cardinality $\kappa^{+\omega+1}$ in which
            $R\cap N$ has cardinality $\kappa^{+\omega}$ (of course
             $R\cap N$ is the interpretation of $R$ in $N$ because
            $N\prec M$).

            The interested reader will want to know that it is shown in
            \cite{MR1045371} that if $\kappa$ is a 2-huge cardinal
            and $j$ is the 2-huge
            embedding with critical point $\kappa$, then with
             $\lambda = j(\kappa)$ one has that
            $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
            (\kappa^{+\omega+1},\kappa^{+\omega})$ holds. There is no loss of
            generality to  also assume that GCH holds in this model.

            \bigskip

            Let $\{ h_\xi : \xi \in \lambda^{+\omega+1}\}$
            be a scale in $\Pi\{ \lambda^{+n+1} : n\in \omega\}$
             ordered by the usual mod finite coordinatewise
            ordering. For convenience we may assume that $h_\xi(n) \geq
            \lambda^{+n}$ for all $\xi$ and all $n$. For each integer
             $m$ the cofinality of the mod finite ordering on
             $\Pi\{ \lambda^{+n+1} : m< n\in \omega\}$
            is the same as it is for the entire product
             $\Pi\{ \lambda^{+n+1} : n\in \omega\}$. If $P$
            is any poset of cardinality less than
             $\lambda^{+m}$ then, in the forcing extension by $P$,
            every function in
             $\Pi\{ \lambda^{+n+1} : m< n\in \omega\}$ is bounded
            above by a ground model function.
            It therefore
            follows easily that  in the forcing
            extension by $P$,  the sequence
            $\{ h_\xi : \xi \in \lambda^{+\omega+1}\}$  remains cofinal in
             $\Pi\{ \lambda^{+n+1} : n\in \omega\}$.

            \bigskip

            The forcing notion $\mathbb P_0$ is simply the finite condition
            collapse of $\kappa^{+\omega}$, i.e. $\mathbb P_0 = \left(
             \kappa^{+\omega}\right)^{<\omega}$. In the forcing extension by
             $\mathbb P_0$, one now has that the ordinal $\kappa^{+\omega+1}$
             from $V$ is the first uncountable cardinal $\aleph_1$. Then in this
             forcing extension we let $\mathbb P_1$ be the countable condition
             Levy collapse, $Lv(\lambda,\omega_2)$, which collapses all cardinals
             less than $\lambda$ to have cardinality at most $\aleph_1$. The poset
             $\mathbb P_1$ has cardinality $\lambda$. We treat
            $\mathbb P*\mathbb P_1$ as containing $\mathbb P_0$
            as a subposet by identifying each $(p_0,1)$ with $p_0$.
            After forcing with $\mathbb P_0*\mathbb P_1$ we will have
            that $\omega_1$ is the ordinal $\left(\kappa^{+\omega+1}\right)^V$,
             $\omega_2$ is the ordinal $\lambda$, and $\omega_\omega$ is the
             ordinal $\left(\lambda^{+\omega}\right)^V$.

            \bigskip

            \newpage

            Now we assume that we have an assignment $\dot f_{\dot A}$ of a
             $\mathbb P_0*\mathbb P_1$-name of a finite-to-one function from
             $\dot A$ into $\omega$ for each
             $\mathbb P_0*\mathbb P_1$-name of a countable subset of
             $\lambda^{+\omega+1}$. We will obtain a contradiction.

            Let $\{ \dot A_\xi : \xi \in \lambda^{+\omega+1}\}$ be an enumeration
            of all the  nice $\mathbb P_0$-names of countable
            subsets of $\lambda^{+\omega}$. For each $\xi\in \lambda^{+\omega+1}$,
             let $\dot f_\xi$ be another notation for $\dot f_{\dot A_\xi}$.
            Since $\mathbb P_0$ forces that $\mathbb P_1$ is countably closed,
            the collection of all nice  $\mathbb P_0$-names will produce all
            the countable sets in the extension by $\mathbb P_0* \mathbb
            P_1$, but $\mathbb P_0*\mathbb P_1$ can introduce new enumerations of
            these names.
            For each $\xi\in \lambda^{+\omega+1}$,
            there is a minimal $\zeta_\xi$ so that $\dot A_{\zeta_\xi}$ is the
            canonical name for the range of $h_\xi$. This means that
             $\dot f_{\zeta_\xi} \circ h_\xi$ is simply the
             $\mathbb P_0 *\mathbb P_1$-name of a finite-to-one function from
             $\omega$ to $\omega$.
              For each $\xi\in \lambda^{+\omega+1}$,
             choose any $p_\xi \in \mathbb P_0*\mathbb P_1$ so that
             there is a nice $\mathbb P_0$-name, $\dot H_\xi$,
             that is forced by
             $p_\xi$  to equal $\dot f_{\zeta_\xi}\circ h_\xi$.
              Choose $\Lambda\subset \lambda^{+\omega+1}$
              of cardinality $\lambda^{+\omega+1}$
              and so that there is a pair $p,\dot H$
              satisfying that $p_\xi = p$ and $\dot H_\xi = \dot H$
              for all $\xi\in \Lambda$. We may assume that $p$
              is in a generic filter $  G$.

            Let $\{ x_\xi : \xi\in \lambda^{+\omega+1}\}$ be any enumeration of
             $H(\lambda^{+\omega+1})$  such that
             $\{ x_\xi : \xi\in \lambda^{+\omega}\}$ is also equal to
             $H(\lambda^{+\omega})$. We choose this enumeration in such a way that
             $x_\xi\in x_\eta$ implies $\xi<\eta$.
            We use relation symbol $R_0$ to code
            (and well order)
             $(H(\lambda^{+\omega+1}), \in)$ as follows: $(\xi,\eta)\in R_0$
            if and only if $x_\xi\in x_\eta$.
            Let $R_1$ be a binary relation on $\kappa^{+\omega}$ so that
            $(\kappa^{+\omega},R_1)$ is isomorphic to $\mathbb P_0$. Let
             $R_2$ be a binary relation on $\lambda$ so that
             $R_2\cap (\kappa^{+\omega}\times \kappa^{+\omega})=R_1$ and
             $(\lambda,R_2)$ is isomorphic to $\mathbb P_0*\mathbb P_1$.
            Let $\psi$ be the poset isomorphism from $(\lambda,R_2)$ to
             $\mathbb P_0*\mathbb P_1$.

            We continue coding. We can code the sequence
             $\{ h_\xi : \xi\in \lambda^{+\omega+1}\}$ as another binary  relation
             $R_3$ on $\lambda^{+\omega+1}$ where $R_3\cap \left(\{\xi\}\times
             \lambda^{+\omega+1}\right) = \{ (\xi,h_\xi(n) ) : n\in \omega\}$
            for each $\xi\in \lambda^{+\omega+1}$. The relation symbol $R_4$ can
            code the sequence $\{ \dot A_\xi : \xi \in \lambda^{+\omega+1}\}$
            where $(\xi, \alpha, \zeta) \in R_4$ if and only if
            $(\check \alpha, \psi(\zeta))$ is in the name $\dot A_\xi$.
             Let $R_5$ code this collection, i.e.
             $(\gamma,n,m,\eta)\in R_5$ if and only if
             $(\check{(n,m)}, \psi(\eta))\in \dot H_\gamma$. Also let $R_6$ code
             (equal)  the set $\Lambda$.
            Finally we use the relation symbol $R_7$ to similarly code the
            sequence $\{ \dot f_\xi  : \xi \in \lambda^{+\omega+1}\}$:
             $(\xi, \alpha, n, \zeta) \in R_7$ if and only if
            $(\check {(\alpha,n)}, \psi(\zeta))$ is in the name $\dot f_\xi$.

            Needless to say, the unary relation symbol $R$
            is interpreted as the set $\lambda^{+\omega}$ for the application of
            $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
            (\kappa^{+\omega+1},\kappa^{+\omega})$.
             Now  we have defined our model
            $M$  of the language $L= \{\in, R, R_0,\ldots, R_7\}$,
             and we choose an elementary submodel
            $N$ witnessing
            $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
            (\kappa^{+\omega+1},\kappa^{+\omega})$. Of course $N$ is really just
            a $\kappa^{+\omega+1}$ sized subset of $\lambda^{+\omega+1}$
            with the additional property that $N\cap \lambda^{+\omega}$ has
            cardinality $\kappa^{+\omega}$.  In the
            forcing extension $N$ has cardinality $\omega_1$ and
             $A= N\cap \lambda^{+\omega}$ is countable.

            We will need the following claim from \cite{MR1045371}:

            \begin{claim*} We may assume that $N$ satisfies that
             $N\cap \kappa^{+\omega+1}$ is transitive (i.e. an initial segment).
            \end{claim*}

            \bgroup
            \def\proofname{Proof of Claim:\/}

            \begin{proof}
            Suppose  our originally supplied $N$ fails the conclusion of the
            claim.
            We   know that $\kappa^{+\omega}\in N$,
            (  via $R_1$)
            in which case so is $\kappa^{+\omega+1}$.

            Then set $\beta_0 = \sup(N\cap \kappa^{+\omega+1})$ and consider the
            Skolem closure $\mathop{Hull}(N\cup \beta_0, M)$.  A little informally
            (in that we have to formalize the enumeration of formulas as per
            G\"odel coding)
            let $\{\varphi_n : n\in \omega\}$ be an enumeration of all formulas
            in the language $L$, and let  $\ell_n$ be the minimal integer such
            that the free variables of $\varphi_n$ are among
             $\{ v_0, \ldots, v_{\ell_n}\}$.
            Then,   for each tuple
            $\langle \xi_1,\ldots, \xi_{\ell_n} \rangle$
            of elements of $\lambda^{+\omega+1}$,
             we define $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be
             the minimal $\xi_0\in \lambda^{+\omega+1}$ such that
             $M\models \varphi_n(\xi_0,\ldots, \xi_{\ell_n})$.
            If there is no such $\xi_0$,
               in other words if $M\models \lnot\exists x~~  \varphi_n(x,\xi_1,
            \ldots,\xi_{\ell_n})$, then set
             $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be $0$.
            Now $\mathop{Hull}(N\cup \beta_0,M)$ is just the minimal superset
             $X$
            of $N\cup \beta_0$ that satisfies that
             $f_n[X^{\{1,\ldots,\ell_n\}}] \subset X$ for all $n$. Since this is
             simply a large algebra, we can generate all the terms $t$ of
            the algebraic operations $\{ f_n : n\in \omega\}$.
            It is easily seen that for each $\zeta \in X$, there is a term
            $t(v_1,\ldots, v_m)$ such that $\zeta  = t(\delta_1,\ldots, \delta_m)$
             for some sequence $\langle \delta_1,\ldots, \delta_m\rangle$ with
             each $\delta_i\in  N\cup \beta_0$.
             Assume that
              $\zeta\in \kappa^{+\omega+1}$. By
              re-indexing the variables in the term we can assume
              that there is an $n\leq m$ so that
               $\delta_i  <\beta_0$ for $1\leq i\leq n$
                and $\kappa^{+\omega+1}\leq \delta_i$ for $n<i\leq m$.
                Let $\vec a $ denote
                the tuple $\langle \delta_{n+1},\ldots, \delta_m\rangle$.
              Choose $\eta\in N\cap \kappa^{+\omega+1}$ large enough so that
             $ \{  \delta_1, \ldots, \delta_n\}$ is contained in  $\eta$. Since set-membership
             in $M$ is coded by $R_0$ rather than $\in$ we have to argue a little
             less naturally.
             Consider the set
               $s_0(\eta,\vec a) = \{ t(\gamma_1,\ldots, \gamma_n,\vec a) :
            \{\gamma_1,\cdots,\gamma_n\} \in [\eta]^{\leq n}\}$. Clearly
             $s_0(\eta,\vec a)$ is a member of $H(\lambda^{+\omega+1})$.
             Now define $s_1(\eta,\vec a) $ to be
             $\{ x_\alpha : \alpha\in s_0(\eta, \vec a)\}$,
             and choose the
             unique $\zeta_1\in \lambda^{+\omega+1}$ such
             that $x_{\zeta_1} = s_1(\eta,\vec a)$. We claim that $\zeta_1\in N$.
             Note that $\alpha R_0\zeta_1$ holds if and only if
              $ \alpha \in s_0(\eta,\vec a)$, and therefore
             $$M\models (\forall \alpha)\left[\alpha R_0 \zeta_1 ~~\mbox{iff}~~
             (\exists   \gamma_1\in \eta)\cdots(\exists \gamma_n\in\eta)(\alpha=
             t(\gamma_1,\ldots, \gamma_n,\vec a))\right]~.$$
             By elementarity then we have that $\zeta_1\in N$,
             and by similar reasoning the supremum, $\zeta_0$,
            of $\zeta_1\cap \kappa^{+\omega+1}$ is also in $N$.
               This of course means that $\zeta < \xi_0$.
            \end{proof}

            \egroup

            We use the elementarity of
             $N$  to deduce properties of the families
             $\{ \dot A_\xi : \xi\in N\}$ and $\{ \dot f_\xi : \xi \in
             N\}$.  Actually the collection we are most interested in
             is the family $\{ h_\xi : \xi\in \Lambda\cap N\}$.

            Now we need a result from Shelah's pcf theory which
            is proven in  Jech \cite[24.9]{MR1940513}.
             Since  $\aleph_1=\mathfrak c < \kappa^{+\omega+1}$ there
              is a function $\langle \varrho_n : n\in \omega\rangle$
              in $\Pi_n \lambda^{+\omega}   $ such that the sequence
               $\{ h_\xi : \xi \in N\}$ is unbounded mod finite
               in $\Pi_n \varrho_n$
               For each $n$, $\varrho_n \leq \sup(N\cap \lambda^{+n+2})$.
                  Since $\mathbb P_0$ has cardinality $\kappa^{+\omega}$,
            and so less than $|N|=\kappa^{+\omega+1}$, a standard argument
            (analogous to the fact that adding a Cohen real does
             not add a dominating real)
             shows that    the sequence $\{ h_\xi : \xi \in \Lambda\cap N\}$ remains
             unbounded
                mod finite in $\Pi_n \varrho_n$ (and in
                 $\Pi_n (\varrho_n\cap N)$).

            Now pass to the extension by
                  $G\cap \mathbb P_0$ and let $H$ be the function
                   $\operatorname{val}_{G}(\dot H)$, and we recall
                   that $f_{\zeta_\xi}(h_\xi(n)) = H(n)$ for all $n\in \omega$
            and $\xi\in \Lambda$.
             Now pass to the full extension $V[G]$ and again, since
              $\mathbb P_1$ was forced to be countably closed,
               the family $\{ h_\xi : \xi \in \Lambda\cap N\}$ is still
               unbounded in $\Pi_n (\varrho_n \cap N)$ (no new elements
            were added). We let
               $A$ be the countable set $N\cap \lambda^{+\omega}$,
               and for  each
               $\xi\in \Lambda\cap N$, there is an $n_\xi$ such that
                $f_{\xi}(h_\xi(m)) = f_A(h_\xi(m))$ for all $m>n_\xi$.
                There is a single $n$ so that $\Lambda_n
                 = \{\xi\in \Lambda\cap N       : n_\xi = n\}$ has cardinality
                  $\omega_1 $, and thus
                  $\{ h_\xi : \xi\in \Lambda_n\cap N\}$ is also unbounded
                  in $\Pi_n (\rho_n\cap N)$.
            This certainly implies that there is an $m>n$
            such that $\{ h_\xi (m) : \xi\in \Lambda_n\cap N\}$ is infinite.
            This completes the proof since  $f_{A}(h_\xi(m)) = H(m)$
            for all $\xi\in \Lambda_n\cap N$.
            \end{proof}

  \section{Applications to infinite length games}

  We introduce three variations of Scheeper's game which we defined in
  the introduction.

  \begin{game}
    Let \(\schFillGame\kappa\) denote the
    \term{Scheepers countable-finite union game} which proceeds analogously
    to \(\schFillStrictGame\kappa\), except that \(\pl C\)'s restriction in round \(n+1\)
    is weakened to \(C_{n+1}\supseteq C_n\).
  \end{game}

  \begin{game}
    Let \(\schFillInitialGame\kappa\) denote the
    \term{Scheepers countable-finite initial game} which proceeds analogously
    to \(\schFillGame\kappa\), except that \(\pl F\)'s winning condition is
    weakened to \(\bigcup_{n<\omega}F_n\supseteq C_0\).
  \end{game}

  \begin{game}
    Let \(\schFillIntGame\kappa\) denote the
    \term{Scheepers countable-finite intersection game} which proceeds analogously
    to \(\schFillInitialGame\kappa\), except that \(\pl C\) may choose any
    \(C_n\in[\kappa]^{\leq\omega}\) each round, and \(\pl F\)'s winning condition
    is weakened to
    \(\bigcup_{n<\omega}F_n\supseteq\bigcap_{n<\omega}C_n\).
  \end{game}

  In \cite{clontzMengerGamePreprint} Clontz extended Scheepers' application of
  almost-compatible injections to these game variants as well as
  \(\menGame{\oneptlind\kappa}\). However, when considering Markov strategies,
  finite-to-one functions suffice.

  \begin{figure}[h]
\begin{center}
\begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=1em,minimum width=2em]
  {
    &
    \pl F \kmarkwin2\menGame{\oneptlind\kappa} &
    \pl F \kmarkwin2\schFillIntGame\kappa \\

    \alcompSP\kappa \\

    &
    \pl F \kmarkwin2\schFillInitialGame\kappa &
    \pl F \kmarkwin2\schFillGame\kappa \\
  };
  \path[>=latex,->]
    (m-2-1) edge (m-1-2)
    (m-2-1) edge (m-3-2)
    (m-1-2) edge (m-3-2)
    (m-1-3) edge (m-3-3);
  \path[>=latex,<->]
    (m-1-2) edge (m-1-3)
    (m-3-2) edge (m-3-3);
\end{tikzpicture}
\end{center}
\caption{Diagram of Scheeper/Menger game implications with \(\alcompSP{\kappa}\)}
\label{GamesDiagram2}
\end{figure}

%   \begin{figure}[h]
% \begin{center}
% \begin{tikzpicture}
%   \matrix (m) [matrix of math nodes,row sep=3em,column sep=1em,minimum width=2em]
%   {
%     &
%     \pl F \kmarkwin2\menGame{\oneptlind\kappa} &
%     \pl F \kmarkwin2\schFillIntGame\kappa \\

%     \alcompSP\kappa &
%     \pl F \kmarkwin2\schFillInitialGame\kappa &
%     \pl F \kmarkwin2\schFillGame\kappa \\

%     &
%     &
%     \pl F \ktactwin2\schFillStrictGame\kappa \\
%   };
%   \path[>=latex,->]
%     (m-2-1) edge (m-1-2)
%     (m-2-1) edge (m-2-2)
%     (m-2-1) edge (m-3-3)
%     (m-1-2) edge (m-2-2)
%     (m-1-3) edge (m-2-3)
%     (m-3-3) edge (m-2-3);
%   \path[>=latex,<->]
%     (m-1-2) edge (m-1-3)
%     (m-2-2) edge (m-2-3);
% \end{tikzpicture}
% \end{center}
% \caption{Diagram of Scheeper/Menger game implications with \(\alcompSP{\kappa}\)}
% \label{GamesDiagram2}
% \end{figure}

  \begin{theorem}
    \(\alcompSP\kappa\) implies the game-theoretic conditions
    shown in Figure \ref{GamesDiagram2}.
  \end{theorem}

  \begin{proof}
    The weaker claim
    \(\alcompS\kappa\Rightarrow\pl F \kmarkwin2\schFillIntGame\kappa\)
    was proven in \cite{clontzMengerGamePreprint};
    however, the strategy
    only required that the \(f_A\) be pairwise almost-compatible and
    that preimages of finite sets in each \(f_A\)
    are finite, which is also possible assuming \(\alcompSP\kappa\).
    Note that the relationships amongst the games were all shown in
    \cite{clontzMengerGamePreprint}.
  \end{proof}

  We include the following proof from \cite{MR1129143} for the convenience
  of the reader.

  \begin{theorem}
    \(\alcompS\kappa\Rightarrow\pl F \ktactwin2\schFillStrictGame\kappa\)
  \end{theorem}

  \begin{proof}
    Let \(\{f_A:A\in[\kappa]^{\leq\aleph_0}\}\) witness
    \(\alcompS\kappa\), and define \(g_A:A\to\omega\) by
    \(g_A(\alpha)=f_A(\alpha)-|\{\beta\in A:f_A(\beta)<f_A(\alpha)\}|\).

    We claim that \(\{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}\)
    must be finite as it is bounded above by
    \(\max\{M,f_A(\alpha),f_B(\alpha):f_A(\alpha)\not=f_B(\alpha)\}\)
    where \(M=f_B(\alpha)\) for some \(\alpha\in B\setminus A\).
    To see this, let \(f_A(\alpha)=f_B(\alpha)=N>M\) and assume
    \(f_A(\beta)\not=f_B(\beta)\) implies \(f_A(\beta),f_B(\beta)<N\).
    Then
    \[
      g_A(\alpha)
        =
      N-|\{\beta\in A:f_A(\beta)<N\}|
        >
      N-|\{\beta\in B:f_B(\beta)<N\}|
        =
      g_B(\alpha)
    \]
    with the strictness of the inequality witnessed by \(f_B(\alpha)=M<N\)
    for some \(\alpha\in B\setminus A\).

    As a result,
    \[
      \sigma(\<A,B\>)
        =
      \{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}
    \]
    is a legal \(2\)-tactic for \(\pl F\). Let \(C=\<C_0,C_1,\dots\>\) be
    a strictly increasing sequence of countable sets and
    \(\alpha\in C_n\). Noting that \(f_A\) is an injection and not
    just finite-to-one, \(0\leq g_{C_{n+m}}(\alpha)\) for all
    \(m<\omega\), and it follows that
    \(g_{C_{n+m}}(\alpha)\leq g_{C_{n+m+1}}(\alpha)\) for some \(m<\omega\).
    Therefore \(\alpha\in\sigma(\<C_{n+m},C_{n+m+1}\>)\).
  \end{proof}

  So it would seem that \(\alcompSP\kappa\) is sufficient only when
  considering Markov strategies. (Of course,
  \(\alcompSP\kappa\Rightarrow\pl F\kmarkwin2\schFillGame\kappa
  \Rightarrow\pl F\kmarkwin2\schFillStrictGame{\kappa}\).)
  We would like to demonstrate that \(\alcompSP\kappa\) is not
  necessary.

  \begin{theorem}
    Let \(\alpha\) be the limit of increasing ordinals \(\beta_n\) for \(n<\omega\).
    If \(\pl F \kmarkwin2\schFillIntGame{\aleph_{\beta_n}}\) for all
    \(n<\omega\), then \(\pl F\kmarkwin2\schFillIntGame{\aleph_\alpha}\).
  \end{theorem}

  \begin{proof}
    Let \(\sigma_n\) be a winning \(2\)-mark for \(\pl F\) in
    \(\schFillIntGame{\aleph_{\beta_n}}\). Define the \(2\)-mark \(\sigma\)
    for \(\pl F\) in \(\schFillIntGame{\aleph_\alpha}\) as follows:
    \[
      \sigma(\<C\>,0)
        =
      \sigma_0(\<C\cap\aleph_{\beta_0}\>,0)
    \]
    \[
      \sigma(\<C,D\>,n+1)
        =
      \sigma_{n+1}(\<D\cap\aleph_{\beta_{n+1}}\>,0)
        \cup
      \bigcup_{m\leq n}
      \sigma_m(\<C\cap\aleph_{\beta_m},D\cap\aleph_{\beta_m}\>,n-m+1)
    \]

    Let \(\<C_0,C_1,\dots\>\) be an attack by
    \(\pl C\) in \(\schFillIntGame{\aleph_\alpha}\), and
    \(\alpha\in\bigcap_{n<\omega}C_n\).
    Choose \(N<\omega\) with \(\alpha<\aleph_{\beta_{N+1}}\). Consider the
    attack
    \(\<C_{N+1}\cap\aleph_{\beta_{N+1}},C_{N+2}\cap\aleph_{\beta_{N+1}},\dots\>\)
    by \(\pl C\) in \(\schFillIntGame{\aleph_{\beta_{N+1}}}\). Since
    \(\sigma_{N+1}\) is a winning strategy and
    \(\alpha\in\bigcap_{n<\omega}C_{N+n+1}\cap\aleph_{\beta_{N+1}}\), either
    \(\alpha\in\sigma_{N+1}(\<C_{N+1}\cap\aleph_{\beta_{N+1}}\>,0)\) and thus
    \(\alpha\in\sigma(\<C_N,C_{N+1}\>,N+1)\), or
    \(
      \alpha
        \in
      \sigma_{N+1}(
        \<C_{N+M+1}\cap\aleph_{\beta_{N+1}},C_{N+M+2}\cap\aleph_{\beta_{N+1}}\>
        ,M+1
      )
    \)
    for some \(M<\omega\) and thus
    \(
      \alpha
        \in
      \sigma(
        \<C_{N+M+1},C_{N+M+2}\>,
        N+M+2
      )
    \). Thus \(\sigma\) is a winning strategy.
  \end{proof}

  \begin{theorem}
    Let \(\alpha\) be the limit of increasing ordinals \(\beta_n\) for \(n<\omega\).
    If \(\pl F \kmarkwin2\schFillInitialGame{\aleph_{\beta_n}}\) for all
    \(n<\omega\), then \(\pl F\kmarkwin2\schFillInitialGame{\aleph_\alpha}\).
  \end{theorem}

  \begin{proof}
    The proof proceeds nearly identically to the previous proof.
  \end{proof}
    % Let \(\sigma_n\) be a winning \(2\)-mark for \(\pl F\) in
    % \(\schFillInitialGame{\omega_{\beta_n}}\). Define the \(2\)-mark \(\sigma\)
    % for \(\pl F\) in \(\schFillInitialGame{\omega_\alpha}\) as follows:
    % \[
    %   \sigma(\<C\>,0)
    %     =
    %   \sigma_0(\<C\cap\omega_{\beta_0}\>,0)
    % \]
    % \[
    %   \sigma(\<C,D\>,n+1)
    %     =
    %   \sigma_{n+1}(\<D\cap\omega_{\beta_{n+1}}\>,0)
    %     \cup
    %   \bigcup_{m\leq n}
    %   \sigma_m(\<C\cap\omega_{\beta_m},D\cap\omega_{\beta_m}\>,n-m+1)
    % \]

    % Let \(\<C_0,C_1,\dots\>\) be an attack by
    % \(\pl C\) in \(\schFillInitialGame{\omega_\alpha}\), and
    % \(\alpha\in C_0\).
    % Choose \(N<\omega\) with \(\alpha<\omega_{\beta_{N+1}}\). Consider the
    % attack
    % \(\<C_{N+1}\cap\omega_{\beta_{N+1}},C_{N+2}\cap\omega_{\beta_{N+1}},\dots\>\)
    % by \(\pl C\) in \(\schFillInitialGame{\omega_{\beta_{N+1}}}\). Since
    % \(\sigma_{N+1}\) is a winning strategy and
    % \(\alpha\in C_{N+1}\cap\omega_{\beta_{N+1}}\), either
    % \(\alpha\in\sigma_{N+1}(\<C_{N+1}\cap\omega_{\beta_{N+1}}\>,0)\) and thus
    % \(\alpha\in\sigma(\<C_N,C_{N+1}\>,N+1)\), or
    % \(
    %   \alpha
    %     \in
    %   \sigma_{N+1}(
    %     \<C_{N+M+1}\cap\omega_{\beta_{N+1}},C_{N+M+2}\cap\omega_{\beta_{N+1}}\>
    %     ,M+1
    %   )
    % \)
    % for some \(M<\omega\) and thus
    % \(
    %   \alpha
    %     \in
    %   \sigma(
    %     \<C_{N+M+1},C_{N+M+2}\>,
    %     N+M+2
    %   )
    % \). Thus \(\sigma\) is a winning strategy.

  \begin{corollary}
    It is consistent that \(\alcompSP{\aleph_\omega}\) fails,
    but as \(\alcompSP{\aleph_k}\) holds in \(ZFC\) for all \(k<\omega\),
    both
    \(\pl F\kmarkwin2\schFillIntGame{\aleph_\omega}\) and
    \(\pl F\kmarkwin2\schFillInitialGame{\aleph_\omega}\) hold in \(ZFC\).
  \end{corollary}

  Note that Question \ref{mainQuestion} remains unsolved; however,
  our results have revealed that we cannot hope to find a \(ZFC\)
  counterexample where \(X=\oneptlind\kappa\). This is
  because if we also assume \(V=L\), it follows that
  \(\alcompSP{\kappa}\) and therefore \(\pl F\kmarkwin2\menGame{X}\)
  for every cardinal. Although, one may consider the following weaker question.

  \begin{question}
    Is there a model of \(ZFC\) where both
    \(\pl F\kmarkwin2\menGame{\oneptlind{\aleph_{\omega+1}}}\) and
    \(\alcompSP{\aleph_\omega}\) fail?
  \end{question}

  % \section{Generalized Kurepa families and almost-compactible functions}

  % TODO

\bibliographystyle{plain}
\bibliography{../../bibliography}

\end{document}