### 2016-09-08 articles/almost-compatible-functions-and-topological-games/article.tex
many many revisions to address the referee's concerns

\documentclass{amsart}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}

\usepackage{tikz}
% \usetikzlibrary{matrix}
\usetikzlibrary{arrows}

\usepackage{../../clontzDefinitions}

      \theoremstyle{plain}
      \newtheorem{theorem}{Theorem}
      \newtheorem{lemma}[theorem]{Lemma}
      \newtheorem{corollary}[theorem]{Corollary}
      \newtheorem{proposition}[theorem]{Proposition}
      \newtheorem{conjecture}[theorem]{Conjecture}
      \newtheorem{question}[theorem]{Question}
      \newtheorem{claim}[theorem]{Claim}

      \theoremstyle{definition}
      \newtheorem{definition}[theorem]{Definition}
      \newtheorem{example}[theorem]{Example}
      \newtheorem{game}[theorem]{Game}

      \theoremstyle{remark}
      \newtheorem{remark}[theorem]{Remark}

      \theoremstyle{plain}
      \newtheorem*{theorem*}{Theorem}
      \newtheorem*{lemma*}{Lemma}
      \newtheorem*{corollary*}{Corollary}
      \newtheorem*{proposition*}{Proposition}
      \newtheorem*{conjecture*}{Conjecture}
      \newtheorem*{question*}{Question}
      \newtheorem*{claim*}{Claim}

      \theoremstyle{definition}
      \newtheorem*{definition*}{Definition}
      \newtheorem*{example*}{Example}
      \newtheorem*{game*}{Game}

      \theoremstyle{remark}
      \newtheorem*{remark*}{Remark}

\parskip=0.5em

% dow stuff
          \def\Succ{\mathop{Succ}}
          \def\dom{\mathop{dom}}
          \def\val{\operatorname{val}}
          \def\N{\mathbb N}
          \def\Nstar{\mathbb N^*}

\begin{document}

% \title{Proximal compact spaces are Corson compact\tnoteref{t1}}
% \tnotetext[t1]{2010 Mathematics Subject Classification. 54E15, 54D30, 54A20.}
\title{Almost compatible functions and infinite length games}

% \author[aub]{S.~Clontz\fnref{fn1}}
% \ead{steven.clontz@auburn.edu}
% \author[aub]{G.~Gruenhage\fnref{fn2}}
% \ead{gruengf@auburn.edu}

% \address[aub]{Department of Mathematics, Auburn University,
%  Auburn, AL 36830}

\author{Steven Clontz}
\address{Department of Mathematics and Statistics,
The University of South Alabama,
Mobile, AL 36606}
\email{steven.clontz@gmail.com}
\author{Alan Dow}
\address{Department of Mathematics and Statistics, UNC Charlotte,
Charlotte, NC 28262}
\email{adow@uncc.edu}

% \keywords{TODO}
% \subjclass[2010]{}

\begin{abstract}
  \(\alcompSP\kappa\) asserts the existence of pairwise
  almost compatible finite-to-one functions \(A\to\omega\) for each
  countable subset \(A\) of \(\kappa\).
  The existence of winning \(2\)-Markov strategies in
  several infinite-length games, including
  the Menger game on the one-point Lindel\"ofication \(\oneptlind\kappa\)
  of \(\kappa\), are guaranteed by \(\alcompSP\kappa\). \(\alcompSP\kappa\) is
  implied by the existence of cofinal Kurepa families of size \(\kappa\),
  and thus holds for all cardinals less than \(\aleph_\omega\).
  It's consistent that \(\alcompSP{\aleph_\omega}\) fails, but
  there must always be a winning \(2\)-Markov strategy for the second
  player in the Menger game on \(\oneptlind{\omega_\omega}\).
\end{abstract}

\maketitle
\footnotetext[1]{The second author acknowledges support provided by
NSF-DMS 1501506}

\section{Introduction}

  \begin{definition}
    Two functions \(f,g\) are almost compatible, that is, \(f\sim g\) when
    \(\{a\in\dom f\cap\dom g:f(a)\not=g(a)\}\) is finite.
  \end{definition}

  Marion Scheepers used almost compatible functions in \cite{MR1129143} in order
  to study the existence of limited information strategies on a variation
  of the meager-nowhere dense game he introduced in \cite{MR1183703}.

  \begin{game}
    Let \(\schFillStrictGame\kappa\) denote
    \term{Scheepers' strict countable-finite union game}
    with two players \(\pl C\), \(\pl F\). In round \(0\), \(\pl C\) chooses
    \(C_0\in[\kappa]^{\leq\omega}\), followed by \(\pl F\) choosing
    \(F_0\in[\kappa]^{<\omega}\). In round \(n+1\), \(\pl C\) chooses
    \(C_{n+1}\in[\kappa]^{\leq\omega}\) such that \(C_{n+1}\supset C_n\), followed
    by \(\pl F\) choosing \(F_{n+1}\in[\kappa]^{<\omega}\).

    \(\pl F\) wins the game if
    \(\bigcup_{n<\omega} F_n\supseteq\bigcup_{n<\omega} C_n\); otherwise,
    \(\pl C\) wins.
  \end{game}

  Of course, with perfect information this game is trivial: during round
  \(n\) player \(\pl F\) simply chooses \(n\) ordinals from each of the
  \(n\) countable sets played by \(\pl C\). However, if \(\pl F\) is limited
  to using information from the last \(k\) moves by \(\pl C\) during each
  round, the task becomes more difficult. Call such a strategy a
  \term{\(k\)-tactical strategy} or \term{\(k\)-tactic}; if using the round
  number is allowed, then the strategy is called a \term{\(k\)-Markov strategy}
  or a \term{\(k\)-mark}.

  \begin{definition}
    The statement \(\alcompS\kappa\) (given as \(S(\kappa,\aleph_0,\omega)\)
    in \cite{MR1129143})
    claims that there exist one-to-one functions
    \(f_A:A\to\omega\) for each \(A\in[\kappa]^{\leq\aleph_0}\) such that
    the collection \(\{f_A:A\in[\kappa]^{\leq\aleph_0}\}\) is pairwise
    almost compatible.
  \end{definition}

  In \cite{MR1129143}, Scheepers noted that \(\alcompS{\omega_1}\) holds in \(ZFC\),
  and that it's possible to force \(\mf c\) to be arbitrarily large
  while preserving \(\alcompS{\mf c}\). However, \(\alcompS{\mf c^+}\) always fails.
  This axiom may be applied to obtain a winning \(2\)-tactic for \(\pl F\)
  in the countable-finite game.

  In \cite{clontzMengerGamePreprint}, Clontz related this game to a game
  which may be used to characterize the Menger covering property of a
  topological space.

  \begin{game}
    Let \(\menGame{X}\) denote the \term{Menger game} with players \(\pl C\),
    \(\pl F\).
    In round \(n\), \(\pl C\) chooses an open cover \(\mc U_n\), followed by
    \(\pl F\) choosing a
    subset \(F_n\) of \(X\) which may be finitely covered by \(\mc U_n\).

    \(\pl F\) wins the game if \(X = \bigcup_{n<\omega}F_n\),
    and \(\pl C\) wins otherwise.
  \end{game}

  This characterization is slightly different than the typical characterization
  in which the second player first chooses a specific
  finite subcollection \(\mc F_n\) of the cover itself and lets
  \(F_n=\bigcup \mc F_n\), denoted as \(G_{fin}(\mc O,\mc O)\)
  in \cite{MR1378387}.
  However, it's easily seen that these games are equivalent
  for perfect information strategies (so both characterize
  the Menger property in the same way), and this characterization is
  more convenient for our concerns.

  \begin{definition}
    Let \(\oneptlind\kappa=\kappa\cup\{\infty\}\) where \(\kappa\) is
    discrete and \(\infty\)'s neighborhoods are the co-countable sets
    containing it.
  \end{definition}

  The relationship between \(\schFillStrictGame\kappa\) and
  \(\menGame{\oneptlind\kappa}\) is strong; in both games \(\pl C\) essentially
  chooses a countable subset of \(\kappa\) followed by \(\pl F\) choosing
  a finite subset of that choice, and it's easy to see the winning perfect
  information strategy for \(\pl F\) in both games.
  In addition, it was shown in
  \cite{clontzMengerGamePreprint} that when \(\alcompS{\kappa}\) holds,
  \(\pl F\) has a winning \(2\)-Markov strategy in
  \(\menGame{\oneptlind\kappa}\).

  One source of motivation is to make progress on the following open question:

  \begin{question}\label{mainQuestion}
    Does there exist a topological space \(X\) for which
    \(\pl F\win\menGame{X}\) but \(\pl F\notkmarkwin{2}\menGame{X}\)?
    (That is, the second player can win the Menger game on \(X\)
    with perfect information but not with \(2\)-Markov information.)
  \end{question}

  \section{One-to-one and finite-to-one almost compatible functions}

  We may weaken Scheeper's \(\alcompS{\kappa}\) as follows:

  \begin{definition}
    The statement \(\alcompSP{\kappa}\) weakens \(\alcompS{\kappa}\) by only
    requiring the witnessing almost-compatible functions \(f_A:A\to\omega\)
    to be finite-to-one.
  \end{definition}

  \begin{proposition}
    \(\alcompS{\kappa}\) and \(\alcompSP{\kappa}\) need only be witnessed by
    functions \(\{f_A:A\in\mc S\}\) for some family \(\mc S\)
    cofinal in \([\kappa]^{\leq\aleph_0}\).
  \end{proposition}

  \begin{proof}
    For each \(A\in[\kappa]^{\leq\aleph_0}\) choose \(A'\supseteq A\)
    from \(\mc S\) and let \(g_A=f_{A'}\rest A\).
  \end{proof}

  In the final section we will show that \(\alcompSP{\kappa}\) is sufficient
  for many applications to the Scheepers and Menger games.
  In the meantime, we will demonstrate that \(\alcompSP{\kappa}\) is strictly
  weaker than \(\alcompS{\kappa}\).

  Recall the following.

  \begin{definition}
    A Kurepa family \(\mc K\subseteq[\kappa]^{\aleph_0}\) on \(\kappa\)
    satisfies that
    \(\mc K\rest A=\{K\cap A:K\in\mc K\}\) is countable
    for each \(A\in[\kappa]^{\aleph_0}\).
    Let \(\kurepaS\kappa\) be the statement claiming there exists
    a Kurepa family on \(\kappa\)
    cofinal in \([\kappa]^{\aleph_0}\).
  \end{definition}

  \begin{theorem}
    \(\kurepaS\kappa \Rightarrow \alcompSP{\kappa}\).
  \end{theorem}

  \begin{proof}
    Let \(\mc K=\{K_\alpha:\alpha<\theta\}\)
    be a cofinal Kurepa family on \(\kappa\).
    We first define \(f_\alpha:K_\alpha\to\omega\) for each \(\alpha<\theta\).

    Suppose we've already defined pairwise almost compatible finite-to-one functions
    \(\{f_\beta:\beta<\alpha\}\). To define
    \(f_\alpha\), we first recall that \(\mc K\rest K_\alpha\) is countable,
    so we may choose \(\beta_n<\alpha\) for \(n<\omega\) such that
    \(
      \{K_\beta:\beta<\alpha\}\rest K_\alpha \setminus \{\emptyset\}
        =
      \{K_\alpha\cap K_{\beta_n}:n<\omega\}
    \).
    Let \(K_\alpha=\{\delta_{i,j}:i\leq\omega,j<w_i\}\) where
    \(w_i\leq\omega\) for each \(i\leq\omega\),
    \(
      K_\alpha\cap \left(K_{\beta_n}\setminus\bigcup_{m<n}K_{\beta_m}\right)
        =
      \{\delta_{n,j}:j<w_n\}
    \),
    and
    \(
      K_\alpha\setminus\bigcup_{n<\omega}K_{\beta_n}
        =
      \{
        \delta_{\omega,j}:j<w_\omega
      \}
    \).
    Then let \(f_\alpha(\delta_{n,j})=\max(n,f_{\beta_n}(\delta_{n,j}))\) for
    \(n<\omega\) and \(f_\alpha(\delta_{\omega,j})=j\) otherwise.

    We should show that \(f_\alpha\) is finite-to-one. Let \(n<\omega\).
    Since \(f_\alpha(\delta_{m,j})\geq m\), we only consider the finite cases
    where \(m\leq n\). Since each
    \(f_{\beta_m}\) is finite-to-one, \(f_{\beta_m}(\delta_{m,j})\leq n\)
    for only finitely many \(j\). Thus
    \(f_\alpha(\delta_{m,j})=\max(m,f_{\beta_m}(\delta_{m,j}))\) maps to
    \(n\) for only finitely many \(j\).

    We now want to demonstrate that \(f_\alpha\sim f_{\beta_n}\) for all
    \(n<\omega\). Note \(\delta_{m,j}\in K_{\beta_n}\) implies
    \(m\leq n\).
    For \(m=n\), we have
    \(f_\alpha(\delta_{n,j})=\max(n,f_{\beta_n}(\delta_{n,j}))\) which differs
    from \(f_{\beta_n}(\delta_{n,j})\) for only the finitely many \(j\) which
    are mapped below \(n\) by \(f_{\beta_n}\).
    For \(m<n\) and \(\delta_{m,j}\in K_{\beta_n}\),we have
    \(f_\alpha(\delta_{m,j})=\max(m,f_{\beta_m}(\delta_{m,j}))\) which can
    only differ
    from \(f_{\beta_n}(\delta_{m,j})\) for only the finitely many \(j\) which
    are mapped below \(m\) by \(f_{\beta_m}\) or the finitely many \(j\)
    for which the
    almost compatible \(f_{\beta_n}\sim f_{\beta_m}\) differ.

    Finally for any \(\beta<\alpha\), we may conclude \(f_\alpha\sim f_\beta\)
    since there is some \(\beta_n\) with
    \(K_\alpha\cap K_\beta=K_\alpha\cap K_{\beta_n}\),
    \(f_\alpha\sim f_{\beta_n}\), and \(f_{\beta_n}\sim f_\beta\).
  \end{proof}

  We now construct a topology on \(\omega_n\) for each \(n<\omega\) which
  will witness a Kurepa family of size \(\aleph_n\), exploiting a theorem of
  Juh\'asz et. al. in \cite{MR539228}.

  \begin{definition}
    A topological space is said to be \term{\(\omega\)-bounded}
    if each countable subset of the space has compact closure.
    As in \cite{MR539228} we call a
    \(T_2\), locally countable, \(\omega\)-bounded space
    \term{splendid}, and let \(\splendidS\kappa\) represent
    the claim that there exists a splendid space of cardinality \(\kappa\).
  \end{definition}

  \begin{theorem}[\cite{MR539228}]
    \(\splendidS{\aleph_k}\) for \(k<\omega\).
  \end{theorem}

  \begin{lemma}
    The family of compact open sets in a locally countable, \(\omega\)-bounded
    topological space \(X\)
    is a Kurepa family cofinal in \([X]^\omega\). That is,
    \(\splendidS\kappa \Rightarrow \kurepaS\kappa\).
  \end{lemma}

  \begin{proof}
    Let \(\mc K\) collect all compact
    open subsets of \(X\). Of course, every Lindel\"of set in
    a locally countable space is countable, and the closure of every
    countable set is a compact countable set; thus \(\mc K\)
    is cofinal in \([X]^\omega\).
    It is Kurepa since every countable set is contained in a countable
    compact open subspace of \(X\); this subspace has a countable
    base of compact open sets, which closed under finite unions enumerates
    all compact open subsets of the subspace.
  \end{proof}

  \begin{corollary}
    \(\kurepaS{\aleph_k}\) for all \(k<\omega\).
  \end{corollary}

  Alternatively, the previous corollary may be obtained
  via an observation of Todorcevic
  communicated by Dow in \cite{MR1229125}:
  if every Kurepa family of size at most \(\kappa\)
  extends to a cofinal Kurepa family, then the same is true of \(\kappa^+\).

  Nyikos points out in \cite{NYIKOSKUREPA} that a cofinal Kurepa family may
  be used to construct a locally metrizable, \(\omega\)-bounded,
  zero-dimensional space with appropriate cardinality,
  but whether this can be
  strengthened to locally countable and \(\omega\)-bounded (as asked in
  \cite{MR539228}) remains an open question.

  Also left open is this extension of the question asked in \cite{NYIKOSKUREPA}
  and \cite{MR539228} on the possible equivalence of \(\splendidS\kappa\)
  and \(\kurepaS\kappa\).

  \begin{question}
    May any of the implications in the theorem
    \(\splendidS\kappa\Rightarrow\kurepaS\kappa\Rightarrow\alcompSP\kappa\)
    be reversed?
  \end{question}

  Regardless, we have obtained our desired result.

  \begin{corollary}
    \(\alcompSP{\aleph_k}\) for all \(k<\omega\).
  \end{corollary}

  \section{Consistency results}

  As noted in \cite{MR1229125}, Jensen's one-gap two-cardinal theorem under
  \(V=L\) introduced in \cite{MR0309729} implies that \(\kurepaS\kappa\)
  holds for all cardinals \(\kappa\).

  \begin{corollary}[\(V=L\)]\label{alcompSPunderV=L}
    \(\alcompSP{\kappa}\) for all cardinals \(\kappa\).
  \end{corollary}

  As noted earlier, Scheepers proved in \cite{MR1129143}
  that \(\neg\alcompS{\mf c^+}\) is a theorem of ZFC, showing
  \(\alcompSP{\kappa}\) is not equivalent to \(\alcompS{\kappa}\).

  We now demonstrate that \(CH\) is not required to have \(\alcompS{\aleph_2}\)
  fail.

            The forcing extension of a model $M$ by a poset $\mathbb P\in M$ is
            obtained simply by evaluating all $\mathbb P$-names from $M$ by a
            generic filter $G$.
             A set $\tau$ is a $\mathbb P$-name if $\tau$ is a
             (possibly empty)
            set of ordered pairs $(\sigma,p)$ where $p\in \mathbb P$ and
             $\sigma$ is also itself a $\mathbb P$-name.
            If $G$ is a $\mathbb P$-generic filter, then
             $\val_G(\tau) $ is defined to equal
              $\{ \val_G(\sigma) : ~~(\exists p\in G)~~(\sigma,p)\in \tau\}$.

            If $x\in M$, then the canonical
            $\mathbb P$-name, $\check x$, is generally, and recursively,
             taken to be
             $\{ (\check y, 1) : y\in x\}$ where $1$ is the maximum element of
             $\mathbb P$.  However, it will be convenient to consider, when the
            context is clear, $(x,p)$ (for any $p\in \mathbb P$)
            to be a kind of $\mathbb P$-name.
            In particular
            if $\tau \subset X\times \mathbb P$ (for some fixed $X\in M$),
            then we may let $\tau[G] = \{ x : ~~(\exists p\in
            G)~~(x,p)\in\tau\}$.

            Thus,  $\val_G(\tau)$ will denote the recursive evaluation by $G$ and
             $\tau[G]$ will be defined as above.
            In fact, if $\tau\in M$ is any set then each of
             $\val_G(\tau)$ and $\tau[G]$ are well defined.
            It is a standard convention to use a dotted letter, such
            as $\dot x$, to indicate that we are discussing a $\mathbb P$-name.

            One says that a condition $p\in \mathbb P$ forces a statement
            $\varphi$
              to hold, denoted $p\Vdash \varphi$,
            if that statement holds in $M[G]$ for all $\mathbb P$-generic filters
            with $p\in G$. The forcing theorem states that if $M[G]\models
            \varphi$,
             then there is some $p\in G$ forcing that $\varphi$ holds.
            The following is an immediate consequence of the forcing theorem.

            \begin{lemma} If $X\in M$ and $\dot x$ is a $\mathbb P$-name,
            then\label{easyname}
             there is a $\tau\subset X\times \mathbb P$, such that
            for any generic $G$, $\tau[G] = X\cap \val_G(\dot x)$.
            \end{lemma}

            In other words, the family of subsets of any $X\in M$ in
            the extension $M[G]$ is equal to
             $\{ \tau[G] : \tau \subset X\times \mathbb P, \ \
            \tau\in M\ \}$.  We will be using
            the forcing poset $\operatorname{Fn}(\omega_2, 2)$.
            The elements of this poset are all the finite partial functions
            from $\omega_2$ into $2$ ordered by reverse inclusion.
            It follows that, for any $\lambda\in \omega_2$,
             each of
             $\operatorname{Fn}(\lambda,2)$ and $\operatorname{Fn}(\omega_2\setminus
             \lambda,2)$ are subposets. For any
             $\operatorname{Fn}(\omega_2,2)$-generic
            filter  $G$, it easily follows that $G_\lambda = G\cap
            \operatorname{Fn}(\lambda,2)$ and
             $G^\lambda = G\cap
            \operatorname{Fn}(\omega_2\setminus
             \lambda,2)$ are also generic filters. But a much stronger statement is
             true.

            \begin{lemma}\cite{MR597342}
            Assume that $G\subset \operatorname{Fn}(\omega_2,2)$ is a
             generic\label{factor} filter,
             and let $\lambda\in \omega_2$. Then the final model $M[G]$
            is equal to $(M[G_\lambda])[G^\lambda]$ in the sense that
            $G^\lambda$ is a $\operatorname{Fn}(\omega_2\setminus \lambda,
             2)$-generic filter   over the model $M[G_\lambda]$.

            In addition, for each $X\in M$ and name $\dot A\subset
            X\times \operatorname{Fn}(\omega_2,2)$, we get that
             \[ (\dot A(G_\lambda))[G^\lambda] = \dot A [G]\ \ \mbox{where}\
            \dot A(G_\lambda) =
             \{ (x,p\restriction [\lambda,\omega_2))  :
             (x,p)\in \dot A\ \ \mbox{and} \ \ p\restriction \lambda\in G_\lambda\}
            \]
            \end{lemma}

            With these lemmas in hand we are ready to prove the theorem. The idea
            of the proof comes from  Kunen's result about no $\omega_2$ length mod
            finite chains of subsets of $\omega$. We consider any  family of
            names  of suitable one-to-one functions from countable subsets of
            $\omega_2$ into $\omega$. We identify a large enough $\lambda\in \omega_2$
            so that a pattern has emerged and we pass to the model $M[G_\lambda]$. We
            then show that this pattern can not continue out to $\omega_2$.

            \begin{theorem} There exists a model of ZFC for which
             $\mathfrak c=\aleph_2$ and $\lnot \mathcal A(\aleph_2)$.
            \end{theorem}

            \begin{proof}
            We start with a model $M$ of GCH and suppose that $G$ is a
            $\operatorname{Fn}(\omega_2,2)$-generic filter. The argument takes
            place in $M$.
            Let $\{ \dot f_{A} : A\in [\omega_2]^\omega\}$ be a family of
            names (in $M$)
             such that, for any generic $G$ and each $A\in [\omega_2]^\omega\cap
             M$,
             $\dot f_{A}[G]$ is a one-to-one function from $A$ into
            $\omega$.  We also assume that whenever $B\subset A$ are members of
            $[\omega_2]^\omega$, we have that $\dot f_{B}[G]\subset^* \dot
            f_{A}[G]$. If we now obtain a contradiction
            then we will have shown that $\mathcal A(\aleph_2)$ fails.

            By \cite[1.5]{MR1031969}, there is a set $H\subset H(\aleph_3)$
            such that the family $\{ \dot f_{A} : A\in [\omega_2]^{\omega}\}$ is an
            element of $H$,  $H$ is an elementary submodel of
            $ H(\aleph_3)$, $H$ has cardinality
            $\aleph_1$, and $H^\omega\subset H$ (every
             countable subset of $H$ is an element of $H$).

            Let $\lambda = H\cap \omega_2$ (same as the supremum of $H\cap
            \omega_2$). Consider the name  $\dot f_{[\lambda,\lambda+\omega)}$.
            What is such a name?  By Lemma \ref{easyname},
            we can assume that it is a set of pairs
             of the form $( (\lambda+k,m), p)$ where $p\in \mathop{Fn}(\omega_2,
             2)$ and, of course, $k,m\in \omega$.
            Furthermore,  for each $k,m$ it is enough
            (see \cite[5.11,5.12]{MR597342})
            to take a  countable set of such $p$ to get an equivalent
            (\textit{nice}\/) name.
            Given any such nice name $\dot f$, let $\operatorname{supp}(\dot f)$
            denote the union of the domains of conditions $p$ appearing in the
            name.

             Now let $Y$ equal $\operatorname{supp}(\dot
             f_{[\lambda,\lambda+\omega)})\setminus \lambda$. Furthermore,
             fix any $\mu\in \lambda\subset H$ such that
               $\operatorname{supp}(\dot
             f_{[\lambda,\lambda+\omega)})\cap \lambda$ is contained in $\mu$.
             Let $\delta\in \omega_1$ denote the order type of
             $Y$ and let  $\varphi_{\mu,\lambda}$ be the order-preserving function
             from  $\mu\cup Y$ onto  the ordinal $\mu+\delta$.  This lifts
             canonically to
            an order-preserving bijection $\varphi_{\mu,\lambda}:
            \operatorname{Fn}(\mu\cup Y,2) \mapsto
            \operatorname{Fn}(\mu+\delta,2)$.
            We can similarly make sense of the name
             $\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$, call it $F_H$.
            Here simply, for each tuple $(\,(\lambda+k,m), p)\in \dot
            f_{[\lambda,\lambda+\omega)}$,
             we have that $(\,(\mu+k,m),\varphi_{\mu,\lambda}(p))$ is in $F_H$.
            Again, let $\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$
            be interpreted in the above sense as giving $F_H$ (which is an element
            of $H$).

             Other values
            replacing $\lambda>\mu$ will result in their own set $Y$
            and canonical map $\varphi_{\mu,\lambda}$.
            Now the object $F_H$ is an
            element of $H$, and $H$ believes this statement is true:
            $$
            (\forall \beta\in\omega_2)~ (\exists \lambda\in \omega_2\setminus \beta)~~~
            \operatorname{supp}(\dot f_{[\lambda,\lambda+\omega)})
            \cap \lambda\subset \mu \ \ \mbox{and}\ \
            F_H = \varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$$

            But now, this  means that,
            not only is there  an $\alpha\in H$,
            $ F_H = \varphi_{\mu,\alpha}(\dot f_{[\alpha,\alpha+\omega)})$
            but also that there is an increasing sequence
             $\{\alpha_\xi : \xi  \in \omega_1\}\subset \lambda$ of
             such $\alpha$'s satisfying that, for each $\xi$
            we have that $\operatorname{supp}(\dot
            f_{[\alpha_\xi,\alpha_\xi+\omega)})$ is contained in $\alpha_{\xi+1}$.

            Choose such a sequence.
             This means that if we let $A = \bigcup_{n>0} [\alpha_n,\alpha_n+\omega)$ we
             have the name $\dot f_A$ in $H$. This then means that all
             the $( (\beta,m) , p)$ appearing in (the nice name)
            $\dot f_A$ have the property
             that $\mathop{dom}(p)$ is contained in $H$.
            There is, also within $H$, a name $\dot g$ satisfying that
             $\dot f_A(\alpha_n+k) = \dot f_{[\alpha_n,\alpha_n+\omega)}(\alpha_n+k)$
            for all $k >\dot g(n)$,
            or more precisely,
             $\dot g\subset (\omega\times\omega) \times
             \operatorname{Fn}(\omega_2,2)$ satisfies that
             $\dot g[G]\in \omega^\omega$ and
              $\dot f_{A}[G](\alpha_n+k) = \dot
              f_{[\alpha_n,\alpha_n+\omega)}[G](\alpha+k)$
            for all $k>\dot g[G](n)$.

            We now apply  Lemma \ref{factor} and we are now
            working in the extension $M[G_\mu]$.
            We work for a contradiction. Something special has now happened,
             namely,  the supports of the names $\{
            \dot f_{[\alpha_n, \alpha_n+\omega)} (G_\mu):  0< n<\omega\}$ are pairwise disjoint
            and also disjoint from the support of the name
             $\dot f_{[\lambda,\lambda+\omega)}(G_\mu)$.
             And not only that, these names are pairwise isomorphic (in
             the way that they all map to $F_H$).

             Since $A$ is disjoint from $[\lambda,\lambda+\omega)$,
            there must be an integer $\ell$
            together with a condition
             $q\in \mathop{Fn}(\omega_2\setminus \mu,2)$ satisfying that for all
             $n>\ell$ , $q$ forces that

            \centerline{
             ``if $k>\dot g(n)$ then
             $(\dot f_{[\alpha_n,\alpha_n+\omega)}(G_\mu))(\alpha_n+k) \neq
            (\dot f_{[\lambda,\lambda+\omega)}(G_\mu))(\lambda +k)$''.}

            Choose $n>\ell$ large enough so that $\mathop{dom}(q) \cap [\alpha_n,
            \alpha_{n+1})$ is
            empty.
             Choose $q_1<q\restriction \lambda$ (in $H$) so that
            $$ \varphi_{\mu,\alpha_n}(q_1\restriction \operatorname{supp}(
             \dot f_{[\alpha_n,\alpha_n+\omega)}) =
            \varphi_{\mu,\lambda}(q\restriction \operatorname{supp}(
             \dot f_{[\lambda,\lambda+\omega)}) $$
            and then (again in $H$) choose $q_2 < q_1$
            so that it both forces a value
             $L$ on $\ell+\dot g(n)$
            and subsequently forces a value $m$ on
             $\dot f_{[\alpha_n, \alpha_n+\omega)}(\alpha_n+L+1)$.
            But now, again calculate
            $$ q_3 = \varphi_{\mu,\lambda}^{-1} \circ \varphi_{\mu,\alpha_n}
            (q_2\restriction \operatorname{supp}(\dot
            f_{[\alpha_n,\alpha_n+\omega)}))$$
            and, by the isomorphisms, we have that $q_3$ forces that
             $\dot f_{[\lambda,\lambda+\omega)}(\lambda+L+1) = m$.

            Technically (or with more care)  all of this is taking place in the
            poset $\operatorname{Fn}(\omega_2\setminus \mu,2)$ and this means
            that $q_3$ and $q$ are  with each other.
            To verify this it suffices to consider
             $q(\beta)=e$ and to assume that $q_3(\beta)$ is defined.
            Since $q_3(\beta)$ is defined, we have that
            there is a  $\beta'\in\mathop{dom}(q_2)$ such that $
             \varphi_{\mu,\lambda}(\beta) = \varphi_{\mu,\alpha_n}(\beta')$,
            and that $q_3(\beta) = q_2(\beta')$.
            But, by definition of $q_1$, $\beta'\in\mathop{dom}(q_1)$
            and even  that $q_1(\beta') = q(\beta)$.
             Then, since $q_2<q_1$, we have that $q_2(\beta')=q_1(\beta') =
             q(\beta)$. This completes the circle that $q_3(\beta) = q(\beta)$.

            Finally, our contradiction is that $q_3\cup q_2\cup q$
            forces that
             $k=L+1$ violates the quoted statement above.
            \end{proof}

  We are also able to force \(\alcompSP{\kappa}\) to fail for every cardinal
  other than the first \(\omega\)-many we've already guaranteed.

            \begin{theorem} It follows from the existence of a $2$-huge cardinal
            that there is a model of $ZFC$ for which $\lnot \mathcal A'(\aleph_{\omega})$.
            \end{theorem}

            \begin{proof}
            We will need the model constructed in \cite{MR1045371} in
             which an instance of Chang's conjecture
             $(\aleph_{\omega+1},\aleph_\omega) \rlap{$\rightarrow$}\,{\rightarrow}
             (\aleph_1,\aleph_0)$ is shown to hold.

            We can take as a given (as shown in \cite[Theorem 5]{MR1045371}) that we may
            assume that we have a model $V$
            of GCH in which there are regular limit cardinals $\kappa<\lambda$
            satisfying that
            $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
            (\kappa^{+\omega+1},\kappa^{+\omega})$.

            What this says is that if $L$ is a countable language
            with at least one unary relation symbol $R$ and
             $M$ is a model of $L$  with base set $\lambda^{+\omega+1}$
            in which the interpretation of $R$ has cardinality
             $\lambda^{+\omega}$, then $M$ has an elementary submodel
             $N$ of cardinality $\kappa^{+\omega+1}$ in which
            $R\cap N$ has cardinality $\kappa^{+\omega}$ (of course
             $R\cap N$ is the interpretation of $R$ in $N$ because
            $N\prec M$).

            The interested reader will want to know that it is shown in
            \cite{MR1045371} that if $\kappa$ is a 2-huge cardinal
            and $j$ is the 2-huge
            embedding with critical point $\kappa$, then with
             $\lambda = j(\kappa)$ one has that
            $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
            (\kappa^{+\omega+1},\kappa^{+\omega})$ holds. There is no loss of
            generality to  also assume that GCH holds in this model.

            Let $\{ h_\xi : \xi \in \lambda^{+\omega+1}\}$
            be a scale in $\Pi\{ \lambda^{+n+1} : n\in \omega\}$
             ordered by the usual mod finite coordinatewise
            ordering. For convenience we may assume that $h_\xi(n) \geq
            \lambda^{+n}$ for all $\xi$ and all $n$. For each integer
             $m$ the cofinality of the mod finite ordering on
             $\Pi\{ \lambda^{+n+1} : m< n\in \omega\}$
            is the same as it is for the entire product
             $\Pi\{ \lambda^{+n+1} : n\in \omega\}$. If $P$
            is any poset of cardinality less than
             $\lambda^{+m}$ then, in the forcing extension by $P$,
            every function in
             $\Pi\{ \lambda^{+n+1} : m< n\in \omega\}$ is bounded
            above by a ground model function.
            It therefore
            follows easily that  in the forcing
            extension by $P$,  the sequence
            $\{ h_\xi : \xi \in \lambda^{+\omega+1}\}$  remains cofinal in
             $\Pi\{ \lambda^{+n+1} : n\in \omega\}$.

            The forcing notion $\mathbb P_0$ is simply the finite condition
            collapse of $\kappa^{+\omega}$, i.e. $\mathbb P_0 = \left(
             \kappa^{+\omega}\right)^{<\omega}$. In the forcing extension by
             $\mathbb P_0$, one now has that the ordinal $\kappa^{+\omega+1}$
             from $V$ is the first uncountable cardinal $\aleph_1$. Then in this
             forcing extension we let $\mathbb P_1$ be the countable condition
             Levy collapse, $Lv(\lambda,\omega_2)$, which collapses all cardinals
             less than $\lambda$ to have cardinality at most $\aleph_1$. The poset
             $\mathbb P_1$ has cardinality $\lambda$. We treat
            $\mathbb P_0*\mathbb P_1$ as containing $\mathbb P_0$
            as a subposet by identifying each $(p_0,1)$ with $p_0$.
            After forcing with $\mathbb P_0*\mathbb P_1$ we will have
            that $\omega_1$ is the ordinal $\left(\kappa^{+\omega+1}\right)^V$,
             $\omega_2$ is the ordinal $\lambda$, and $\omega_\omega$ is the
             ordinal $\left(\lambda^{+\omega}\right)^V$.

            Now we assume that we have an assignment $\dot f_{\dot A}$ of a
             $\mathbb P_0*\mathbb P_1$-name of a finite-to-one function from
             $\dot A$ into $\omega$ for each
             $\mathbb P_0*\mathbb P_1$-name of a countable subset of
             $\lambda^{+\omega+1}$. We will obtain a contradiction
             to the claim of coherence.

            Let $\{ \dot A_\xi : \xi \in \lambda^{+\omega+1}\}$ be an enumeration
            of all the  nice $\mathbb P_0$-names of countable
            subsets of $\lambda^{+\omega}$. For each $\xi\in \lambda^{+\omega+1}$,
             let $\dot f_\xi$ be another notation for $\dot f_{\dot A_\xi}$.
            Since $\mathbb P_0$ forces that $\mathbb P_1$ is countably closed,
            the collection of all nice  $\mathbb P_0$-names will produce all
            the countable sets in the extension by $\mathbb P_0* \mathbb
            P_1$, but $\mathbb P_0*\mathbb P_1$ can introduce new enumerations of
            these names.
            For each $\xi\in \lambda^{+\omega+1}$,
            there is a minimal $\zeta_\xi$ so that $\dot A_{\zeta_\xi}$ is the
            canonical name for the range of $h_\xi$. This means that
             $\dot f_{\zeta_\xi} \circ h_\xi$ is simply the
             $\mathbb P_0 *\mathbb P_1$-name of a finite-to-one function from
             $\omega$ to $\omega$.
              For each $\xi\in \lambda^{+\omega+1}$,
             choose any $p_\xi \in \mathbb P_0*\mathbb P_1$ so that
             there is a nice $\mathbb P_0$-name, $\dot H_\xi$,
             that is forced by
             $p_\xi$  to equal $\dot f_{\zeta_\xi}\circ h_\xi$.
              Choose $\Lambda\subset \lambda^{+\omega+1}$
              of cardinality $\lambda^{+\omega+1}$
              and so that there is a pair $p,\dot H$
              satisfying that $p_\xi = p$ and $\dot H_\xi = \dot H$
              for all $\xi\in \Lambda$. We may assume that $p$
              is in a generic filter $  G$.

            Let $\{ x_\xi : \xi\in \lambda^{+\omega+1}\}$ be any enumeration of
             $H(\lambda^{+\omega+1})$  such that
             $\{ x_\xi : \xi\in \lambda^{+\omega}\}$ is also equal to
             $H(\lambda^{+\omega})$. We choose this enumeration in such a way that
             $x_\xi\in x_\eta$ implies $\xi<\eta$.
            We use relation symbol $R_0$ to code
            (and well order)
             $(H(\lambda^{+\omega+1}), \in)$ as follows: $(\xi,\eta)\in R_0$
            if and only if $x_\xi\in x_\eta$.
            Let $R_1$ be a binary relation on $\kappa^{+\omega}$ so that
            $(\kappa^{+\omega},R_1)$ is isomorphic to $\mathbb P_0$. Let
             $R_2$ be a binary relation on $\lambda$ so that
             $R_2\cap (\kappa^{+\omega}\times \kappa^{+\omega})=R_1$ and
             $(\lambda,R_2)$ is isomorphic to $\mathbb P_0*\mathbb P_1$.
            Let $\psi$ be the poset isomorphism from $(\lambda,R_2)$ to
             $\mathbb P_0*\mathbb P_1$.

            We continue coding. We can code the sequence
             $\{ h_\xi : \xi\in \lambda^{+\omega+1}\}$ as another binary  relation
             $R_3$ on $\lambda^{+\omega+1}$ where $R_3\cap \left(\{\xi\}\times
             \lambda^{+\omega+1}\right) = \{ (\xi,h_\xi(n) ) : n\in \omega\}$
            for each $\xi\in \lambda^{+\omega+1}$. The relation symbol $R_4$ can
            code the sequence $\{ \dot A_\xi : \xi \in \lambda^{+\omega+1}\}$
            where $(\xi, \alpha, \zeta) \in R_4$ if and only if
            $(\check \alpha, \psi(\zeta))$ is in the name $\dot A_\xi$.
             Let $R_5$ code this collection, i.e.
             $(\gamma,n,m,\eta)\in R_5$ if and only if
             $(\check{(n,m)}, \psi(\eta))\in \dot H_\gamma$. Also let $R_6$ code
             (equal)  the set $\Lambda$.
            Finally we use the relation symbol $R_7$ to similarly code the
            sequence $\{ \dot f_\xi  : \xi \in \lambda^{+\omega+1}\}$:
             $(\xi, \alpha, n, \zeta) \in R_7$ if and only if
            $(\check {(\alpha,n)}, \psi(\zeta))$ is in the name $\dot f_\xi$.

            Needless to say, the unary relation symbol $R$
            is interpreted as the set $\lambda^{+\omega}$ for the application of
            $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
            (\kappa^{+\omega+1},\kappa^{+\omega})$.
             Now  we have defined our model
            $M$  of the language $L= \{\in, R, R_0,\ldots, R_7\}$,
             and we choose an elementary submodel
            $N$ witnessing
            $(\lambda^{+\omega+1},\lambda^{+\omega})\rlap{$\rightarrow$}\,{\rightarrow}
            (\kappa^{+\omega+1},\kappa^{+\omega})$. Of course $N$ is really just
            a $\kappa^{+\omega+1}$ sized subset of $\lambda^{+\omega+1}$
            with the additional property that $N\cap \lambda^{+\omega}$ has
            cardinality $\kappa^{+\omega}$.  In the
            forcing extension $N$ has cardinality $\omega_1$ and
             $A= N\cap \lambda^{+\omega}$ is countable.

            We will need the following claim from \cite{MR1045371}:

            \begin{claim*} We may assume that $N$ satisfies that
             $N\cap \kappa^{+\omega+1}$ is transitive (i.e. an initial segment).
            \end{claim*}

            \bgroup
            \def\proofname{Proof of Claim:\/}

            \begin{proof}
            Suppose  our originally supplied $N$ fails the conclusion of the
            claim.
            We   know that $\kappa^{+\omega}\in N$,
            (  via $R_1$)
            in which case so is $\kappa^{+\omega+1}$.

            Then set $\beta_0 = \sup(N\cap \kappa^{+\omega+1})$ and consider the
            Skolem closure $\mathop{Hull}(N\cup \beta_0, M)$.  A little informally
            (in that we have to formalize the enumeration of formulas as per
            G\"odel coding)
            let $\{\varphi_n : n\in \omega\}$ be an enumeration of all formulas
            in the language $L$, and let  $\ell_n$ be the minimal integer such
            that the free variables of $\varphi_n$ are among
             $\{ v_0, \ldots, v_{\ell_n}\}$.
            Then,   for each tuple
            $\langle \xi_1,\ldots, \xi_{\ell_n} \rangle$
            of elements of $\lambda^{+\omega+1}$,
             we define $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be
             the minimal $\xi_0\in \lambda^{+\omega+1}$ such that
             $M\models \varphi_n(\xi_0,\ldots, \xi_{\ell_n})$.
            If there is no such $\xi_0$,
               in other words if $M\models \lnot\exists x~~  \varphi_n(x,\xi_1,
            \ldots,\xi_{\ell_n})$, then set
             $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be $0$.
            Now $\mathop{Hull}(N\cup \beta_0,M)$ is just the minimal superset
             $X$
            of $N\cup \beta_0$ that satisfies that
             $f_n[X^{\{1,\ldots,\ell_n\}}] \subset X$ for all $n$. Since this is
             simply a large algebra, we can generate all the terms $t$ of
            the algebraic operations $\{ f_n : n\in \omega\}$.
            It is easily seen that for each $\zeta \in X$, there is a term
            $t(v_1,\ldots, v_m)$ such that $\zeta  = t(\delta_1,\ldots, \delta_m)$
             for some sequence $\langle \delta_1,\ldots, \delta_m\rangle$ with
             each $\delta_i\in  N\cup \beta_0$.
             Assume that
              $\zeta\in \kappa^{+\omega+1}$. By
              re-indexing the variables in the term we can assume
              that there is an $n\leq m$ so that
               $\delta_i  <\beta_0$ for $1\leq i\leq n$
                and $\kappa^{+\omega+1}\leq \delta_i$ for $n<i\leq m$.
                Let $\vec a $ denote
                the tuple $\langle \delta_{n+1},\ldots, \delta_m\rangle$.
              Choose $\eta\in N\cap \kappa^{+\omega+1}$ large enough so that
             $ \{  \delta_1, \ldots, \delta_n\}$ is contained in  $\eta$. Since set-membership
             in $M$ is coded by $R_0$ rather than $\in$ we have to argue a little
             less naturally.
             Consider the set
               $s_0(\eta,\vec a) = \{ t(\gamma_1,\ldots, \gamma_n,\vec a) :
            \{\gamma_1,\cdots,\gamma_n\} \in [\eta]^{\leq n}\}$. Clearly
             $s_0(\eta,\vec a)$ is a member of $H(\lambda^{+\omega+1})$.
             Now define $s_1(\eta,\vec a) $ to be
             $\{ x_\alpha : \alpha\in s_0(\eta, \vec a)\}$,
             and choose the
             unique $\zeta_1\in \lambda^{+\omega+1}$ such
             that $x_{\zeta_1} = s_1(\eta,\vec a)$. We claim that $\zeta_1\in N$.
             Note that $\alpha R_0\zeta_1$ holds if and only if
              $ \alpha \in s_0(\eta,\vec a)$, and therefore
             $$M\models (\forall \alpha)\left[\alpha R_0 \zeta_1 ~~\mbox{iff}~~
             (\exists   \gamma_1\in \eta)\cdots(\exists \gamma_n\in\eta)(\alpha=
             t(\gamma_1,\ldots, \gamma_n,\vec a))\right]~.$$
             By elementarity then we have that $\zeta_1\in N$,
             and by similar reasoning the supremum, $\zeta_0$,
            of $\zeta_1\cap \kappa^{+\omega+1}$ is also in $N$.
               This of course means that $\zeta < \beta_0$.
            \end{proof}

            \egroup

            We use the elementarity of
             $N$  to deduce properties of the families
             $\{ \dot A_\xi : \xi\in N\}$ and $\{ \dot f_\xi : \xi \in
             N\}$.  Actually the collection we are most interested in
             is the family $\{ h_\xi : \xi\in \Lambda\cap N\}$.

            Now we need a result from Shelah's pcf theory which
            is proven in  Jech \cite[24.9]{MR1940513}.
             Since  $\aleph_1=\mathfrak c < \kappa^{+\omega+1}$ there
              is a function $\langle \varrho_n : n\in \omega\rangle$
              in $\Pi_n \lambda^{+\omega}   $ such that the sequence
               $\{ h_\xi : \xi \in N\}$ is unbounded mod finite
               in $\Pi_n \varrho_n$
               For each $n$, $\varrho_n \leq \sup(N\cap \lambda^{+n+2})$.
                  Since $\mathbb P_0$ has cardinality $\kappa^{+\omega}$,
            and so less than $|N|=\kappa^{+\omega+1}$, a standard argument
            (analogous to the fact that adding a Cohen real does
             not add a dominating real)
             shows that    the sequence $\{ h_\xi : \xi \in \Lambda\cap N\}$ remains
             unbounded
                mod finite in $\Pi_n \varrho_n$ (and in
                 $\Pi_n (\varrho_n\cap N)$).

            Now pass to the extension by
                  $G\cap \mathbb P_0$ and let $H$ be the function
                   $\operatorname{val}_{G}(\dot H)$, and we recall
                   that $f_{\zeta_\xi}(h_\xi(n)) = H(n)$ for all $n\in \omega$
            and $\xi\in \Lambda$.
             Now pass to the full extension $V[G]$ and again, since
              $\mathbb P_1$ was forced to be countably closed,
               the family $\{ h_\xi : \xi \in \Lambda\cap N\}$ is still
               unbounded in $\Pi_n (\varrho_n \cap N)$ (no new elements
            were added). We let
               $A$ be the countable set $N\cap \lambda^{+\omega}$,
               and for  each
               $\xi\in \Lambda\cap N$, there is an $n_\xi$ such that
                $f_{\xi}(h_\xi(m)) = f_A(h_\xi(m))$ for all $m>n_\xi$.
                There is a single $n$ so that $\Lambda_n
                 = \{\xi\in \Lambda\cap N       : n_\xi = n\}$ has cardinality
                  $\omega_1 $, and thus
                  $\{ h_\xi : \xi\in \Lambda_n\cap N\}$ is also unbounded
                  in $\Pi_n (\rho_n\cap N)$.
            This certainly implies that there is an $m>n$
            such that $\{ h_\xi (m) : \xi\in \Lambda_n\cap N\}$ is infinite.
            This completes the proof since  $f_{A}(h_\xi(m)) = H(m)$
            for all $\xi\in \Lambda_n\cap N$.
            \end{proof}

  \section{Applications to infinite length games}

  We introduce three variations of Scheeper's game which we defined in
  the introduction.

  \begin{game}
    Let \(\schFillGame\kappa\) denote the
    \term{Scheepers countable-finite union game} which proceeds analogously
    to \(\schFillStrictGame\kappa\), except that \(\pl C\)'s restriction in round \(n+1\)
    is weakened to \(C_{n+1}\supseteq C_n\).
  \end{game}

  \begin{game}
    Let \(\schFillInitialGame\kappa\) denote the
    \term{Scheepers countable-finite initial game} which proceeds analogously
    to \(\schFillGame\kappa\), except that \(\pl F\)'s winning condition is
    weakened to \(\bigcup_{n<\omega}F_n\supseteq C_0\).
  \end{game}

  \begin{game}
    Let \(\schFillIntGame\kappa\) denote the
    \term{Scheepers countable-finite intersection game} which proceeds analogously
    to \(\schFillInitialGame\kappa\), except that \(\pl C\) may choose any
    \(C_n\in[\kappa]^{\leq\omega}\) each round, and \(\pl F\)'s winning condition
    is weakened to
    \(\bigcup_{n<\omega}F_n\supseteq\bigcap_{n<\omega}C_n\).
  \end{game}

  In \cite{clontzMengerGamePreprint} Clontz extended Scheepers' application of
  almost-compatible injections to these game variants as well as
  \(\menGame{\oneptlind\kappa}\). However, when considering Markov strategies,
  finite-to-one functions suffice.

  \begin{figure}[ht]
\begin{center}
\begin{tikzpicture}[x=0.8in,y=0.5in]
  \node at (-3.5,2) {\(\alcompS\kappa\)};
  \node at (-3.5,0) {\(\pl F \ktactwin2\schFillStrictGame\kappa\)};
  \node at (0,2) {\(\alcompSP\kappa\)};
  \node at (-1,1) {\(\pl F \kmarkwin2\menGame{\oneptlind\kappa}\)};
  \node at (0,1) {\(\Leftrightarrow\)};
  \node at (1,1) {\(\pl F \kmarkwin2\schFillIntGame\kappa\)};
  \node at (-1,0) {\(\pl F \kmarkwin2\schFillInitialGame\kappa\)};
  \node at (0,0) {\(\Leftrightarrow\)};
  \node at (1,0) {\(\pl F \kmarkwin2\schFillGame\kappa\)};
  \node at (0,-1) {\(\pl F \kmarkwin2\schFillStrictGame\kappa\)};
  \draw[-implies,double equal sign distance] (-3.2,2) -- (-0.3,2);
  \draw[-implies,double equal sign distance] (-2.6,0) -- (-1.9,0);
  \draw[-implies,double equal sign distance] (-3.5,1.65) -- (-3.5,0.35);
  \draw[-implies,double equal sign distance] (0,1.65) -- (0,1.35);
  \draw[-implies,double equal sign distance] (0,0.65) -- (0,0.35);
  \draw[-implies,double equal sign distance] (0,-0.35) -- (0,-0.65);
\end{tikzpicture}
\end{center}
\caption{Diagram of Scheeper/Menger game implications with \(\alcompS{\kappa},\alcompS{\kappa}\)}
\label{GamesDiagram}
\end{figure}

  \begin{theorem}
    Figure \ref{GamesDiagram}.
  \end{theorem}

  \begin{proof}
    \(\alcompS\kappa\Rightarrow\pl F \ktactwin2\schFillStrictGame\kappa\) was
    shown by Scheepers in \cite{MR1129143} (also, see the following theorem).
    Most of the other results in the figure were proven
    in \cite{clontzMengerGamePreprint}, with the exception that
    \(\alcompSP\kappa\) was not considered at the time.
    The following proof that
    \(\alcompSP\kappa\Rightarrow\pl F \kmarkwin2\schFillIntGame\kappa\)
    is a trivial modification of the proof presented in
    \cite{clontzMengerGamePreprint} assuming \(\alcompS\kappa\), but as
    that paper is under review at the time
    of this writing, we provide it here.

    Let \(f_A\) for \(A\in[\kappa]^{\leq\omega}\) witness \(\alcompSP\kappa\).
    We define a \(2\)-mark \(\sigma\) for \(\schFillIntGame\kappa\) as follows:
      \[
        \sigma(\<A\>,0) = \{\alpha\in A: f_A(\alpha) = 0 \}
      \]
      \[
        \sigma(\<A,B\>,n+1)
          =
        \{
          \alpha \in A\cap B
        :
          f_B(\alpha) \leq n+1 \text{ or }
          f_A(\alpha)\not=f_B(\alpha)
        \}
      \]
    For any attack \(\<A_0,A_1,\dots\>\) by \(\pl C\) and
    \(\alpha\in\bigcap_{n<\omega}A_n\), either \(f_{A_n}(\alpha)\) is constant for
    all \(n\), or \(f_{A_n}(\alpha)\not=f_{A_{n+1}}(\alpha)\) for some \(n\);
    either way, \(\alpha\) is covered.
  \end{proof}

  We include the following proof from \cite{MR1129143} to point
  out why \(\alcompSP\kappa\) seems insufficient for providing \(\pl F\) a
  winning \(2\)-tactic in \(\schFillStrictGame\kappa\), despite that it
  witnesses a winning \(2\)-mark.

  \begin{theorem}[\cite{MR1129143}]
    \(\alcompS\kappa\Rightarrow\pl F \ktactwin2\schFillStrictGame\kappa\)
  \end{theorem}

  \begin{proof}
    Let \(\{f_A:A\in[\kappa]^{\leq\aleph_0}\}\) witness
    \(\alcompS\kappa\), and define \(g_A:A\to\omega\) by
    \(g_A(\alpha)=f_A(\alpha)-|\{\beta\in A:f_A(\beta)<f_A(\alpha)\}|\).

    We claim that \(\{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}\)
    must be finite as it is bounded above by
    \(\max\{M,f_A(\alpha),f_B(\alpha):f_A(\alpha)\not=f_B(\alpha)\}\)
    where \(M=f_B(\alpha)\) for some \(\alpha\in B\setminus A\).
    To see this, let \(f_A(\alpha)=f_B(\alpha)=N>M\) and assume
    \(f_A(\beta)\not=f_B(\beta)\) implies \(f_A(\beta),f_B(\beta)<N\).
    Then
    \[
      g_A(\alpha)
        =
      N-|\{\beta\in A:f_A(\beta)<N\}|
        >
      N-|\{\beta\in B:f_B(\beta)<N\}|
        =
      g_B(\alpha)
    \]
    with the strictness of the inequality witnessed by \(f_B(\alpha)=M<N\)
    for some \(\alpha\in B\setminus A\).

    As a result,
    \[
      \sigma(\<A,B\>)
        =
      \{\alpha\in A:g_A(\alpha)\leq g_B(\alpha)\}
    \]
    is a legal \(2\)-tactic for \(\pl F\). Let \(C=\<C_0,C_1,\dots\>\) be
    a strictly increasing sequence of countable sets and
    \(\alpha\in C_n\).

    Noting that \(f_A\) is an injection
    (not just finite-to-one), \(0\leq g_{C_{n+m}}(\alpha)\) for all
    \(m<\omega\), and it follows that
    \(g_{C_{n+m}}(\alpha)\leq g_{C_{n+m+1}}(\alpha)\) for some \(m<\omega\).
    Therefore \(\alpha\in\sigma(\<C_{n+m},C_{n+m+1}\>)\).
  \end{proof}

  While the above proof cannot be trivially modified to utilize
  the finite-to-one functions witnessed by
  \(\alcompSP\kappa\) in constructing a winning \(2\)-tactical strategy
  for \(\schFillStrictGame{\kappa}\), whether
  \(\alcompSP\kappa\) is sufficient for
  \(\pl F\ktactwin2\schFillStrictGame{\kappa}\) after all
  does remain open:

  \begin{question}
    May the previous theorem be improved by replacing \(\alcompS\kappa\)
    with \(\alcompSP\kappa\)?
  \end{question}

  We would like to demonstrate that \(\alcompSP\kappa\) is not
  necessary for constructing winning \(2\)-Markov strategies in
  \(\schFillIntGame\kappa\).

  \begin{theorem}
    Let \(\alpha\) be the limit of increasing ordinals \(\beta_n\) for \(n<\omega\).
    If \(\pl F \kmarkwin2\schFillIntGame{\aleph_{\beta_n}}\) for all
    \(n<\omega\), then \(\pl F\kmarkwin2\schFillIntGame{\aleph_\alpha}\).
  \end{theorem}

  \begin{proof}
    Let \(\sigma_n\) be a winning \(2\)-mark for \(\pl F\) in
    \(\schFillIntGame{\aleph_{\beta_n}}\). Define the \(2\)-mark \(\sigma\)
    for \(\pl F\) in \(\schFillIntGame{\aleph_\alpha}\) as follows:
    \[
      \sigma(\<C\>,0)
        =
      \sigma_0(\<C\cap\aleph_{\beta_0}\>,0)
    \]
    \[
      \sigma(\<C,D\>,n+1)
        =
      \sigma_{n+1}(\<D\cap\aleph_{\beta_{n+1}}\>,0)
        \cup
      \bigcup_{m\leq n}
      \sigma_m(\<C\cap\aleph_{\beta_m},D\cap\aleph_{\beta_m}\>,n-m+1)
    \]

    Let \(\<C_0,C_1,\dots\>\) be an attack by
    \(\pl C\) in \(\schFillIntGame{\aleph_\alpha}\), and
    \(\alpha\in\bigcap_{n<\omega}C_n\).
    Choose \(N<\omega\) with \(\alpha<\aleph_{\beta_{N+1}}\). Consider the
    attack
    \(\<C_{N+1}\cap\aleph_{\beta_{N+1}},C_{N+2}\cap\aleph_{\beta_{N+1}},\dots\>\)
    by \(\pl C\) in \(\schFillIntGame{\aleph_{\beta_{N+1}}}\). Since
    \(\sigma_{N+1}\) is a winning \(2\)-mark and
    \(\alpha\in\bigcap_{n<\omega}C_{N+n+1}\cap\aleph_{\beta_{N+1}}\), either
    \(\alpha\in\sigma_{N+1}(\<C_{N+1}\cap\aleph_{\beta_{N+1}}\>,0)\) and thus
    \(\alpha\in\sigma(\<C_N,C_{N+1}\>,N+1)\), or
    \(
      \alpha
        \in
      \sigma_{N+1}(
        \<C_{N+M+1}\cap\aleph_{\beta_{N+1}},C_{N+M+2}\cap\aleph_{\beta_{N+1}}\>
        ,M+1
      )
    \)
    for some \(M<\omega\) and thus
    \(
      \alpha
        \in
      \sigma(
        \<C_{N+M+1},C_{N+M+2}\>,
        N+M+2
      )
    \). Thus \(\sigma\) is a winning \(2\)-mark.
  \end{proof}

  \begin{theorem}
    Let \(\alpha\) be the limit of increasing ordinals \(\beta_n\) for \(n<\omega\).
    If \(\pl F \kmarkwin2\schFillInitialGame{\aleph_{\beta_n}}\) for all
    \(n<\omega\), then \(\pl F\kmarkwin2\schFillInitialGame{\aleph_\alpha}\).
  \end{theorem}

  \begin{proof}
    The proof proceeds nearly identically to the previous proof: substitute
    \(\alpha\in C_0\) in place of \(\alpha\in\bigcap_{n<\omega}C_n\) and proceed.
  \end{proof}

  \begin{corollary}
    It is consistent that \(\alcompSP{\aleph_\omega}\) fails,
    but as \(\alcompSP{\aleph_k}\) holds in \(ZFC\) for all \(k<\omega\),
    both
    \(\pl F\kmarkwin2\schFillIntGame{\aleph_\omega}\) and
    \(\pl F\kmarkwin2\schFillInitialGame{\aleph_\omega}\) hold in \(ZFC\).
  \end{corollary}

  We conclude by returning our attention to Question \ref{mainQuestion},
  which asks whether there exists a space for which the second player
  \(\pl F\) in the game \(\menGame{X}\) has a winning strategy
  without a winning \(2\)-mark.

  \begin{question}
    Does
    \(\pl F\kmarkwin2\schFillIntGame{\kappa}\) hold for all cardinals
    \(\kappa\) in ZFC?
  \end{question}

  If not, the model producing \(\kappa>\aleph_\omega\) where
  \(\pl F\notkmarkwin2\schFillIntGame{\kappa}\) yields a positive
  answer to Question \ref{mainQuestion}:
  \(X=\oneptlind{\kappa}\). On the other hand, under \(V=L\)
  Corollary \ref{alcompSPunderV=L} shows that
  \(\alcompSP{\kappa}\) and therefore
  \(\pl F\kmarkwin2\menGame{\oneptlind\kappa}\)
  for every cardinal \(\kappa\), so a more exotic example than
  \(X=\oneptlind{\kappa}\) would be required to
  answer Question \ref{mainQuestion} in \(ZFC\).

  Solving the following weaker question would not answer Question
  \ref{mainQuestion} by itself, but a solution would be interesting nonetheless.

  \begin{question}
    Does
    \(\pl F\kmarkwin2\schFillGame{\kappa}\) hold for all cardinals
    \(\kappa\) in ZFC?
  \end{question}

  Although, it's still open whether the previous two questions are even
  distinct.

  \begin{question}
    Can a winning \(2\)-Markov strategy in \(\schFillGame{\kappa}\)
    be used to construct a winning \(2\)-Markov strategy in
    \(\schFillIntGame{\kappa}\)?
  \end{question}

\bibliographystyle{plain}
\bibliography{../../bibliography}

\end{document}