### 2018-02-23 postdoctoral/k-markov-tactic-banach-mazur.tex
almost finish poset banach mazur game result

\documentclass[11pt]{article}

\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amsthm}

\usepackage[letterpaper,margin=1in]{geometry}

\usepackage{enumerate}

      \theoremstyle{plain}
      \newtheorem{theorem}{Theorem}
      \newtheorem{lemma}[theorem]{Lemma}
      \newtheorem{corollary}[theorem]{Corollary}
      \newtheorem{proposition}[theorem]{Proposition}
      \newtheorem{conjecture}[theorem]{Conjecture}
      \newtheorem{question}[theorem]{Question}
      \newtheorem{claim}[theorem]{Claim}

      \theoremstyle{definition}
      \newtheorem{definition}[theorem]{Definition}
      \newtheorem{example}[theorem]{Example}
      \newtheorem{observation}[theorem]{Observation}
      \newtheorem{game}[theorem]{Game}

      \theoremstyle{remark}
      \newtheorem{remark}[theorem]{Remark}

      \theoremstyle{plain}
      \newtheorem*{theorem*}{Theorem}
      \newtheorem*{lemma*}{Lemma}
      \newtheorem*{corollary*}{Corollary}
      \newtheorem*{proposition*}{Proposition}
      \newtheorem*{conjecture*}{Conjecture}
      \newtheorem*{question*}{Question}
      \newtheorem*{claim*}{Claim}

      \theoremstyle{definition}
      \newtheorem*{definition*}{Definition}
      \newtheorem*{example*}{Example}
      \newtheorem*{observation*}{Observation}
      \newtheorem*{game*}{Game}

      \theoremstyle{remark}
      \newtheorem*{remark*}{Remark}

\title{Tactics and Marks in Banach Mazur Games}
\author{Steven Clontz}

\usepackage{../clontzDefinitions}

\newcommand{\bmPoGame}[2]{BM_{po}(#1,#2)}

\begin{document}

\maketitle

  \section*{marks and tactics}

  My notes on Galvin/Telgarsky's Theorem 5 from \cite{MR831181}.

  \begin{definition}
    Let \(\mb P\) be partially ordered by \(\leq\).
    Let \(\mb P^{\omega,\downarrow}=\{f\in\mb P^\omega : f(n)\geq f(n+1)\}\).
    Then for \(f,g\in\mb P^{\omega,\downarrow}\), we say that \(f,g\) zip into each
    other if for all \(m<\omega\) there exists \(n<\omega\) such that
    \(f(m)\geq g(n)\) and \(g(m)\geq f(n)\).
  \end{definition}

  \begin{definition}
    \(\bmPoGame{\mb P}{W}\) is a game defined for all non-empty partial orders
    \(\mb P\) and all subsets \(W\subseteq\mb P^{\omega,\downarrow}\).
    During round \(0\), \(\plI\) chooses \(a_0\in\mb P\),
    and then \(\plII\) chooses \(b_0\leq a_0\); during around \(n+1\),
    \(\plI\) chooses \(a_{n+1}\leq b_n\), and then \(\plII\) chooses
    \(b_{n+1}\leq a_{n+1}\). \(\plII\) wins this game if
    \(\<a_0,a_1,\dots\>\in W\).
  \end{definition}

  \begin{theorem}
    Let \(W\subseteq\mb P^{\omega,\downarrow}\) be closed under zipping.
    \(\plII\markwin\bmPoGame{\mb P}{W}\) if and only if
    \(\plII\tactwin\bmPoGame{\mb P}{W}\).
  \end{theorem}

  \begin{proof}
    Let \(\tau(p,n+1)\) be a winning mark for \(\plII\), where \(p\)
    is the most recent move by \(\plI\) and \(n+1\)
    is the number of moves made by \(\plI\).
    Define \(\tau^0(p)=p\) and \(\tau^{n+1}(p)=\tau(\tau^n(p),n+1)\).
    Let \(\preceq\) well-order \(\mb P\).

    For \(p,q\in \mb P\), say \(p\geq_n q\) if there exist
    \(s_m(p)\in\mb P\) for \(m\leq n\)
    such that
    \[
      p
    \geq
      s_m(p)
    \geq
      \tau(s_m(p),n+1)
    \geq
      q
    .\]
    Note that \(p'\geq p\geq_n q\geq q'\) implies \(p'\geq_n q'\),
    and \(p\geq_n \tau^n(p)\).

    Say \(p\geq_\omega q\) whenever \(p\geq_n q\) for all \(n<\omega\).
    If \(p\geq_\omega l(p)\) for some \(l(p)\), then say \(p\) is long;
    otherwise call \(p\) short.

    For \(p\) short, let
    \[
      \mu(p)
        =
      \min_{\preceq}\{
        r\text{ short}
          :
        r\geq p
      \}
    \]
    and since \(\mu(p)\not\geq_n p\) for some \(n\), let
    \[
      N(p)
        =
      \min\{
        n<\omega
      :
        \mu(p)\not\geq_n p
      \}
    .\]
    Note that whenever \(\mu(p)=\mu(q)\) for \(p\geq_n q\),
    it follows that \(\mu(p)\geq_n q\) and therefore \(N(p)<N(q)\).

    We define
    \[
      \sigma( p)
        =
      \begin{cases}
        l(p) & p \text{ is long} \\
        \tau^{N(p)+1}( p) &  p \text{ is short}
     \end{cases}
    .\]

    Suppose \(\sigma\) is legally attacked by \(a\in\mb P^\omega\).
    For \(n\leq\omega\), if \(a(n)\) is long, then \(a(n)\geq_n l(a(n))\).
    Therefore,
    \[
      a(n)
        \geq
      s_n(a(n))
        \geq
      \tau(s_n(a(n)),n+1)
        \geq
      l(a(n))
        =
      \sigma(a(n))
        \geq
      a(n+1)
    .\]
    Thus if \(a(n)\) is long for \(n<\omega\), it follows that
    \(c\in\mb P^{\omega,\downarrow}\) defined by \(c(n)=s_n(a(n))\)
    is a legal attack against \(\tau\). Since \(\tau\) is winning,
    \(c\in W\), and since \(c\) zips into \(a\),
    \(a\in W\) as well.

    Otherwise, we may choose a final subsequence \(b\) of \(a\) such that
      \begin{itemize}
        \item \(b(n)\) is short for all \(n<\omega\),
              since \(a(m)\) short implies \(a(n+m)\) short for all
              \(n<\omega\).
        \item \(\mu(b(n))=\mu'\) is fixed for all \(n<\omega\), since
              there cannot be an infinite \(\preceq\)-decreasing sequence.
      \end{itemize}
    As a result,
    \[
      b(n)
        \geq_{N(b(n))}
      \tau^{N(b(n))+1}(b(n))
        =
      \sigma(b(n))
        \geq
      b(n+1)
    \]
    and therefore \(N(b(n))<N(b(n+1))\). In particular, \(N(b(n))\geq n\).

    Thus for \(n<\omega\),
    \[
      b(n)
        \geq
      \tau^{n}(b(n))
        \geq
      \tau(\tau^{n}(b(n)),n+1)
        \geq
      \tau^{N(b(n))+1}(b(n))
        =
      \sigma(b(n))
        \geq
      b(n+1)
    .\]
    As a result, \(c\in\mb P^{\omega,\downarrow}\) defined by \(c(n)=\tau^n(b(n))\)
    is a legal attack against the winning strategy \(\tau\). Therefore
    \(c\in W\), and since \(c\) zips into \(b\) and \(a\), we conclude
    \(a\in W\).
  \end{proof}

  \begin{observation}
    When \(\mb P=T(X)\setminus\{\emptyset\}\) is ordered by set-inclusion
    and \(W=\{U\in\mb P^{\omega,\downarrow}:\bigcap_{n<\omega}U(n)\not=\emptyset\}\),
    then \(\bmPoGame{\mb P}{W}\) is exactly the topological Banach Mazur game
    \(\bmGame{X}\). Note \(W\) is closed under zipping.
  \end{observation}

  \begin{corollary}
  \(\plII\markwin\bmGame{X}\) if and only if
  \(\plII\tactwin\bmGame{X}\).
  \end{corollary}

  \section*{2+ marks and tactics}

  And this stuff is based on section 4.5.1 of \cite{bartoszynski1993covering}.

  \begin{definition}
    Let \(f\in S^{\leq\omega}\). Then \(f\rest n\in S^n\) is defined by
    \((f\rest n)(i)=f(i)\). (\(f\rest n\) gives the first \(n\) terms of \(f\).)

    Let \(t\in S^{<\omega}\). Then \(t\last k\in S^k\) is defined by
    \((t\last k)(i)=t(i+|t|-k)\). (\(t\last k\) gives the last \(n\) terms
    of \(t\).)
  \end{definition}

  \begin{definition}
    For every partial order \(\mb P\) and compatible \(p,q\in\mb P\),
    write \(p\not\perp q\) and let \(p\wedge q\) satisfy \(p\wedge q\leq p,q\).
    If \(p,q\) are incompatible, write \(p\perp q\).
  \end{definition}

  \begin{definition}
    For every partial order \(\mb P\) and compatible \(p\in\mb P\),
    let \(p^\downarrow=\{q\in\mb P:q\leq p\}\).
\end{definition}

  \begin{lemma}
    \(\mb P\) contains no infinite antichains if and only if every
    antichain in \(\mb P\) is of size \(n\) or less for some \(n<\omega\).
  \end{lemma}

  \begin{proof}
    This was shown to be true for \(\mb P=\tau\setminus\{\emptyset\}\) in
    Lemma 2.10 of \cite{comfort_negrepontis_1982}. It's likely known
    for general \(\mb P\), but I can't find a citation, so let's roll
    our own proof here. Assume \(\mb P\) has antichains of size \(n\)
    for all \(n<\omega\).

    Say \(p\in\mb P\) is bad if there exists \(r_p\leq p\) such that
    \(r_p^\downarrow\) is pairwise compatible. Let
    \(\mb P_{bad}\) collect all bad points in \(\mb P\), and say
    \(p\sim q\) for \(p,q\in\mb P_{bad}\) if \(r_p\not\perp r_q\).
    This is obviously symmetric and reflexive, and if we assume
    \(p\sim q,q\sim t\), then let \(s_p=r_p\wedge r_q\) and
    \(s_r=r_q\wedge r_t\). Since \(r_q^\downarrow\in\mb P_{bad}\),
    \(s_p\not\perp s_r\), so \(r_p\not\perp r_t\) and thus
    \(p\sim t\). Thus \(\sim\) is an equivalence relation.

    If \(\mb P_{bad}/\sim\) is infinite, we may choose \(p_i\in\mb P_{bad}\)
    such that \(p_i\not\sim p_j\) for \(i<j<\omega\). Thus
    \(r_{p_i}\perp r_{p_j}\) for \(i<j<\omega\), giving us
    an infinite antichain \(\{r_{p_i}:i<\omega\}\).

    Otherwise \(|\mb P_{bad}/\sim|=n<\omega\), and choose an antichain
    \(\{p_i:i\leq n\}\) in \(\mb P\). If \(\{p_i:i<n\}\subseteq\mb P_{bad}\),
    \(p_i\perp p_j\) implies \(r_{p_i}\perp r_{p_j}\)
    and \(p_i\not\sim p_j\) for all
    \(i\leq n\). Thus \(\mb P_{bad}=\bigcup_{i<n}\overset{\sim}{p_i}\),
    and \(p_n\not\in\mb P_{bad}\).

    So we've found \(b_0\in\mb P\setminus\mb P_{bad}\).
    Given \(b_n\in\mb P\setminus\mb P_{bad}\), we may choose
    \(a_n,b_{n+1}\leq b_n\) such that \(a_n\perp b_{n+1}\).
    Thus by construction, \(a_n\perp a_{m+1}\) for all
    \(n\leq m<\omega\). Therefore
    \(\{a_n:n<\omega\}\) is an antichain.
  \end{proof}

  \begin{proposition}
    Let \(W\subseteq\mb P^{\omega,\downarrow}\) be closed under zipping.
    Suppose every antichain in \(\mb P\) is of size \(n<\omega\) or less,
    and \(\plII\win\bmPoGame{\mb P}{W}\). Then
    \(\plII\autowin\bmPoGame{\mb P}{W}\) (i.e. \(\plII\) wins every
    play of \(\bmPoGame{\mb P}{W}\), i.e. \(W=\mb P^{\omega,\downarrow}\)).
  \end{proposition}

  \begin{proof}
    First, let \(\{p_i:i<n\}\) be an antichain of size \(n<\omega\),
    then let \(\mb P_i\) be a maximal pairwise-compatible subset of \(\mb P\)
    containing \(p_i\). Note that if there existed
    \(q\in\mb P\setminus\bigcup_{i<n}\mb P_i\),
    \(q\) must be incompatible with some \(q_i\in\mb P_i\) for \(i<n\).
    Since \(p_i,q_i\in\mb P_i\), they are compatible, so let
    \(r_i=p_i\wedge q_i\). Since \(q\) is incompatible with \(q_i\) for \(i<n\),
    \(q\) is incompatible with \(r_i\) for \(i<n\). Since \(p_i\) is
    incompatible with \(p_j\) for \(i<j<n\), \(r_i\) is incompatible with
    \(r_j\) for \(i<j<n\). But that makes \(\{q\}\cup\{r_i:i<n\}\) an
    antichain of size \(n+1\), contradicting the assumption of the proposition.
    Thus \(\mb P=\bigcup_{i<n}\mb P_i\).

    We now show that if \(s\in\mb P_i^\downarrow\) for some \(i\),
    then \(s\in W\). Let \(\sigma\) be a
    winning strategy for \(\plII\) in \(\bmPoGame{\mb P}{W}\),
    and attack \(\sigma\) with \(q(0)=s(0)\wedge p_i\) and
    \(q(n+1)=s(n+1)\wedge\sigma(\<q(0),\dots,q(n)\>)\). Note that the choice
    of \(q(0)\) is valid as \(s(0),p_i\in\mb P_i\). Similarly,
    \(\sigma(\<q(0),\dots,q(n)\>)\leq q(0)\leq p_i\), so
    \(\sigma(\<q(0),\dots,q(n)\>)\) cannot be compatible with any \(p_j\)
    where \(j\not=i\). Thus \(s(n+1),\sigma(\<q(0),\dots,q(n)\>)\in\mb P_i\),
    making the choice of \(q(n+1)\) valid. Since \(\sigma\) is winning for
    \(\plII\), we see that \(q\in W\), and therefore \(s\in W\).

    Finally, consider any play of \(\bmPoGame{\mb P}{W}\). It must contain
    have a subsequence \(s\in\mb P_i^\downarrow\) for some \(i<n\),
    so \(s\in W\) and therefore
    the play is also in \(W\), securing a victory for \(\plII\).
  \end{proof}

  \begin{lemma}
    Let \(W\subseteq\mb P^{\omega,\downarrow}\) be closed under zipping.
    Suppose that for every \(p\in\mb P\), there exists an infinite
    antichain \(A_p=\{a_p(n):n<\omega\}\subseteq\{q\in\mb P:q\leq p\}\).
    Then \(\plII\kmarkwin{(k+2)}\bmPoGame{\mb P}{W}\) if and only if
    \(\plII\ktactwin{(k+2)}\bmPoGame{\mb P}{W}\).
  \end{lemma}

  \begin{proof}
    The intuition of the following proof is simple: consider the case
    \(k=0\). During the first round,
    \(\plI\) plays some \(p_0\in\mb P\), and
    \(\plII\) can store the round number \(0\) (known by \(\plII\) since they
    only have knowledge of one move) by pretending \(\plI\) chose
    \(a_{p_0}(0)\leq p_0\) instead, and applying the winning \(2\)-mark.
    Thus when \(\plI\) plays \(p_1\leq a_{p_0}(0)\), \(\plII\) will have
    knowledge of both \(p_0\) and \(p_1\), and thus can observe that as
    \(p_1\leq a_{p_0}(0)\), it must be round \(1\) rather than some future
    round, and can repeat this process by pretending \(\plI\) chose
    \(a_{p_1}(1)\leq p_1\) and \(a_{p_0}(0)\leq p_0\) instead.

    We now proceed with a formal proof.
    Let \(\sigma\) witness \(\plII\kmarkwin{(k+2)}\bmPoGame{\mb P}{W}\).
    Define \(\tau(t)=\sigma(\<a_{t(0)}(0)\>,1)\) for \(t\in\mb P^1\).
    Since \(\tau(t)=\sigma(\<a_{t(0)}(0)\>,1)\leq a_{t(0)}(0)\leq t(0)\),
    this is a legal move.

    Consider \(t\in\mb P^{j+2}\) for \(j\leq k\).
    If there exists \(l_t<\omega\) such that
    \(t(j+1)\leq a_{t(j)}(l_t+j)\),
    define \(t'\in\mb P^{j+2}\) by \(t'(i)=a_{t(i)}(l_t+i)\) and let
    \(\tau(t)=\sigma(t',l_t+|t|)\). Note that since
    \[
      \tau(t)
        =
      \sigma(t',l_t+|t|)
        \leq
      t'(j+1)
        =
      a_{t(j+1)}(l_t+j+1)
        \leq
      t(j+1)
    \]
    this is a legal move. (If \(l_t\) failed to exist,
    we could arbitrarily let, say,
    \(\tau(t)=t(|t|-1)\); as we will see, this case will never occur
    for any legal attack against \(\tau\).)

    Let \(f\) be a legal attack against \(\tau\).
    We may quickly verify
    that \(l_{f\rest 2}=0\) since
    \begin{align*}
      (f\rest 2)(1)
        &=
      f(1)
        \\&\leq
      \tau(f\rest 1)
        \\&=
      \sigma(\<a_{f(0)}(0)\>,1)
        \\&\leq
      a_{f(0)}(0)
        \\&=
      a_{(f\rest 2)(0)}(0+0)
    \end{align*}
    We claim in general that \(l_{f\rest(j+2)}=0\) for \(j\leq k\).
    Assuming \(l_{f\rest(j+2)}=0\) for \(j<k\),
    \begin{align*}
      (f\rest(j+3))(j+2)
        &=
      f(j+2)
        \\&\leq
      \tau(f\rest(j+2))
        \\&=
      \sigma(f\rest(j+2)',0+(j+2))
        \\&\leq
      f\rest(j+2)'(j+1)
        \\&=
      a_{(f\rest(j+2))(j+1)}(0+(j+1))
        \\&=
      a_{(f\rest(j+3))(j+1)}(0+(j+1))
    \end{align*}
    proving \(l_{f\rest(j+3)}=0\).

    Now we show that \(l_{f\rest(n+2)\last(k+2)}=j-k\) for \(n\geq k\).
    We've just shown that this is true for our base case \(n=k\)
    since in that case \(f\rest(n+2)\last(k+2)=f\rest(k+2)\).
    Now assuming \(l_{f\rest(n+2)\last(k+2)}=n-k\) for some \(n\geq k\),
    we observe
    \begin{align*}
      (f\rest(n+3)\last(k+2))(k+1)
        &=
      f(n+2)
        \\&\leq
      \tau(f\rest(n+2)\last(k+2))
        \\&=
      \sigma((f\rest(n+2)\last(k+2))',(n-k)+(k+2))
        \\&\leq
      (f\rest(n+2)\last(k+2))'(k+1)
        \\&=
      a_{(f\rest(n+2)\last(k+2))(k+1)}((n-k)+(k+1))
        \\&=
      a_{(f\rest(n+3)\last(k+2))(k)}((n+1-k)+(k))
    \end{align*}
    and conclude \(l_{f\rest(n+3)\last(k+2)}=n+1-k\).

    Define \(g\in\mb P^{\omega,\downarrow}\) by \(g(0)=f(0)\) and
    \(g(j+1)=a_{f(j+1)}(j+1)\).
    Reviewing the above, the reader may confirm that we have shown for
    \(n<k+2\)
    \[
      f(n+1)
        \leq
      \tau(f\rest(n+1))
        =
      \sigma(g\rest(n+1),n+1)
        \leq
      g(n)
        \leq
      f(n)
    \]
    and for \(n\geq k+2\)
    \[
      f(n+1)
        \leq
      \tau(f\rest(n+1)\last(k+2))
        =
      \sigma(g\rest(n+1)\last(k+2),n+1)
        \leq
      g(n)
        \leq
      f(n)
    \]
    Thus \(g\) is a legal attack against \(\sigma\), and since \(\sigma\)
    is winning, \(g\in W\). Since \(W\) is closed under zipping,
    \(f\in W\), and therefore \(\tau\) is also winning.
  \end{proof}

  \bibliographystyle{plain}
  \bibliography{../bibliography}

\end{document}