### 2018-03-27 postdoctoral/go-first.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amsthm}
\begin{document}

\begin{theorem}
	Suppose \(T\) be a formula representing a gofirst-fair set of \(n+1\) dice
	extending a formula \(S\) for \(n\) dice such that for each symbol in
	\(T\) representing a face of the extra die, \(S\) restricted below that
	symbol is permutation-fair. Then \(T\) is permutation-fair.
\end{theorem}

\begin{proof}
	Since gofirst-fair implies permutation-fair in the base case of one
	(or two, or zero, whatever) dice, assume the theorem holds for formulas
	\(T\) of length \(n+1\), and let \(T'\) be length \(n+2\) extending \(S'\).
\end{proof}
\end{document}